% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Spring 1990}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Work 9 of the following problems. Do not turn in more than 9.

For a subset $A$ contained in a topological space $X$, let $A^-$ denote
the closure of $A$ in $X$, let $A^\circ$ denote the interior of $A$ in $X$,
and let $Fr(A)$ denote the frontier of $A$ in $X$. All Euclidean spaces will
have the usual topology, and all product spaces will have the product
topology.

\begin{description}
\item[1.]
For a topological space $X$, prove that ``o--" and ``--o" are both idempotent
operations on the subsets of $X$; that is, $A^{o-o-} = A^{o-}$ and
$A^{-o-o} = A^{-o}$ for all $A \subseteq X$.

\item[2.] (a)
Define ``net".

\item[\quad] (b)
Give an example of a net that is not a sequence.

\item[3.]
Prove that $[0,1]$ (with its usual topology) is compact.

\item[4.] (a)
Characterize the compact subsets of $\R$.

\item[\quad] (b)
State and prove a maximum value theorem from calculus.

\item[\quad] (c)
Using maximum, minimum, and intermediate value theorems from calculus, prove
every continuous open function $f: [0,1] \to [0,1]$ is surjective.

\item[5.]
Prove that there is no smallest ($\subseteq$ - minimal) base for the usual
topology on $\R$.

\item[6.]
Prove that if $e: X \to Y$ is an embedding and $h: Y \to Z$ is a homeomorphism,
then $h \circ e$ is an embedding.

\item[7.]
Give an example of two topologies $\sigma$ and $\tau$ on the set of integers
$\Z$ for which $\sigma \subsetneq \tau$ and $(\Z, \sigma)$ is
homeomorphic to $(\Z, \tau)$.

\item[8.] (a)
Prove that the product of two connected spaces is connected.

\item[\quad] (b)
Is the product of uncountably many connected spaces necessarily connected?

\item[9.]
Prove that if $X \stackrel{f}{\longrightarrow} Y$ and
$Y \stackrel{g}{\longrightarrow} Z$ are quotient maps, then
$X \stackrel{g \cdot f}{\longrightarrow} Z$ is a quotient map.

\item[10.]
Show that $\Q \times \Q \cap S^1$ is dense in $S^1$.

\item[11.]
Prove that if $A$ is a compact subset of a regular (non Hausdorff) space
$X$, then $A^-$ is compact.

\item[12.]
Prove that any uncountable subset of $\R$ has a limit point.

\item[13.]
Give statements of the following theorems:

\item[\quad] (a)
Urysohn's Lemma.

\item[\quad] (b)
Urysohn's Metrization Theorem.

\item[\quad] (c)
Tychonoff's Theorem.

\item[\quad] (d)
Baire Category Theorem.

\item[14.]
Prove that the circle $(S^1)$ is not a retract of the plane $(\R^2)$.

\item[15.]
Prove or disprove that every separable space is second countable.

\item[16.]
Prove that the following are equivalent for any topological space $Y$:

\item[\quad] (a)
$Y$ is Hausdorff.

\item[\quad] (b)
The diagonal $\Delta_Y = \{(y,y) |y \in Y\}$ is closed subset of $Y \times Y$.

\item[\quad] (c)
For each space $Z$ and each pair of continuous functions $f,g: Z \to Y$
that agree on a dense subset of $Z$, it follows that $f = g$.





\end{description}    
%\end{Large}
\end{document}














