% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}


\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Spring 1993}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

\begin{description}
\item[1.]
Prove that the interval [0,1] with its usual topology is compact.

\item[2.]
Prove or disprove that a closed subspace of a normal space must be normal.

\item[3.]
Prove or disprove that if $X$ and $Y$ are any sets, $f : X \to Y$ is any
function, $A \subseteq X$, and $B \subseteq Y$, then
$f[A \cap f^{-1} [B]] = f[A] \cap B$.

Use the following definition for problems 4 and 5.

For any topological property $P$ a space is $X$ called {\bf locally} $P$
if and only if every point $x$ of $X$ has a base of neighborhoods each of
which has property $P$ (as a subspace).

\item[4.]
Prove that every quotient of a locally pathwise connected space is locally
pathwise connected.

\item[5.] (a)
Prove that every locally $T_1$ space is $T_1$.

\item[\quad] (b)
Prove that a locally $T_2$ space need not be $T_2$.

\item[6.]
Prove that a space $X$ is Hausdorff if and only if the diagonal
$\Delta \subseteq X \times X$ is closed.

\item[7.]
Let $X$ be a regular space and assume that $\mathcal U$ is a countable open
covering of $X$ with the property that the closure of each of its members is
paracompact. Prove that $X$ is paracompact.

\item[8.]
State the Axiom of Choice and prove that it is equivalent to Statement $S$
below.

{\bf Statement} $S$: The product of any set (including the empty set) of
nonempty sets is nonempty.

\item[9.]
Prove that if $A \times B$ is a compact subset of the space
$X \times Y$ that is contained in an open set $U$ of $X \times Y$, then there
is a ``rectangular" open set of the form $V \times W$ such that
$A \times B \subseteq V \times W \subset U$.

\item[10.]
Prove that if $A$ is a retract of a Hausdorff space $X$, then $A$ is closed
in $X$.

\item[11.] (a)
Prove that in a first-countable Hausdorff space $X$, every one point set is a
countable intersection of open subsets of $X$.

\item[\quad] (b)
Give an example of a space in which every one point set is a countable
intersection of open sets, but which is not first-countable.

\item[12.]
Let $X$ be a completely regular space. Prove that $X$ is connected if and
only if its Stone-\u Cech compactifiction $\beta X$ is connected.





\end{description}    
%\end{Large}
\end{document}














