% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}


\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Fall 1992}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

\begin{description}
\item[1.]
Prove that the interval [0,1] with the usual topology is connected.

\item[2.]
Prove or disprove: Every compact Hausdorff space is separable.

\item[3.]
Let $\pi_1 : X \times Y \to X$ be the projection map, and let $Y$ be
compact. Prove $\pi_1$ is a closed map (where $X \times Y$ has the product
topology).

\item[4.]
Prove that a paracompact Hausdorff space is regular.

\item[5.]
Use Zorn's lemma to prove that for every set $X$ and relation $R$ there is
maximal $A \subseteq X$ such that $A \times A \subseteq R$.

\item[6.]
Prove that the plane $\R^2$ is not a countable union of straight lines.

\item[7.]
If $\Omega$ is the first uncountable ordinal, prove that $[0, \Omega]$
with the order topology is compact.

\item[8.]
Prove that a connected normal Hausdorff space containing more than one point
is uncountable.

\item[9.]
Let $A$ be a connected subset of a connected space $X$, and let $C$ be a
component of $X-A$. Prove that $X-C$ is connected.

\item[10.]
Let $C(X,Y)$ be the set of continuous functions from $X$ to $Y$, given the
compact-open topology. Let $e : C(X,Y) \times X \to Y$ be the evaluation
map $e(f,x) = f(x)$. Prove that if $X$ is locally compact Hausdorff, then
$e$ is continuous.

\item[11.]
Let $f : X \to Y$ be a continuous surjective map from a compact space $X$ to
a Hausdorff space $Y$. Prove that $f$ is a quotient map.

\item[12.]
Find an incorrect statement in the proof of the following theorem and prove
that it is an incorrect statement.

{\bf Theorem.}
If $B^2 = \{(x_1, x_2) \in \R^2 |x_1^2 + x^2_2 \leq 1\}$
has the usual topology, then each continuous function $f : B^2 \to B^2$
has a fixed point.

{\bf Proof:}
Suppose that $f : B^2 \to B^2$ is a continuous function with no fixed points.
Let $\pi_1: B^2 \times B^2 \to B^2$ be first projection, let
$$\Delta = \{(z,z) |z \in B^2\}$$
be the diagonal in $B^2 \times B^2$, and let
$$F = \{(z, f(z)|z \in B^2\}$$
be the graph of $f$. Since $f$ is continuous, $\pi_1|_F : F \to B^2$ is a
homeomorphism. Since $B^2$ is connected, $F$ is therefore connected. However,
$$F \subseteq B^2 \times B^2 - \Delta$$
which is a contradiction since $B^2 \times B^2 - \Delta$ is disconnected.

\item[13.]
Prove that the topologist's comb
$$C= \left( \bigcup^\infty_{n=1} \left\{\frac{1}{n} \right\} \times
  [0,1] \right) \cup \left( [0,1] \times \{0\} \right) \cup
  \left( \{0\} \times [0,1] \right)$$
is {\bf not} a retract of the square $S=[0,1] \times [0,1]$
(with both $C$ and $S$ having the usual Euclidean subspace topologies).





\end{description}    
%\end{Large}
\end{document}














