% latex file
%Topology QE F98 - Maginnis & Strecker

\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)

\input prepictex.tex
\input pictex.tex
\input postpictex.tex

%\documentstyle[bbb]{report}
\documentclass[bbb]{report}

\vcorrection{-.7in}
\hcorrection{-1in}

\topmargin  0in
\textwidth 6.4in
\textheight 9.5in

\mathsurround=2pt

\def\ds{\displaystyle}
\def\R{{\Bbb R}}

\begin{document}

%\pagestyle{empty}

\begin{Large}

\hfill Name \rule{2.5in}{.01in}
\par
\vspace{.1in}

\begin{center}
   TOPOLOGY QUALIFYING EXAM \\
   Fall 1998 \\
   (Maginnis and Strecker) \\
\end{center}

\end{Large}

\vspace{.1in}

\begin{large}
\noindent {\bf Choose and work any 6 of the following 15 problems.
Start each problem on a new sheet of paper.
\underbar{Do not turn in more than six problems}.
In the problems below, a space always means a topological space.}

\vspace{.2in}

\begin{description}

\item[1.]
Prove or disprove:
For any space $X$ with topology ${\cal J}$,
the family 
${\cal B}=$
\linebreak $\{ A\subseteq X\,\bigg|\,A$ equals the interior of its
closure\} forms a base for some topology ${\cal J}'$ on $X$.

\vspace{.5in}

\item[2.]
Prove that a quotient space of a locally connected space is
locally connected.

\vspace{.5in}

\item[3.]
Let $f:X\to Y$ be an open map between the spaces $X$ and $Y$.
Let $B\subseteq Y$ and $A=f^{-1}[B]$.
Prove that the restriction $\overline{f}:A\to B$
(i.e., $\overline{f}(a)=f(a)$) is an open map
from $A$ to $B$.

\vspace{.5in}

\item[4.]
(a) State the Axiom of Choice.

\vspace{.1in}
\item[\quad (b)] State the Well-Ordering Theorem.

\vspace{.1in}
\item[\quad (c)] Either use the Axiom of Choice to prove the
Well-Ordering Theorem or use the Well-Ordering Theorem to prove the
Axiom of Choice.

\vspace{.5in}
\item[5.] Prove that the plane ${\Bbb R}^2$ with its usual
topology is not equal to a countable union of straight lines.

\vspace{.5in}
\item[6.]
Prove that the Sorgenfrey line $X={\Bbb R}$ with basis
$\{[a,b)\,\bigg|\,a,b\in{\Bbb R}\}$ is a paracompact space.

\vspace{.5in}
\item[7.]
Let $f:[a,b]\to{\Bbb R}$ be a real-valued function on a
closed interval and let
$G=\{(x,f(x))\in{\Bbb R}\times\R
\,\bigg|\, a\leq x\leq b\}$ be its graph.
Prove or give a counterexample for the following.

\vspace{.1in}
\item[\quad (a)] If $G$ is connected, then $f$ is continuous.

\vspace{.25in}
\item[\quad (b)] If $f$ is continuous, then $G$ is connected.


\vspace{.5in}
\item[8.]
Prove that the net based on an ultrafilter is an ultranet.

\vspace{.5in}
\item[9.]
Let $Y$ be a compact space. Prove that the projection map
$\pi_1:X\times Y\to X$ is a closed map.

\vspace{.5in}
\item[10.]
Let $A$ be a subset of a complete metric space $X$.
Prove that $A$ is totally bounded if and only if the closure
$\overline{A}$ is compact.

\vspace{.5in}
\item[11.]
Let $X$ be a completely regular $T_1$ space
(i.e., one point sets are closed, and for each closed set $C$
and point $x\not\in C$, there exists a continuous function
$f:X\to[0,1]$ with $f(x)=1$ and $f[C]=\{0\}$).
Prove that the Stone-\v Cech compactification $\beta(X)$ is
connected if and only if $X$ is connected.

\vspace{.5in}
\item[12]
Let $f:X\to Y$ be a continuous surjective map from a compact space
$X$ to a Hausdorff space $Y$. Prove that $f$ is a quotient map.

\vspace{.5in}
\item[13.] Let $D$ be a dense subset of a metric space $X$,
and let $Y$ be a complete metric space.
Prove that any uniformly continuous function
$f:D\to Y$ can be extended to a uniformly continuous function
$F:X\to Y$ (i.e., $F|_D=f$).

\vspace{.5in}
\item[14.]
Prove that a metric space is compact if and only if every sequence
has a convergent subsequence.

\vspace{.5in}
\item[15.]
Prove that each metric space is a normal space.

\vfill
\clearpage


\end{description}
\end{large}
\end{document}
