% latex file
%Topology QE F99 - Maginnis & Strecker

\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)

\input prepictex.tex
\input pictex.tex
\input postpictex.tex

%\documentstyle[bbb]{report}
\documentclass[bbb]{report}

\vcorrection{-.7in}
\hcorrection{-1in}

\topmargin  0in
\textwidth 6.4in
\textheight 9.5in

\mathsurround=2pt

\def\ds{\displaystyle}
\def\R{{\Bbb R}}
\def\N{{\Bbb N}}

\begin{document}

%\pagestyle{empty}

\begin{Large}

\hfill Name \rule{2.5in}{.01in}
\par
\vspace{.1in}

\begin{center}
   TOPOLOGY QUALIFYING EXAM \\
   Fall 1999 \\
   (Maginnis and Strecker) \\
\end{center}

\end{Large}

\vspace{.1in}

\begin{large}
\noindent {\bf Choose and work any 6 of the following 14 problems.
Start each problem on a new sheet of paper.
\underbar{Do not turn in more than six problems}.
In the problems below, a space always means a topological space.}

\vspace{.2in}

\begin{description}

\item[1.]
Prove or give a counterexample:
 If  $f:X\to Y$ is an injective continuous function and $Y$
 is a Hausdorff space, then $X$ is a Hausdorff space.

%\vspace{.15in}

\item[2.]
Let $f:X\to Y$ be a quotient map such that for all $y\in Y$,
$f^{-1}[\{y\}]$ is connected. Let $W\subseteq Y$ be an
open connected subset. Prove $f^{-1}[W]$ is connected.
%\vspace{.15in}


\item[3.]
A space $X$ is called an $R_0$-space if for every pair of points
$\{x,y\}\subseteq X$, $x$ is in the closure $\overline{ \{y\} }$
if and only if $y$ is in the closure $\overline{\{ x\} }$.
Prove or give a counterexample for each of the following statements.
\item[\quad] a) Every $T_0$-space is $R_0$.
\item[\quad] b) Every $R_0$-space is $T_0$.
\item[\quad] c) A space is $T_1$ iff it is both $T_0$ and $R_0$.

%\vspace{.15in}

\item[4.]
Let $\R$ be the real numbers with the usual topology.
Prove that any open set in $\R$ is a countable union of pairwise disjoint
intervals.

%\vspace{.15in}
\item[5.]
Let $\omega_1$ be the first uncountable ordinal number, let $X$ be the
set of all ordinals less than or equal to $\omega_1$,
and let $Y=X-\{\omega_1\}$. ($Y$ is often called the minimal uncountable
well-ordered set).
\item[\quad] a) Prove that \underbar{no} sequence in $Y$ converges
to $\omega_1$.
\item[\quad] b) Give \underbar{either} a net in $Y$ \underbar{or}
a filter in $Y$ that converges to $\omega_1$.

%\vspace{.15in}
\item[6.]
Let $X$ be a compact Hausdorff space, and let $f:X\to X$ be a
continuous function. Prove that there exists a nonempty subset
$A\subseteq X$ such that $f[A]=A$.


%\vspace{.15in}
\item[7.]
Let $X$ be a normal (and Hausdorff) space with a countable basis.
Prove that every subspace of $X$ is normal.

%\vspace{.15in}
\item[8.]
Prove or give a counterexample for each of the following statements.
\item[\quad] a) A compact subset of a $T_1\hbox{-space}$ is closed.
\item[\quad] b) A closed subset of a paracompact space is paracompact.

%\vspace{.15in}
\item[9.]
Let $\R$ be the real numbers with the usual topology.
Prove there does \underbar{not} exist a function $f:\R\to\R$ which
is continuous precisely at the rational numbers.
(i.e., continuous at the rationals and discontinuous at the irrationals.)

%\vspace{.15in}
\vfill\pagebreak

\item[10.]
Let $f:X\to Y$ be a continuous function.
\item[\quad] a)
Prove that there exists a  space $Z$, a continuous surjection
$s:X\to Z$, and an embedding $e:Z\to Y$ such that $f$ is the
composite $F=e\circ s$.
\item[\quad] b)
Prove that if $f=e\circ s$ and $f=\hat e\circ\hat s$, where
$X\buildrel s\over\to Z\buildrel e\over\to Y$
and
$X\buildrel{\hat s}\over \to \hat Z \buildrel{\hat e}\over \to Y$,\
are two factorizations of $f$ as a composite of a continuous
surjection and an embedding, then there exists a unique
homeomorphism $h:Z\to\hat Z$ with $h\circ s=\hat s$ and
$\hat e\circ h=e$.

%\vspace{.15in}
\item[11.]
Let $X$ be a Hausdorff space such that every infinite subset
has a limit point. Prove that every countable open cover of $X$ has
a finite subcover.

%\vspace{.15in}
\item[12]
Let $X$ be the countable product $X=\R^\N=\ds\prod^\infty_{n=1}\R$,
where $\R$ is the real numbers with the usual topology.
Prove that $X$, with the product topology, is locally connected but
is not locally compact.

%\vspace{.15in}
\item[13.]
Let $X$ and $Y$ be spaces which are homotopy equivalent, and assume
$X$ is connected. Prove $Y$ is connected.
(Recall that $X$ and $Y$ are \underbar{homotopy equivalent} if there
exist four continuous functions
$f:X\to Y$, $g:Y\to X$, $H_1:X\times I\to X$, and $H_2:Y\times I\to Y$\,
where $I$ is the interval $[0,1]$, satisfying
$H_1(x,0)=g(f(x))$, $H_1(x,1)=x$, $H_2(y,0)=f(g(y))$,
and $H_2(y,1)=y$, for all $x\in X$ and all $y\in Y$.)

\vspace{.15in}
\item[14.]
Prove or give a counterexample for each of the following:

\item[\quad] a)
Every bounded subset of a well ordered set has both a greatest
lower bound and a least upper bound.

\item[\quad] b)
Every linearly ordered set (also called totally ordered) in which
every bounded subset has both a greatest lower bound and a least upper
bounded is well ordered.

\item[\quad] c)
Every linearly ordered subset of a partially ordered set has
an upper bound.

\item[\quad] d)
In any well ordered set, the minimal element is the only element
not having an immediate predecessor.


\end{description}
\end{large}
\end{document}
