% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
%\begin{Large}

 
\begin{center}\begin{LARGE}
{\bf Topology   Qualifying Exam}\\ 
{\bf Fall 1996}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Choose and work any 6 of the following 14 problems. Start each problem
on a new sheet of paper. {\bf Do not turn in more than six problems}. A
space always means a topological space below.

\begin{description}

\item[1.]
Prove that the unit interval $[0,1]$ is connected.

\item[2.]
Prove that if $X$ is a space with the property that each filter on $X$
has a cluster point, then $X$ is compact.

\item[3.]
Prove or disprove: There exists an infinite strictly decreasing sequence of
ordinal numbers.

\item[4.]
Prove or disprove:

\item[\quad] (a)
Every path connected space is connected.

\item[\quad] (b)
Every connected space is path connected.

\item[5.]
Let $f:X \to Y$ be a surjective continuous function with compact Hausdorff
domain. Prove that $Y$ is Hausdorff if and only if $f$ is closed.

\item[6.]
For a topological space $(X, \tau)$ define $I: P (X) \to P(X)$ by
$I(A) = A^\circ$ and $C : P(X) \to P(X)$ by $C(A) = \overline A$.
For a set $A$ of $X$, consider the sequence:
$A, I(A), CI(A), ICI(A), CICI(A), \dots$

\item[\quad] (a)
For any space, what is the largest number of distinct sets that this
sequence can contain?

\item[\quad] (b)
Find $A \subseteq \Bbb R$ for which this largest number is obtained (where
$\Bbb R$ has the usual topology).

\item[7.]
Prove or disprove: The product of two normal spaces is normal.

\item[8.]
Prove that a product of spaces $\Pi_{\alpha \in I} X_\alpha$ is locally
connected if and only if each space $X_\alpha$ is locally connected and all
but finitely many of the $X_\alpha$ are connected.

\item[9.]
Prove that a metric space is compact if and only if it is complete and
totally bounded.

\item[10.]
Prove that if $f: X \to Y$ is a continuous function between
Hausdorff spaces with $f[X]$ dense in $Y$, then $|Y| \leq 2^{2|x|}$.

\item[11.]
Let $f: X \to Y$ be a continuous open surjection. Assume that $X$ is a
Baire space. Prove that $Y$ is a Baire space.

\item[12.]
Prove or disprove: There is a completely regular $T_1$ space $X$ such that
the \v Cech-Stone compactification of $X$ is homeomorphic to the
Alexandroff compactification of $X$.

\item[13.]
Find an error in the following purported proof that ``Every compact $T_2$
space is metrizable."

{\bf Proof} If $X$ is compact $T_2$ it is locally compact $T_2$ and so it
is completely regular. Since $X$ is completely regular, it can be embedded
in a Hilbert cube, $I^{\aleph_0}$. But $I^{\aleph_0}$ is metrizable, and
metrizability is hereditary. Thus, since $X$ is homeomorphic to a metrizable
space, it must be metrizable.

\item[14.]
Prove that every space that is paracompact and countably compact must be
compact.




    
\end{description}
%\end{Large}
\end{document}














