%From strecker%math.ksu.edu@KSUVM.KSU.EDU Wed Feb 22 08:31:00 1995
%Date: Tue, 21 Feb 95 15:25:29 CST
%From: George Strecker <strecker%math.ksu.edu@KSUVM.KSU.EDU>
%To: math@ksuvm

% Here is an edited version of what you sent me.  Please
%look it over.  The only other change that I think I might
%make would be to put more of the easy problems earlier in the exam.  Geo.
% LaTeX file


%\documentstyle[12pt]{article}
\documentclass[12pt]{article}
\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)


\renewcommand{\labelenumi}
{\mbox{ \mbox{{\rm {\roman{enumi}}}})}}
\renewcommand{\theenumi}{\roman{enumi}}
%This changes the numbering in itemizing to roman.


\begin{document}

 \begin{center}

 {\bf TOPOLOGY QUALIFYING EXAM \ \ Fall 1994}

(Wu and Strecker)

 \end{center}


        Choose and work any 6 of the following 15 problems. Start each
 problem on a new sheet of paper. {\bf Do not turn in more than six
 problems.}
A space always mean a topological space below.

\vspace{10pt}
\noindent {\bf 1.}
State the axiom of choice and give another statement which is equivalent to
the axiom of choice. Then prove one of the implications for the equivalence.

\vspace{10pt}
\noindent {\bf 2.}  Prove that the following statements are equivalent
for any space $X$:
\begin{enumerate}
  \item $X$ is Hausdorff;
  \item for any space $Y$ and any pair of continuous maps
        $f, g : Y\rightarrow X$ that agree on a dense subset of $Y$,
        then $f=g$.
  \item the diagonal $\Delta\subseteq X \times X$ defined by
        $\Delta := \left\{ (x_1, x_2) \in X\times X \ |\  x_1=x_2 \right\} $
        is a closed subset of $X\times X$;
\end{enumerate}

\vspace{10pt}
\noindent {\bf 3.}
Prove or give a counterexample for each of the following implications: \\
separable $\stackrel{(a)}{\Rightarrow}$ second countable
 $\stackrel{(b)}{\Rightarrow}$ first countable  $\stackrel{(c)}{\Rightarrow}$
separable  $\stackrel{(d)}{\Rightarrow}$ first countable
 $\stackrel{(e)}{\Rightarrow}$ second countable  $\stackrel{(f)}{\Rightarrow}$
separable

\vspace{10pt}
\noindent {\bf 4.}
Prove that every quotient of a locally connected space is locally connected.

\vspace{10pt}
\noindent {\bf 5.}
Give a proof or counterexample for each of the following implications:
\begin{enumerate}
  \item $X$ is locally connected $\Rightarrow$ the connected components of $X$
      are open;
  \item  the connected components of $X$ are open
         $\Rightarrow$ $X$ is locally connected.
\end{enumerate}

\vspace{10pt}
\noindent {\bf 6.}
Let $\{ X_{\alpha}\}_{\alpha\in A}$ be a collection of spaces and let
$x$ be a fixed point in $\prod_{\alpha\in A} X_{\alpha}$. Show that
$$ D =\left\{ y\in \prod_{\alpha\in A} X_{\alpha} \mid
             y\ {\rm and}\ x\ {\rm differ\ in\ at\ most\ finitely\
             many\ coordinates} \right\} $$
is dense in $\prod_{\alpha\in A} X_{\alpha}$.

\vspace{10pt}
\noindent {\bf 7.} Prove or disprove that any two spaces that are
simultaneously discrete and indiscrete must be homeomorphic.

\vspace{10pt}
\noindent {\bf 8.}
Let $X$ be a $T_1$ regular space with a countable basis and
let $U$ be open in $X$. Show that
\begin{enumerate}
  \item $U$ is a countable union of closed sets of $X$;
  \item There is a continuous function $f: X\rightarrow [0, 1]$ such that
        $f(x)>0$ for $x\in U$ and $f(x)=0$ for $x\not\in U$.
\end{enumerate}

\vspace{10pt}
\noindent {\bf 9.}
Let $X$ be a $T_1$ completely regular space and $Y$ a
compactification of $X$; let $\beta(X)$ be the Stone-\v{C}ech
compactification of $X$. Show that there exists a continuous surjective closed
map $g: \beta(X)\rightarrow Y$ that equals the identity on $X$.

\vspace{10pt}
\noindent {\bf 10.}
Prove that the following statements are equivalent for any space $X$:
\begin{enumerate}
  \item $X$ is Tychonoff (i.e., $T_1$ and completely regular);
  \item $X$ can be embedded in a compact Hausdorff space.
\end{enumerate}

\vspace{10pt}
\noindent {\bf 11.}
Find an error in the following purported ``proof''
of the jactitation that $2^{\Bbb R}$
is metrizable, where $2=\{0, 1\}$ is a two point discrete space and
${\Bbb R}$ is the set of real numbers.

\noindent
``\underline{\sc Proof}'':
Consider the inclusion ${\Bbb N}\subseteq {\Bbb R}$, where
${\Bbb N}$ is the set of natural numbers. This induces a ``natural''
embedding $2^{\Bbb N}\hookrightarrow 2^{\Bbb R}$ by the map
$f\mapsto \hat{f}$ where
$$ \hat{f}(r)=\left\{ \begin{array}{ll}
    f(r),     & {\rm if}\ r\in {\Bbb N} \\
    0,        & {\rm if}\ r\in {\Bbb R}\setminus {\Bbb N}.
   \end{array} \right.  %}
$$
% But $2^{\Bbb N}$ is the Cantor space $C$. Thus $C$ is embedded in
% $2^{\Bbb R}$.
Let $U$ be an open subset of $2^{\Bbb R}$.
By the definition of the product topology, the projection of $U$
is $\{ 0, 1\}=2$ in all but finitely many coordinates.
Thus $U\cap 2^{\Bbb N}\not =\emptyset $. So $2^{\Bbb N}$ is dense in
$2^{\Bbb R}$. But $2^{\Bbb N}$ is compact and $2^{\Bbb R}$ is Hausdorff,
so $2^{\Bbb N}$ is closed in $2^{\Bbb R}$.
Thus $2^{\Bbb N}=2^{\Bbb R}$. But $2^{\Bbb N}$ is metrizable
since it is a countable product of metrizable spaces.
Hence $2^{\Bbb R}$ is metrizable.

\vspace{12pt}
\noindent {\bf 12.}
Let ${\frak A}$ be a locally finite collection of
subsets in the space $X$. Show that
$$ \overline{\bigcup_{A\in {\frak A}} A}
        = {\bigcup_{A\in {\frak A}} \overline{A}} . $$

\vspace{12pt}
\noindent {\bf 13.}
Prove or disprove each of the following:
\begin{enumerate}
  \item Every compact subset of a Hausdorff space is closed;
  \item every closed subset of a Hausdorff space is compact.
\end{enumerate}


\vspace{12pt}
\noindent {\bf 14.}
Let $E^n$ denote the Euclidean $n$-dimensional space. For what value
of $n$ is it true that $E^n\setminus \{p\}$ is connected and simply
connected? Prove your answer.

\vspace{12pt}
\noindent {\bf 15.}
Show that in a locally compact Hausdorff space, countable intersections of
dense open sets are dense.

\end{document}





