% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology   Qualifying Exam}\\ 
{\bf Fall 1997}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Choose and work and 6 of the following problems. Start each new problem
on a new sheet of paper. Do not turn in more than six problems. Below
a ``space" always means a ``topological space".

\begin{description}

\item[1.]
Prove or disprove:

\item[\quad] (a)
Closed subspaces of path connected spaces are path connected.

\item[\quad] (b)
If $f: X \to Y$ is continuous and $X$ is path connected, then $f[X]$
is path connected.

\item[2.]
Let ${\cal A}$ be a collection of subsets of the topological space $X$
such that $X = \cup {\cal A}$. Consider the function $f: X \to Y$; suppose
that $f|A$ is continuous for each $A \in {\cal A}$.

\item[\quad] (a)
Show that if ${\cal A}$ is finite and each member of ${\cal A}$ is closed,
then $f$ is continuous.

\item[\quad] (b)
Give an example to show that the word ``finite" in part (a) cannot be
changed to ``countable".

\item[3.]
Let $A$ and $B$ be disjoint compact subsets in the Hausdorff space $X$.
Show that there are disjoint open subsets $U$ and $V$ of $X$ such that
$A \subseteq U$ and $B \subseteq V$.

\item[4.]
Let $Y$ be an ordered set with the order topology. Let $f, g :X \to Y$ be
continuous.

\item[\quad] (a)
Let $h : X \to Y$ be the function given by
$$h(x) := \min \{f(x), g(x)\}.$$
Show that $h$ is continuous.

\item[\quad] (b)
Show that the set $\{x \in X|f (x) \leq g(x)\}$ is closed in $X$.

\item[5.]
Let $X$ be a complete metric space and $f: X \to \Bbb R$ a continuous real-valued
function on $X$. Show that every nonempty open subset of $X$ contains a
nonempty open subset on which $f$ is bounded.

\item[6.]
Let $f: X\to Y$ be a continuous surjective map, where $X$ is compact and
$Y$ is Hausdorff. Show that $f$ is a quotient map.

\item[7.]
A space $X$ is said to be {\it completely regular} if one-point sets are
closed and if for each point $x_0$ and each closed subset $A$ not containing
$x_0$, there is a continuous function $f: X \to [0,1]$ such that
$f(x_0) = 1$ and $f[A] \subset \{0\}$.

Show that every locally compact Hausdorff space is completely regular.

\item[8.]
If $f: X \to Y$ and $g: Y \to X$ are continuous functions such that $g o f$
is the identity function on $X$, prove that $f$ is topological embedding
and that $g$ is a quotient map.

\item[9.]
If $f$ and $g$ are real-valued continuous functions with the same domain,
prove that $f+g$ is continuous, where $(f+g) (x) \equiv f(x) + g(x)$ for
any $x$ in the domain.

\item[10.]
Prove that a filter ${\cal G}$ on a set $X$ is an ultrafilter if and only if for
each subset $A$ of $X$, either $A \in {\cal G}$ or
$X \backslash A \in {\cal G}$.

\item[11.]
Prove or disprove:

\item[\quad] (a)
Every compact subset of a Hausdorff space is closed.

\item[\quad] (b)
Every closed subset of a Hausdorff space is compact.

\item[12.]
Show that a metrizable space $X$ has a countable dense subset if and only
if it has a countable basis.

\item[13.]
Prove or disprove that closed subspaces of normal spaces are normal.

\item[14.]
Let $Y$ be a metric space and let $f_n: X \to Y$ be a sequence of continuous
functions and $f:X \to Y$ a (not necessarily continuous) function. Suppose
that $\{f_n\}$ is equicontinuous and $f_n(x) \to f(x)$ for each
$x \in X$ (point-wise convergence). Show that $f$ is continuous.







\end{description}
%\end{Large}
\end{document}














