\documentclass[10pt]{article}
\usepackage{amsxtra,latexsym,amssymb,amsthm}


\setlength{\parindent}{0cm}

\def\ds{\displaystyle}


\pagestyle{empty}

\textwidth=7.2truein
\textheight=9.5truein
\hoffset=-1.5truein
\voffset=-1.25truein

\begin{document}

\def\R{{\mathbb R}}

\begin{large}
\begin{center}
{\bf
  TOPOLOGY QUALIFYING EXAM}\\
  Fall 2001\\
  Muenzenberger and Strecker
\end{center}


\vspace{.1in}
Choose and work any 6 of the following 12 problems. Start each problem on a
new sheet of paper. \underbar{\bf Do not turn in more than six problems.}
A ``space" always means a topological space below.

{
\begin{list}{}
{\itemsep=0em \parsep=0em }\item[]
\end{list}
}

\begin{description}
%\setlength{\itemsep}{0em}
%\newlength{\parsep}{20em}

\item 1.
Assume that every open cover of the space $X$ has a countable subcover.
Let $A\subseteq X$ be an uncountable subset. Prove $A$ has a limit point.


\item 2.
Prove a space $X$ is compact if and only if every net in $X$ has a cluster
point.


\item 3.
A continuous map $f:(X,\tau)\to (Y,\sigma)$ is said to be \textbf{final}
provided that for each topological space $(Z,\mu)$ each set-function
$g:Y\to Z$ is continuous whenever $g\circ f:(X,\tau)\to(Z,\mu)$ is
continuous. Prove any two of the following statements:


\quad a)
The composition of final (continuous) maps is final.


\quad b)
The ``second factor" of a final map is final: i.e., if
$\ds{ (X,\tau) {\buildrel f\over\to} (Y,\sigma) {\buildrel h\over\to}
(W,s)}$
are continuous maps and $h\circ f$ is final, then $h$ is final.

\quad c)
Each quotient map is final.

\item 4.
Use Zorn's Lemma to prove that for each set $X$ and relation $R$ on $X$,
there is a maximal $A\subseteq X$ such that $A\times A\subseteq R$.


\item 5.
Prove that if $Y$ is compact, then the projection $\pi_1:X\times Y\to X$
is closed;
i.e., for each closed subset $A$ of $X\times Y$, $\pi_1[A]$ is closed in $X$.


\item 6.
Let $(X_1,\tau_1)$ and $(X_2,\tau_2)$ be topological spaces and let $\tau$
be the product topology on $X_1\times X_2$.
Let $Y_1\subseteq X_1$ and $Y_2\subseteq X_2$.
There is a ``natural" subspace topology $\sigma$ on $Y_1\times Y_2$ that
is induced from $\tau$ and the fact that $Y_1\times Y_2$ is a subset of
$X_1\times X_2$.
There is also a ``natural" product topology $\rho$ on $Y_1\times Y_2$
that comes from the subspace topologies $(Y_1,\rho_1)$
and $(Y_2,\rho_2)$ induced by $\tau_1$ and $\tau_2$ respectively.
Prove or disprove that $\sigma=\rho$.


\item 7.
Prove that no continuous function $f:S^1\to\R$ is $1-1$.
$$[S^1=\{(x,y)\in\R^2:x^2+y^2=1\}]$$
Note: $S^1$ and $\R$ have the usual topology.

\item 8.
Prove that a quotient of a locally connected space is locally connected.

\item 9.
Show that if $X$ is a compact metrizable space, then every metric
which generates the topology on $X$ is complete.

\item 10.
Prove that every paracompact Hausdorff space is regular.

\item 11.
Show that the Moore Plane (tangent disc space) is not normal.

\item 12.(a)
What is the fundamental group of $B^2=\{(x,y)\in\R^2: x^2+y^2\leq 1\}$?

\item \quad (b)
What is the fundamental group of $S^1=\{(x,y)\in \R^2:x^2+y^2=1\}$?

\item \quad (c)
Prove that there is no continuous function $r:B^2\to S^1$ such that
\newline $r(x,y)=(x,y)$ for all $(x,y)\in S^1$.

Note: $B^2$ and $S^1$ have the usual topology.

\end{description}
\end{large}
\end{document}
