% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}


\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Spring 1994}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

\begin{description}
\item[1.]
Prove or disprove: The continuous image of a locally connected space is
locally connected.

\item[2.]
Prove or disprove: A space $X$ is Hausdorff if and only if the diagonal
$\Delta: X \hookrightarrow X \times X$ is closed.

\item[3.] (a)
State the Axiom of Choice.

\item[\quad] (b)
Give another statement equivalent to the Axiom of Choice.

\item[4.]
Prove or disprove: A metric space is compact if and only if it is complete
and totally bounded.

\item[5.]
Give an example of a space which is connected but not path connected.

\item[6.]
Let $f: X \to Y$ be a continuous surjective map from a compact space $X$
to a Hausdorff space $Y$. Prove that $f$ is a quotient map.

\item[7.]
Prove or disprove: Every compact Hausdorff space is a Baire space.

\item[8.]
Prove or disprove: $S_\Omega$, the minimal uncountable well-ordered set
equipped with the order topology is not Lindel\"of.

\item[9.]
Prove or disprove: The product of a family of connected spaces is connected.

\item[10.]
Let $C(X,Y)$ be the set of continuous functions fro $X$ to $Y$, given the
compact-open topology. Let $e: C(X<Y) \times X \to Y$ be the evaluation
map $e(f,x) = f(x)$. Prove that if $X$ is locally compact Hausdorff, then
$e$ is continuous.

\item[11.] (a)
State Urysohn's Lemma

\item[\quad] (b)
Use Urysohn's Lemma to prove that for every finite open cover
${\mathcal U} = \{U_i\}^n_{i=1}$ of a normal space $X$, there is a partition
of unity on $X$ subordinated to $\mathcal U$.

\item[12] (a)
Prove that [0,1] in the usual topology is connected.

\item[\quad] (b)
Use your result in (a) to prove that every path-connected space is connected.

\item[13.]
Let $X$ be a completely regular space. Prove that $X$ is connected if and only
if its Stone-\u Cech compactification $\beta (X)$ is connected.





\end{description}    
%\end{Large}
\end{document}














