% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Fall 1990}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Work 6 of the following problems. Start each problem on a new sheet of
paper. Do not turn in more than 6 problems.

\begin{description}
\item[1.]
If $(X,\tau)$ and $(Y, \sigma)$ are topological spaces and $f: X \to Y$ is
a function, give a detailed set-theoretic proof that the following are
equivalent.

\item[\quad] (i)
For each $U \in \sigma, f^{-1} [U] \in \tau$.

\item[\quad] (ii)
For each $A \subseteq X, f[\overline A^\tau] \subseteq
  \overline{f[A]}^\sigma$.

\item[2.]
Prove that $[0,1]$, with its usual topology, is connected.

\item[3.]
Show that a space $X$ that has the fixed point properly is connected.

\item[4.] (a)
True-False.

\item[\qquad] (i)
An open and closed one-to-one function between topological spaces must be an
embedding.

\item[\qquad] (ii)
Each space that is locally-Hausdorff (in the sense that each point has
neighborhood base of Hausdorff subspaces) must be Hausdorff.

\item[\qquad] (iii)
Each quotient of a locally connected space must be locally connected.

\item[\qquad] (iv)
Each locally compact Hausdorff space is completely regular.

\item[\qquad] (v)
The product of metrizable spaces is metrizable.

\item[\quad] (b)
For each false entry, give a counterexample of other explanation (no proofs).

\item[5.] (a)
State the Axiom of Choice.

\item[\quad] (b)
Give another statement that is equivalent to the Axiom of Choice.

\item[6.]
Prove that if $A$ is a compact subset of a regular (not necessarily
Hausdorff) space $X$, then $\overline A$ is compact.

\item[7.]
Give an example of two topologies $\sigma$ and $\tau$ on the set of integers
$\Z$ for which $\sigma \subsetneq \tau$ and
$(\Z, \sigma)$ is homeomorphic to $(\Z, \tau)$.

\item[8.]
A continuous map $f: (X, \tau) \to (Y, \sigma)$ is said to be final provided
that for each topological space $(Z, \mu)$ each set-function $g: Y \to Z$
is continuous whenever $g \circ f: (X, \tau) \to (Z, \mu)$
is continuous. Prove that:

\item[\quad] (a)
The composition of final (continuous) maps is final.

\item[\quad] (b)
The ``second factor" of a final map is final, i.e., if
$(X, \tau) \stackrel{f}{\longrightarrow} (Y, \sigma)
  \stackrel{h}{\longrightarrow} (w,s)$
are continuous maps and $h \circ f$ is final, then $h$ is final.

\item[9.] (a)
Give an example of topological space $X$ that has both a Stone \u Cech
compactification, $\beta X$, and an Alexandorff compactification,
$\alpha X$, but for which $\alpha X$ and $\beta X$ are not homoemorphic.

\item[\quad] (b)
Give a reason why $\alpha X$ and $\beta X$, in part (a), are not homeomorphic.

\item[10.] (a)
If sequential limits in a space $X$ are unique, must $X$ be Hausdorff?

\item[\quad] (b)
Prove that your answer to (a) is correct.

\item[11.]
Show that the Moore Plane (tangent disc space) is not normal.

\item[12.]
Prove that if $A$ is a connected subset of connected space $X$ and if
$C$ is a component of $X-A$, then $X-C$ is connected.






\end{description}    
%\end{Large}
\end{document}














