\input amssym.def
\input amssym
\input epsf.tex
\input epsfig.sty
\documentstyle[11pt]{article}
\setlength{\textwidth}{16.5cm}
\setlength{\oddsidemargin}{0cm}
\setlength{\topmargin}{-1.4cm}
\setlength{\textheight}{23cm}


\begin{document}
\bibliographystyle{alpha}

\newtheorem{thm}{Theorem}
\newtheorem{defin}[thm]{Definition}
\newtheorem{lemma}[thm]{Lemma}
\newtheorem{propo}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{exa}[thm]{Example}

\centerline{\LARGE \bf Topology Qualifying Exam}

\centerline{\large \bf Spring 2002}

\centerline{\large (Yetter and Muenzenberger)}
\medskip

\noindent Choose and work any 6 of the following 10 problems.  Start each
problem on a new sheet of paper.  Include your name, the date and the question
number on each sheet. {\bf Do not turn in more than six (6) problems.}  
The word space below always means a topological space.
\medskip

\begin{enumerate}
\item Use Zorn's Lemma to prove that for each set $X$ and a (binary) relation
$R$ on $X$, there is a maximal $A \subset X$ such that $A \times A \subset X$.
\smallskip

\item Prove that if $A$ and $B$ are disjoint compact subsets of a Hausdorff
space $X$, then there are disjoint open sets $U$ and $V$ such that 
$A \subset U$ and $B \subset V$.
\smallskip

\item Prove or disprove that the continuous image of a compact space is 
compact.
\smallskip

\item Let $f:X\rightarrow Y$ be a continuous surjective map from a compact
space $X$ to a Hausdorff space $Y$.  Prove that $f$ is a quotient map.
\smallskip

\item Prove that if $Y$ is compact, then the projection $\pi_X:X\times Y
\rightarrow X$ is closed (i.e. for each closed subset $A$ of $X \times Y$,
the image $\pi_X(A)$ is closed in $X$).
\smallskip

\item Prove that a compact metric space has a countable basis.
\smallskip

\item Prove that no continuous function $f:S^1 \rightarrow {\Bbb R}$ is
one-to-one.  (Here $S^1$ is the unit circle $\{ (x,y) \in {\Bbb R}^2 |
x^2 + y^2 = 1 \}$, and both spaces have the usual topologies.)
\smallskip

\item Prove that a quotient of a locally connected space is locally
connected.
\smallskip

\item Let $X$ be a connected normal $T_2$ space having more than one
point.  Prove that $X$ is uncountable.
\smallskip

\item Describe the fundamental groups of the following spaces, where each
has the usual topology
\begin{enumerate}
\item the circle, $S^1$
\item the Moebius strip, M
\item the figure-eight (union of two circles in the plane intersecting in a
single point)
\item the torus $S^1 \times S^1$
\item the real projective plane, ${\Bbb R}{\Bbb P}^2$
\end{enumerate}
\end{enumerate}
\end{document}