% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Fall 1989}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Work 9 of the following problems. Do not turn in more than 9.

For a subset $A$ contained in a topological space $X$, let $A^-$ denote
the closure of $A$ in $X$, let $A^\circ$ denote the interior of $A$ in $X$,
and let $Fr(A)$ denote the frontier of $A$ in $X$. All Euclidean spaces will
have the usual topology, and all product spaces will have the product
topology.

\begin{description}
\item[1.]
For a topological space $X$, prove that ``o--" and ``--o" are both idempotent
operations on the subsets of $X$; that is, $A^{o-o-} = A^{o-}$ and
$A^{-o-o} = A^{-o}$ for all $A \subseteq X$.

\item[2.]
Let $X$ and $Y$ be topological spaces. Prove that a bijection $f: X \to Y$
is a homeomorphism if and only if $f[A^-] = f[A]^-$ for all $A \subseteq X$.

\item[3.]
Prove that the line $L$ defined by the equation $x=0$ separates the plane
$\R^2$ into exactly two components and is the frontier of each of them.

\item[4.]
Prove that $[0,1]$ (with its usual topology) is compact.

\item[5.]
Let $X$ be the unit square in the plane. Give precise descriptions of four
quotient maps with domain $X$ whose images are:

\item[\quad] a)
the sphere $S^2$;

\item[\quad] b)
the torus, $S^1 \times S^1$;

\item[\quad] c)
the cylinder, $S^1 \times I$;

\item[\quad] d)
a M\"obius band $M$ in $\R^3$.

\item[6.]
A continuous map $f: (X, \tau) \to (Y, \sigma)$ is said to be initial
provided that for each topological space $(Z, \mu)$, each set-function
$g:Z \to X$ is continuous whenever $f \circ g : (Z, \mu) \to (Y, \sigma)$
is continuous. Prove that:

\item[\quad] a)
The composition of initial maps is initial.

\item[\quad] b)
The ``first factor" of an initial map is initial; i.e., if
$(X, \tau) \stackrel{f}{\to} (Y, \sigma) \stackrel{h}{\to} (W, \nu)$
are continuous maps and $h \circ f$ is initial, then $f$ is initial.

\item[7.] (a)
True-False.

\item[\qquad] (i)
Every compact Hausdorff space is metrizable.

\item[\qquad] (ii)
If $f,g : X \to Y$ are homotopically equivalent and $h,k: Z \to W$ are
homotopically equivalent, then $f \times h$ and $g \times k$ are
homotopically equivalent.

\item[\qquad] (iii)
The product of connected spaces is connected.

\item[\qquad] (iv)
Every retract of a locally connected space is locally connected.

\item[\qquad] (v)
A space is $T_2$ if and only if it is locally $T_2$; i.e., each pt has a
base of $T_2$ neighborhoods.

\item[\qquad] (vi)
Every metrizable space is normal.

\item[\qquad] (vii)
Every locally compact Hausdorff space is completely reuglar.

\item[\qquad] (viii)
Every subspace of a separable Hausdorff space is separable and Hausdorff.

\item[\quad] (b)
For each false entry, give a counter example (no proofs).

\item[8.]
Does there exist a subset $C$ of the unit circle
$S^1 = \{(x,y) : x^2 + y^2 = 1\}$ such that

\item[\quad] a)
$C$ is closed in $S^1$

\item[\quad] b)
$C$ contains exactly one point from each pair of diametrically opposite
points on $S^1$.

\item[9.]
Compute the fundamental group of the compact surface $\hat T$ obtained by
removing an open disc from a torus $T$.

\item[10.]
Prove that every contractible space is pathwise connected.

\item[11.]
Prove that if $X$ is Lindel\"of and $Y$ is compact, then $X \times Y$ is
Lindel\"of.

\item[12.]
Suppose that $X$ and $Y$ are spaces for which the second projection
$\pi_2 : X \times Y \to Y$ is closed. Suppose also that $q: Y \to Y^\prime$.
is a quotient map. Prove or disprove that the second projection
$\pi^\prime_2 : X \times Y^\prime \to Y^\prime$ is closed.

\item[13.] a)
Give an example of an ultrafilter on the set $\R$ of real numbers.

\item[\quad] b)
Give an example of a filter on $\R$ that is not an ultrafilter.

\item[\quad] c)
Prove that every filter on $\R$ is contained in an ultrafilter
on $\R$.

\item[14.]
Prove that if a normal space $X$ contians a closed copy of $[0, \infty)$,
then $X$ does not have the fixed point properly.

\item[15.]
Suppose that there is an continuous open surjection $f: [0,1] \to X$
where $X$ is a nondegenerate IIausdorff space.

\item[\quad] a)
Show that $f$ need not be injective.

\item[\quad] b)
Show that $X$ is homeomorphic to $[0,1]$.

\item[16.]
Let $A$ be a nonempty proper subset of a continuum $X$. If $C$ is
component of $A$, show that $C \cap Fr(A) \neq 0$.





\end{description}    
%\end{Large}
\end{document}














