% latex file
%Topology QE S99 - Maginnis & Strecker

\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)

\input prepictex.tex
\input pictex.tex
\input postpictex.tex

%\documentstyle[bbb]{report}
\documentclass[bbb]{report}

\vcorrection{-1in}
\hcorrection{-1.15in}

\topmargin  0in
\textwidth 7in
\textheight 9.15in

\mathsurround=2pt

\def\ds{\displaystyle}
\def\R{{\Bbb R}}
\def\nd{\noindent}
\def\qed{\hfill\ \vbox{\hrule\hbox{\vrule\kern4pt\vbox{\kern4pt{}
\kern4pt}\kern4pt\vrule}\hrule}}


\begin{document}

%\pagestyle{empty}

\begin{Large}

\hfill Name \rule{2.15in}{.01in}
\par
\vspace{.1in}

\begin{center}
   TOPOLOGY QUALIFYING EXAM \\
   Spring 1999 \\
   (Maginnis and Strecker) \\
\end{center}

\end{Large}

\vspace{.1in}

\begin{large}
\noindent {\bf Choose and work any 6 of the following 15 problems.
Start each problem on a new sheet of paper.
\underbar{Do not turn in more than six problems}.
In the problems below, a space always means a topological space.}

\vspace{.2in}

\begin{description}

\item[1.]
Let $f:X\to Y$ be any function, let ${\cal A}\subseteq {\cal P}(X)$
and ${\cal B}\subseteq{\cal P}(Y)$. Prove or disprove each of the
following:

\item[\quad (a)]
$f[\bigcap {\cal A}]=\bigcap\{ f[A]|A\in{\cal A}\}$

\item[\quad (b)]
$f^{-1}[\bigcap {\cal B}]=\bigcap \{f^{-1}[B]|B\in{\cal B}\}$


\vspace{.15in}
\item[2.\ (a)]
State the Axiom of Choice.

\item[\quad (b)] State the Well-Ordering Theorem.

\item[\quad (c)] Either use the Axiom of Choice to prove the
Well-Ordering Theorem or use the Well-Ordering Theorem to prove the
Axiom of Choice.


\vspace{.15in}
\item[3.]
Let $X$ be a well-ordered set in the order topology, and assume
$X$ has a maximal element. Prove $X$ is a compact space.


\vspace{.15in}
\item[4.]
Prove that every closed subset of a metrizable space is a
countable intersection of open sets.


\vspace{.15in}
\item[5.]
Let $X$ be a connected and locally path connected space. Prove
$X$ is path connected.


\vspace{.15in}
\item[6.]
Let $I$ denote the closed unit interval $[0,1]$ in $\R$
(the real numbers with its usual topology). Prove or disprove
that in $\R^{\R}$ with its usual product topology
$(I^\circ)^{\R}=(I^{\R})^\circ$, (where $A^\circ$ denotes the
interior of $A$).



\vspace{.15in}
\item[7.]
Let $X\buildrel{f}\over{\longrightarrow} Y$
be any continuous function.

\vspace{.1in}
\item[\quad (a)]
Prove that there is a factorization of $f$
$X\buildrel{f}\over{\longrightarrow}Y=
 X\buildrel{q}\over{\longrightarrow}Z\buildrel{m}\over{\longrightarrow}Y$
 where $q$ is a quotient map and $m$ is a one-to-one continuous
 function.

\item[\quad (b)]
Prove that the factorization $f=m\circ q$ of part (a) is essentially
unique in the sense that if
$X\buildrel{f}\over{\longrightarrow}Y=
 X\buildrel{\widehat q}\over{\longrightarrow}\widehat Z
  \buildrel{\widehat m}\over{\longrightarrow}Y$
is also a factorization of $f$ with $\widehat q$ a
quotient map and $\widehat m$ one-to-one and continuous,
then there is a unique homeomorphism
$Z\buildrel{h}\over{\longrightarrow}\widehat Z$
such that $h\circ q=\widehat q$ and $\widehat m\circ h=m$.

\vspace{.15in}
\item[8.]
Prove that a filter $F$ on a space $X$ converges to a point
$x\in X $ if and only if the net $f$ based on $F$ converges to $x$.

\vspace{.15in}
\item[9.]
Let $(X,\leq)$ be a linearly ordered set and let
$\tau$ be the topology on $X$ inherited from the order.
If $A$ is a subset of $X$, there are two natural ways to
topologize $A$. The first, $\tau_1$, is as a subspace of
$(X,\tau)$.
The second, $\tau_2$, is as an ordered space with the order on $A$
inherited from $(X,\leq)$. Prove or disprove that in all cases
$(A,\tau_1)=(A,\tau_2)$.

\vspace{.15in}
\item[10.]
Let $A$ be a connected subset of a connected space $X$, and let
$C$ be a component of $X-A$. Prove $X-C$ is connected.


\vspace{.15in}
\item[11.]
Prove that a Hausdorff space with a basis consisting of sets that
are both open and closed is totally disconnected.


\vspace{.15in}
\item[12.]
Prove or disprove.

\item[\qquad (a)]
Every quotient map is an open map.

\item[\qquad (b)]
Every open map is a quotient map.


\vspace{.15in}
\item[13.]
Prove that a compact subset of a Hausdorff space is closed.

\vspace{.15in}
\item[14.]
Find a flaw in the following purported proof of the statement:

\vspace{.05in}
{\it Every discrete subspace of a topological space that has no isolated
points is nowhere dense.}

\vspace{.05in}
\nd Purported proof:

\vspace{.05in}
Suppose $X$ is a space with no isolated points and $D$ is a
discrete subspace of $X$. If $D$ is not nowhere dense, then
there is a nonempty open set $U$ with $U\subseteq\overline{D}$.
Thus there is some $d\in D\bigcap U$.
Since $D$ is discrete, there is some open set $W\subseteq U$
such that $W\bigcap D=\{d\}$.
Since $X$ has no isolated points $W\not=\{d\}$.
Thus $W \setminus \{d\}$ is a nonempty open set contained in
$\overline{D}$.
Thus there is some $\widehat{d}\in(W\setminus\{d\})\bigcap D
\subseteq W\bigcap D=\{d\}$,
which is a contradiction.
Thus $D$ must be nowhere dense.\qed



\vspace{.15in}
\item[15.]
Assume $X$ is a normal space. Let $\beta(X)$ be the
\v Cech-Stone compactification of $X$, and let $y\in\beta(X)-X$.
Prove that $y$ is not the limit of a sequence of points of $X$.




\vfill
\clearpage


\end{description}
\end{large}
\end{document}
