%From strecker%math.ksu.edu@KSUVM.KSU.EDU Wed Feb 22 08:32:25 1995
%Date: Tue, 21 Feb 95 15:29:51 CST
%From: George Strecker <strecker%math.ksu.edu@KSUVM.KSU.EDU>
%To: math@ksuvm

% LaTeX file


%\documentstyle[12pt]{article}
\documentclass[12pt]{article}
\topmargin -.5in
\textwidth 6.1in
\textheight 9.0 in
\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)


\renewcommand{\labelenumi}
{\mbox{ \mbox{{\rm {\roman{enumi}}}})}}
\renewcommand{\theenumi}{\roman{enumi}}
%This changes the numbering in itermizing to roman.


\begin{document}

 \begin{center}

 {\bf TOPOLOGY QUALIFYING EXAM \ \ Spring 1995}

	(Strecker and Wu)

 \end{center}


 Choose and work any {\bf 6} of the following 15 problems. Start each
  problem on a new sheet of paper. {\bf Do not turn in more than six
  problems.}
 A space always means a topological space below.

\vspace{10pt}
\noindent {\bf 1.}
Let $f, g:X\rightarrow Y$ be continuous. Assume that $Y$ is Hausdorff.
Show that the set$\{ x\in X\ | \ f(x)=g(x) \}$ is closed in $X$.

\vspace{10pt}

\noindent {\bf 2.}
Consider the function $f: X\rightarrow Y$ between topological spaces
and the following statements:

\noindent {\it Statement 1.}  $f$ is continuous iff for every subset $A$ of $X$
$f[cl_X[A]] \subseteq cl_Yf[A]$.

\noindent {\it Statement 2.}  $f$ is continuous iff for each open set $U$ in $Y$
the set $f^{-1}[U]$ is open in $X$.

\noindent {\it Statement 3.}  $f$ is continuous iff for each convergent net
$\alpha : \Lambda\rightarrow X$ the net $f\circ \alpha : \Lambda\rightarrow Y$
 is convergent.

\noindent {\it Statement 4.}  $f$ is continuous iff for each convergent
filter $\cal F$ on $X$ the filter on $Y$ generated by $\{f[F] | F \in \cal F\}$
is convergent.

\vspace{10pt}

Choose one of the above statements as the definition of continuity and use it to
show that one of the other statements is true.

\vspace{10pt}

\noindent {\bf 3.}
Prove that every retract of the real line with the usual topology is a closed
subset of the line.

\vspace{10pt}

\noindent {\bf 4.}
Show that the following two statements about a topological space $X$ are
equivalent:


(a) $X$  is connected.

(b)  No two-point discrete space is a quotient space of $X$.

\vspace{10pt}

\noindent {\bf 5.}
A space $X$ is called {\it locally metrizable} if each point $x$ of
$X$ has a neighborhood base consisting of metrizable subspaces.
Show that a compact Hausdorff space $X$ is metrizable if and only if
it is locally metrizable.

\vspace{10pt}

\noindent {\bf 6.}
Prove that the sequence $a: \Bbb N\rightarrow \Bbb R$ given by
$a_n =\frac{1}{n}$ is {\bf not} an ultranet in $\Bbb R$.

\vspace{10pt}

\noindent {\bf 7.}
Prove that for any function $f$ between spaces, the following are equivalent:

(a) $f$ is a homeomorphism.

(b) $f$ is a one-to-one quotient map.

(c) $f$ is a surjective embedding.

\vspace{10pt}

%\eject
\noindent {\bf 8.}
Prove or disprove the following statements:

(a) A connected open set in a locally path connected space is path
	connected.

(b) A path connected open set in a locally connected space is
	connected.

\vspace{10pt}

\eject

\noindent {\bf 9.}
Let $C_n$ be the subspace of ${\Bbb R}^2$ defined by
$$ C_n := \left\{ (x,y)\ | \
	\left(x-\frac{1}{n}\right)^2+ y^2 =\frac{1}{n^2} \right\}  $$
and let $Y=\cup_{n=1}^{\infty} C_n\subseteq {\Bbb R}^2$ with the
subspace topology. Let
$X=C_1\times \{ 1, 2, \ldots, \}\subseteq {\Bbb R}^2\times {\Bbb R}$
with the subspace topology. Define $g: X\rightarrow Y$ by
$$ g\left( (x, y), n\right)
	:= \left(\frac{x}{n}, \frac{y}{n}\right) .$$
Show that $g$ is continuous and surjective. Is $g$ a quotient map?


\vspace{10pt}
\noindent {\bf 10.}
Consider $X\times Y$ where $Y$ is a compact space.
Let $\{x_0\}\times Y\subseteq N $ where $N$ is open in $X\times Y$.
Show that there is a neighborhood $W$ of $x_0$, $W$ open in $X$ such
that
$$
\{x_0\}\times Y \subseteq W\times Y\subseteq N . $$


\vspace{10pt}
\noindent {\bf 11.}
(a) Give an example of an ultrafilter on the set $\Bbb N$ of natural numbers.

(b) Give an example of a filter on $\Bbb N$ that is not an ultrafilter.

(c) Prove that every filter on $\Bbb N$ in contained in an
ultrafilter on $\Bbb N$

\vspace{10pt}
\noindent {\bf 12.}
Prove or disprove the following statements:

(a) ${\Bbb R}$ (with the standard topology) is not a countable union
	of closed subsets each having empty interior.

(b)  ${\Bbb R}$ (with the standard topology) is not a countable union
	of subsets each having empty interior.


\vspace{10pt}
\noindent {\bf 13.}
Let $X$ be a topological space and $(Y, d)$ a metric space.
Let $C(X, Y)$ be the set of continuous maps from $X$ to $Y$. Show that
the compact-open topology  and the topology of uniform
convergence on compact subsets on $C(X, Y)$ coincide.

\vspace{10pt}
\noindent {\bf 14.}
Find an error in the following purported ``proof''
of the jactitation that if $K$ is compact, $\Bbb R$ has the usual topology,
 and $f: K\rightarrow \Bbb R$ is continuous, then $f$ is an open mapping.

\noindent
``\underline{\sc Proof}'':
Let $U$ be open in $K$. Then $E=K\setminus U$ is closed.  So $E$ is compact.
  Hence $f[E]$ is a compact subset of $\Bbb R$.  Thus $f[E]$ must be closed,
so that $\Bbb R \setminus f[E]=W$ is open.  But $W=f[U]$.  Hence $f$
preserves open sets and is thus an open mapping.

\vspace{10pt}
\noindent {\bf 15.}
Let $\Omega_0$ be the set of all ordinals less than the first uncountable
ordinal, with the order topology.  We know that $\Omega_0$ is $T_4$ and
that every real-valued function on $\Omega_0$ is constant on some tail.
Find the Stone-\v Cech compactification of $\Omega_0$. [Give your reasoning].


\end{document}






