% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}



\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Summer 1991}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Work 9 of the following problems. Start each problem on a new sheet of paper.
Do not turn in more than 9 problems.

\begin{description}
\item[1.]
Prove that a space $X$ is Hausdorff if and only if the diagonal
$\Delta = \{(x,x) |x \in X\}$ is closed in $X \times X$.

\item[2.] (a)
Show that $\overline{A \cup B} = \overline A \cup \overline B$.

\item[\quad] (b)
Construct a counterexample for an infinite union.

\item[3.]
Prove that if $A$ is a retract of a Hausdorff space $X$, then $A$ is closed
in $X$.

\item[4.]
Give an example of a quotient map which is neither open nor closed. Prove all
of your assertions.

\item[5.] (a)
Prove that every compact metric space is closed and bounded.

\item[\quad] (b)
Give an example that shows that not every closed, bounded metric space is
compact.

\item[6.]
Prove that $I \times I$ in the dictionary order topology is locally connected
but not locally path connected. Here $I$ denotes the unit interval
$[0,1]$.

\item[7.]
Let $X$ be a compact Hausdorff space. Prove that if every point of $X$ is a
limit point, then $X$ is uncountable.

\item[8.]
Let $G$ be a topological group such that all points are closed. Prove that
$G$ is a regualr space.

\item[9.]
Let $M$ be a compact, connected, orientable 3-dimensional manifold, and
assume that $H_1(M; \Z)$ is finite. Show that $H_2 (M;\Z) = 0$.

\item[10.]
Compute the homology with $\Z$-coefficients of the space
$\R P^2 \times \R P^3 \times \R P^4$.

\item[11.]
Prove that there does not exist a homeomorphism from $\C P^2$ to itself
which reverses the orientation.

\item[12.]
Let $A$ be a $ 3 \times 3$ matrix of positive real numbers. Prove that $A$
has a positive real eigenvalue.

\item[13.]
Assume that $A$ is a retract of $X$, and that $\pi_1(A)$ is a normal
subgroup of $\pi_1(X)$. Prove that
$\pi_1(X) \simeq \pi_1 (A) \times \frac{\pi_1(X)}{\pi_1(A)}$.

\item[14.]
Let $X=S^1 \vee S^2$ be the wedge of a circle and a sphere.

\item[\quad] (a)
Compute $\pi_1(X)$ and $\pi_2(X)$, giving explicit statements of any
theorems you use.

\item[\quad] (b)
Describe the action of $\pi_1(X)$ on $\pi_2(X)$.

\item[15.] (a)
State (do not prove) a theorem relating the coverings of a given space $X$
and the subgroups of $\pi_1(X)$. Be explicit with the hypotheses and
conclusions.

\item[\quad] (b)
Applying the theorem for $X = T^2$, describe the homeomorphism types of
spaces which cover the torus.

\item[16.]
Prove that if $X$ is a suspension of some space $Y$, then all cup products
in the cohomology ring of $X$ of elements of positive degree
are zero.






\end{description}    
%\end{Large}
\end{document}














