% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}


\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Fall 1991}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Work 6 of the following problems. Start each problem on a new sheet of paper.
Do not turn in more than 6 problems. Assume that all products have the
product topology.

\begin{description}
\item[1.]
Prove that $[0,1]$, with the usual topology, is connected.

\item[2.]
Prove that if $A$ is a retract of a Hausdorff space $X$, then $A$ is closed
in $X$.

\item[3.]
Prove that a quotient of a locally connected space is locally connected.

\item[4.]
Let $f : X \to Y$ be an open, continuous surjection. Prove that $Y$ is
Hausdorff if and only if the set
$$C= \{(x_1, x_2) \in X \times X |f(x_1) = f(x_2)\}$$
is closed subset of $X \times X$.

\item[5.]
Let $S^1$ have the usual topology. Prove that
$(\Bbb Q \times \Bbb Q) \cap S^1$ is dense in $S^1$.

\item[6.]
Let $\Bbb E$ denote the set of real numbers with the Sorgenfrey topology,
which has basis consisting of all half-open intervals of the form
$[x,y)$. Prove that any compact subset of $\Bbb E$ is countable.

\item[7.]
Prove that the first projection $\pi_1: X \times Y \to X$ is closed, if
$Y$ is compact.

\item[8.]
Let $\R^2$ have the usual topology. Prove that if $U$ is a convex open
subset of $\R^2$, then
$$\overline U^0 = U,$$
where ``-- " indicates closure and $``^0"$ indicates interior.

\item[9.]
Let $X$ be a compact Hausdorff space. Prove that if every point of $X$ is
a limit point of $X$, then $X$ is uncountable.

\item[10.]
State the Axiom of Choice and Zorn's Lemma, and prove that Zorn's Lemma
implies the Axiom of Choice.

\item[11.]
Let $X$ be a metric space. Show that if every family of pairwise disjoint
non-empty open subsets of $X$ is countable, then $X$ is separable.

\item[12.]
A space is called {\it functionally Hausdorff} if for every pair of distinct
points $x$ and $y$ in $X$, there exists a continuous function
$f : X \to [0,1]$ with $f(x) = 0$ and $f(y)$ = 1. Either prove or disprove
that every product of functionally Hausdorff spaces is functionally
Hausdorff.





\end{description}    
%\end{Large}
\end{document}














