% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\P{\mathbb P}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Spring 1992}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\
Work 6 of the following problems. Start each problem on a new sheet of
paper. Do not turn in more than six problems.

\begin{description}
\item[1.]
For a topological space $(X, \tau)$ define
$I: {\mathcal P} (X) \to {\mathcal P} (X)$ by $I(A) = A^\circ$ and
$C: {\mathcal P} (X) \to {\mathcal P} (X)$ by
$C(A) = \overline A$. For a set $A$ of $X$, consider the sequence $A$,
$I(A), CI(A), ICI(A), CICI(A), \dots$

\item[\quad] a)
For any space what is the largest number of distinct sets that this sequence
can contain?

\item[\quad] b)
Find $A \subseteq \R$ for which this largest number is obtained (where $\R$
has the usual topology).

\item[2.]
Prove that the projection of  $[0,1] \times \R$ onto $\R$ is a closed map,
where both have their usual topologies.

\item[3.]
Prove that if $A \subseteq X \times X$, then there is a maximal subset $Y$
of $X$ with $Y \times Y \subseteq A$.

\item[4.]
For any topological space $X$, define an equivalence relation $\sim$ on
$X$ by $x \sim y$ iff $\overline{\{x\}} = \overline{\{y\}}$. Prove that
the resulting quotient space $X/ \sim$ is $T_0$.

\item[5.]
Let $\Bbb E$ denote the set of real numbers with the Sorgenfrey topology,
which has basis consisting of all half-open intervals of the form
$[x,y)$. Prove that $\Bbb E \times \Bbb E$ is not normal.

\item[6.]
Let $f_i : [-1,1] \to \R$ be the function defined by
$f_i(x) = (i + x)^2, i = -1, 0, 1$. Prove or disprove that the (unique)
evaluation map $f: [-1,1] \to R^3$ with the property that
$\pi_i \circ f = f_i$ is an embedding.

\item[7.]
Prove that if $\rho$ is a metric on $M$, then $\rho_1 : M \times M \to R$
defined by $\rho_1(x,y) = \frac{\rho(x,y)}{1+ \rho(x,y)}$
is a metric on $M$ which is equivalent to $\rho$.

\item[8.]
Prove that if $\{A_\lambda | \lambda \in \Lambda\}$ is locally finite
system of sets, then
$\overline{\cup_{\lambda \in \Lambda} A_\lambda} = \cup_{\lambda \in \Lambda}
  \overline{A_\lambda}$.

\item[9.]
Let $f$ be a one-to-one onto function from the real line with the usual
topology itself. If $f$ and $f^{-1}$ map connected subsets onto connected
subsets, prove that $f$ is a homeomorphism.

\item[10.]
Let $\Omega_0$ be the set of all ordinals less than the first uncountable
ordinal with the order topology. We know that $\Omega_0$ is $T_4$ and every
continuous real-valued function on $\Omega_0$ is constant on some tail.
Find the Stone- \u Cech compactification of $\Omega_0$. (Justify your
answer!)

\item[11.]
Let $\P$ be the irrational numbers with the usual (subspace) topology. Show
that the intersection of any countable family of dense open subsets of
$\P$ is dense in $\P$.

\item[12.]
Let $X$ be a locally compact Hausdorff space and let $C(X,Y)$ be the space
of all continuous functions from $X$ into $Y$ with the compact-open
topology. Prove that the map $P:C(X,Y) \times X \to Y$ defined by
$P(f,x) = f(x)$ is continuous.





\end{description}    
%\end{Large}
\end{document}














