% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Spring 1996}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\
Choose and work any 6 of the following 14 problems. Start each problem
on a new sheet of paper. {\bf Do not turn in more than six problems}. A
space always means a topological space below.

\begin{description}
\item[1.] (a)
Prove or disprove that if $m$ and $n$ are any cardinal numbers then
$m+n= n+m$.

\item[\quad] (b)
Prove or disprove that if $m$ and $n$ are any ordinal numbers then
$m+n = n+m$.

\item[2.]
Prove or disprove that every filter on a nonempty set is contained in an
ultra-filter on that set.

\item[3.]
Prove that $\Bbb R$ and $\Bbb R^2$ with their usual topologies are not homeomorphic.

Use the following definition for problems 4 and 5.

For any topological property $P$ a space is $X$ called {\bf locally} $P$
if and only if every point $x$ of $X$ has a base of neighborhoods each of
which has property $P$ (as a subspace).

\item[4.] (a)
Prove or disprove that locally compact spaces are hereditary [i.e., that
every subspace of a locally compact space is locally compact].

\item[\quad] (b)
Prove or disprove that compact spaces are hereditary.

\item[5.]
Prove that exactly one of the properties locally $T_1$ and locally $T_2$ is
hereditary; i.e, that one is hereditary and the other is not.

\item[6.]
Let $X$ be a $T_1$ completely regular space. Prove that $X$ is connected if
and only if its \v Cech-Stone compactification $\beta(X)$ is connected.

\item[7.]
List three different bases for the usual topology of the plane (i.e., $\Bbb R^2$).

\item[8.]
A topological property $P$ is called {\it divisible} iff every topological
quotient of a space with property $P$ has property $P$.

\item[\quad] (a)
Which of the following properties are divisible and which are not? Hausdorff,
connected, locally connected, compact, metrizable, discrete.

\item[\quad] (b)
Prove one of your assertions from part (a).

\item[9.]
A topological property $P$ is called productive iff every topological
product of a family of spaces with property $P$ has propery $P$.

\item[\quad] (a)
Which of the following properties are productive and which are not? Hausdorff,
connected, locally connected, compact, metrizable, discrete.

\item[\quad] (b)
Prove one of your assertions from part (a).

\item[10.]
Find an incorrect statement in the proof of the following theorem and prove
that it is an incorrect statement.

{\bf Theorem.} If $B^2 = \{(x_1, x_2) \in \Bbb R^2 |x^2_1 + x^2_2 \leq 1\}$
has the usual topology, then each continuous function
$f: B^2 \to B^2$ has a fixed point.

{\bf Proof:}
Suppose that $f:B^2 \to B^2$ is a continuous function with no fixed points.
Let $\pi_1 : B^2 \times B^2 \to B^2$ be first projection, let
$$\Delta = \{(z,z)|z \in B^2\}$$
be the diagonal in $B^2 \times B^2$, and let
$$F= \{(z,f(z)|z \in B^2\}$$
be the graph of $f$. Since $f$ is continuous, $\pi_1 |F:F \to B^2$ is a
homeomorphism. Since $B^2$ is connected, $F$ is therefore connected. However,
$$F \subseteq B^2 \times B^2 - \Delta$$
which is a contradiction since $B^2 \times B^2 - \Delta$ is connected.

\item[11.] (a)
Prove or disprove that the subspace topology on the set of integers, $\Bbb Z$,
in $\Bbb R$ (with its usual topology) is the sames as the discrete topology on
$\Bbb Z$.

\item[\quad] (b)
Prove or disprove that the subspace topology on the set of rationals, $\Bbb Q$,
in $\Bbb R$ (with its usual topology) is the same as the discrete topology on
$\Bbb Q$.

\item[12.]
Let $(X,d)$ be a metric space. Show that the following are equivalent:

\item[\quad] (i)
$X$ has a countable basis;

\item[\quad] (ii)
$X$ is Lindelof;

\item[\quad] (iii)
$X$ has a countable dense subset.

\item[13.]
Let $X$ be a regular space and assume that $U$ is a countable open covering
of $X$ with the property that the closure of each of its members is
paracompact. Prove that $X$ is paracompact.

\item[14.]
Prove that the topologist's comb
$$C=\left( \bigcup^\infty_{n=1} \left\{ \frac{1}{n}\right\}
  \times [0,1] \right) \cup \left([0,1] \times \{0\} \right) \cup
  \left(\{0\} \times [0,1] \right)$$
is not a retract of the square $S = [0,1] \times [0,1]$ (with both $C$ and
$S$ having the usual Euclidean subspace topologies).
  




    
\end{description}
%\end{Large}
\end{document}














