% latex file
%Topology QE S00 - Lee & Strecker

\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }
\def\subplus{
{\hspace{-.25in}} \hbox{\ }_{~\over{+}}\ } %puts a plus symbol below the line

\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)

\input prepictex.tex
\input pictex.tex
\input postpictex.tex

%\documentstyle[bbb]{report}
\documentclass[bbb]{report}

\vcorrection{-.7in}
\hcorrection{-1in}

\topmargin  0in
\textwidth 6.4in
\textheight 9.5in

\mathsurround=2pt

\def\ds{\displaystyle}
\def\R{{\Bbb R}}
\def\N{{\Bbb N}}

\begin{document}

%\pagestyle{empty}

\begin{Large}

\hfill Name \rule{2.5in}{.01in}
\par
\vspace{.1in}

\begin{center}
   TOPOLOGY QUALIFYING EXAM \\
  Spring 2000 \\
   (Lee and Strecker) \\
\end{center}


\vspace{.1in}

\noindent {\bf Choose and work any 6 of the following problems.
Start each problem on a new sheet of paper.
{\bf Do not turn in more than six problems}.
In the problems below, a ``space" always means a ``topological space".}

\vspace{.2in}

\begin{description}

\item[1.]
Prove or disprove each of the following:

\item[\quad] a)
If $A$ is a connected subset of space $X$ and $Q\subseteq X$
such that $A\subseteq Q\subseteq \overline{A}$.
then $Q$ must be connected.

\item[\quad] b)
Each component of a space must be closed.

\item[\quad] c)
Each component of a space must be open.


%\vspace{.15in}

\item[2.]
Prove or disprove each of the following:

\item[\quad] a)
Every subspace of a compact Hausdorff space is locally compact.

\item[\quad] b)
Every subspace of a connected Hausdorff space is locally connected.

%\vspace{.15in}


\item[3.]
Prove that each compact subset of a Hausdorff space must be closed but a
compact subset of a $T_1\hbox{-space}$ need not be closed.

%\vspace{.15in}

\item[4. a)]
Complete the following sentence so that it is a true statement:

\item[\qquad] Given a set $X$, a family ${\cal B}$ of subsets of $X$
is a base for a topology
on $X$ if, and only if \hrulefill.

\item[\quad] b)
Construct another true sentence that describes a subbase for a topology
on $X$.

%\vspace{.15in}
\item[5.]
Let $(X_1,\tau_1)$ and $(X_2,\tau_2)$ be topological spaces and let
$\tau$ be the product topology on $X_1\times X_2$.
Let $Y_1\subseteq X_1$ and $Y_2\subseteq X_2$.
There is a ``natural" subspace topology $\sigma$ on $Y_1\times Y_2$
that is induced from $\tau$ and the fact that $Y_1\times Y_2$
is a subset of $X_1\times X_2$.
There is also a ``natural" product topology $\rho$ on $Y_1\times Y_2$
that comes from the subspace topologies $(Y_1,\rho_1)$ and $(Y_2,\rho_2)$
induced by $\tau_1$ and $\tau_2$ respectively.
Prove or disprove that $\sigma=\rho$.


%\vspace{.15in}
\item[6.]
Let
$A=\{x|x\hbox{\ is an ordinal number that is less than or equal to }\omega_1\}$,
where $\omega_1$ is the first uncountable ordinal number
and let $\tau$ be the order-topology on $A$ induced by the usual order.
Prove one of the following:


\item[\quad] a)
The space $(A,\tau)$ is a compact Hausdorff space.

\item[\quad] b)
If $f:(A,\tau)\to \R$ is a real-valued continuous function
(where $\R$ is assumed to have its usual topology)
then there exists some
$x_0<\omega_1$ such that $f(x)=f(x_0)$ for all $x\geq x_0$ in $A$.

\item[\quad] c)
The Stone-\v Cech compactification of $A\backslash \{\omega_1\}$
(with its usual subspace topology) is
equivalent to $(A,\tau)$.


%\vspace{.15in}
\item[7.]
Prove that for any set $X$ the cardinality of $X$ is strictly smaller
than the cardinality of ${\cal P}(X)$, the set of all subsets of $X$.

%\vspace{.15in}

%\hfill\eject

\item[8.]
A continuous map $f:(X,\tau)\to (Y,\sigma)$ is said to be {\bf final}
provided that for each topological space $(Z,\mu)$ each set-function
$g:Y\to Z$ is continuous whenever $g\circ f:(X,\tau)\to (Z,\mu)$
is continuous. Prove any two of the following statements:

\item[\quad] a)
The composition of final (continuous) maps is final.

\item[\quad] b)
The ``second factor" of a final map is final: i.e., if
$$(X,\tau)\buildrel{f}\over{\to} (Y,\sigma)\buildrel{h}\over{\to}(W,s)$$
are continuous maps and $h\circ f$ is final, then $h$ is final.

\item[\quad] c)
Each quotient map is final.


%\vspace{.15in}
\item[9.]
Work one of the following:

\item[\quad] a)
Give an example of two topologies $\sigma$ and $\tau$ on the
set of integers $\Bbb Z$ for which $\sigma
\subset_\subplus\tau$ and
$({\Bbb Z},\sigma)$ is homeomorphic to $({\Bbb Z},\tau)$.

\item[\quad] b)
Give an example of a point-finite open covering of the real line
with the usual topology that is not locally finite.


\item[10.]
Prove one of the following statements:

\item[\quad] a)
Each compact Hausdorff space is a Baire space.

\item[\quad] b)
The set of rational numbers is not a $G_\delta\hbox{-set}$ in $\R$
(where $\R$ has its usual topology).

\vfill\eject

\item[11.]
Prove or disprove one of the following:

\item[\quad] a)
For each open covering $\{U_\alpha|\alpha\in A\}$ of a paracompact space $X$,
there is a partition of unity subordinated to $\{U_\alpha\}_{\alpha\in A}$.

\item[\quad] b)
Each locally compact Hausdorff space is paracompact.

%\vspace{.15in}
\item[12.]
Prove one of the following theorems:

\item[\quad] a)
Urysohn Lemma.

\item[\quad] b)
Tychonoff Theorem on Product of Compact Spaces.

\item[\quad] c)
Urysohn Metrization Theorem.

%\vspace{.15in}
\item[13.]
Let $X$, $Y$ and $Z$ be topological spaces and $X$ be locally compact
Hausdorff. Let $C(X,Y)$ be the space of continuous functions from
$X$ into $Y$ with compact-open topology. For a function
$$ F:X\times Z\to Y$$
let $\hat F$ be the function from $Z$ into the set of all functions
from $X$ to $Y$ defined by
$$ \hat F (z)(x) = F(x,z). $$
Prove one of the following:


\item[\quad] a)
If $F$ is continuous, then  $\hat F$ is continuous from $Z$ into
$C(X,Y)$.

\item[\quad] b)
If $\hat F$ is continuous from $Z$ into $C(X,Y)$ then $F$ is continuous.


\vspace{.15in}
\item[14.]
For any $\varepsilon>0$ and any
$(a,b)\in {\Bbb R}^2$ let an
\underbar{$\varepsilon \hbox{-plus centered at\ } (a,b)$}
be the set
  $+_\varepsilon(a,b)=\{(x,b)\in\R^2\bigg| |x-a| <\varepsilon\}
   \cup \{(a,y)\in{\Bbb R}^2\bigg| |y-b|<\varepsilon\}$.
Then call a set $U\subseteq{\Bbb R}^2$ \underbar{plus-open}
if for each $(a,b)\in U$ there is some $\varepsilon>0$ such that
$+_\varepsilon(a,b)\subseteq U$.

\item[\quad] a)
Prove that the plus-open subsets of ${\Bbb R}^2$ form a topology for
${\Bbb R}^2$, (called the plus-topology for ${\Bbb R}^2$).

\item[\quad] b)
Prove that the plus-topology for ${\Bbb R}^2$ is \underbar{strictly}
finer than the usual topology for ${\Bbb R}^2$.



\end{description}
\end{Large}
\end{document}
