% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Fall 1988}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Work 9 of the following problems. Do not turn in more than 9.

\begin{description}
\item[1.]
Let $A$ be a subset of space $X$ and let $\overline A$, $Fr(A)$, and
$A^\circ$ denote the closure, frontier, and interior of $A$ in $X$,
respectively. The set $A$ is called poor if and only if $Fr(A) = A$, and
the set $A$ is called thin if and only if $\overline A^\circ = \emptyset$.
Prove the following

\item[\quad] (i)
If $A$ is poor, then $A$ is thin.

\item[\quad] (ii)
If $A$ is closed, then $A$ is poor if and only if $A$ is thin.

Give an example of space $X$ in which some subset $A$ is thin but not poor.

\item[2.] (a)
True-False

\item[\qquad] (i)
The composition of quotient maps is a quotient map.

\item[\qquad] (ii)
The product of metrizable spaces is metrizable.

\item[\qquad] (iii)
$f: X\to Y$ is a topological embedding if and only if $f$ is one-to-one and
$X$ has the coarset (=weakest) topology making $f$ continuous.

\item[\qquad] (iv)
A space is $T_1$ if and only if it is locally $T_1$; i.e., each pt has a base
of $T_1$ neighborhoods.

\item[\qquad] (v)
A space is $T_2$ if and only if it is locally $T_2$; i.e., each pt has a base
of $T_2$ neighborhoods.

\item[\qquad] (vi)
Every metrizable space is normal.

\item[\qquad] (vii)
Every locally compact Hausdorff space is completely regular.

\item[\qquad] (viii)
Every subspace of a separable Hausdorff space is separable and Hausdorff.

\item[\quad] (b)
For each false entry, give a counter example (no proofs).

\item[3.]
A family of functions $(f_i : A \to B_i)_{i \in I}$ with a common domian
is called a mono-source provided that for each set $C$ and each pair of
functions $h,k: C \to A$, $f_i \circ h = f_i \circ k$
for each $i \in I$ implies that $h=k$.

Prove that $(f_i: A \to B)_{i \in I}$ is a mono-source if and only if
it separates points; i.e., for each $a, a^\prime \in A$ with
$a \neq a^\prime$ there is some $j \in I$ such that
$f_j(a) \neq f_j(a^\prime)$.

\item[4.]
A continuous map $f: (X, \tau) \to (Y, \sigma)$ is said to be initial
provided that for each topological space $(Z, \mu)$, each set-function
$g: Z \to X$ is continuous whenever $f \circ g : (Z, \mu) \to (Y, \sigma)$
is continuous.

Prove that:

\item[\quad] (a)
The composition of initial maps is initial.

\item[\quad] (b)
the product of initial maps is initial; i.e., each of
$(X_i, \tau_i) \stackrel{f_i}{\to} (Y_i, \sigma_i)$
is initial, then so is the unique induced map
$$(\Pi X_i, \tau) \stackrel{f= \langle f_i \rangle}{\longrightarrow}
  (\Pi Y_i, \sigma)$$
where $\tau$ and $\sigma$ are the product topologies.

\item[5.]
Prove that each non-degenerate completely regular $T_1$ connected space is
uncountable.

\item[6.]
Prove that any open subset of a locally connected, second countable space
$X$ can be written as the union of a finite or infinite sequence of
disjoint open, connected subsets of $X$.

\item[7.]
Consider the ordinal spaces $[1, \omega]$ and $[1, \omega_1]$ where $\omega$
is the first inifinite ordinal and $\omega_1$ is the first uncountable
ordinal. Recall that every continuous real-valued function on
$[1, \omega_1)$ (or on $[1, \omega_1]$) is constant on some tail. Prove that
the Stone-\u Cech compactification of the deleted Tychonoff plank
$[1, \omega_1] \times [1, \omega] - \{(\omega_1, \omega)\}$ is the
Tychonoff plank $[1, \omega_1] \times [1, \omega]$.

\item[8.]
Give 4 examples, 1 compact, 1 non-compact, 1 non-locally connected, 1
non-locally compact, of spaces homotopically equivalent to $S^1$, but not
homeomorphic to $S^1$.

\item[9.]
Characterize the sequences $(x_n, y_n)$ in the Moore plane $\Gamma$
(tangent disc space) which converge to $(0,0)$.

\item[10.]
Show that $\prod_{\alpha \in A} X_\alpha$ is connected if and only if each
$X_\alpha$ is connected.

\item[11.]
Let $f: X \to Y$ be perfect. Show that if $Y$ is paracompact, so is $X$.

\item[12.]
Find a flaw in the ``Proof" of the following:

{\bf Theorem.} All real-valued functions are continuous.

{\bf Lemma.} $f: X \to R$ is continuous if and only if $g_f$ defined by
$x \mapsto (x, f(x))$ is continuous.

{\bf Proof.}
If $f$ is continuous, then $g_f$ is determined by continuous functions, $f$
and identify, so it must be continuous. If $g_f$ is continuous then $f$
is $g_f$ followed by the second projection, so it must be continuous.

{\bf Proof of Theorem}
For any $f: X \to R$, $g_f$ is one-to-one. Thus it is continuous if and only
if its inverse is open. However the inverse of $g_f$ is the first projection.
But all projection functions are well-known to be open surjections. Hence
$g^{-1}_f$ is open, so $g_f$ is continuous. Hence, by the lemma, $f$ is
continuous.

\item[13.]
Prove that there is no smallest base for the usual topology on $R$.

\item[14.]
Prove that the following are equivalent for any topological space $Y$:

\item[\quad] (a)
$Y$ is Hausdorff.

\item[\quad] (b)
the diagonal $\Delta_Y = \{(y,y) | y \in Y\}$ is a closed subset of
$Y \times Y$.

\item[\quad] (c)
for each space $Z$ and each pair of continuous functions $f,g: Z \to Y$ that
agree on a dense subset of $Z$, it follows that $f=g$.

\item[15.] (a)
Prove that if $A \times B$ is compact subset of $X \times Y$ contained in an
open set $W$ in $X \times Y$, then there exist open sets $U \subseteq X$ and
$V \subseteq Y$ such that
$$A \times B \subseteq U \times V \subseteq W.$$

\item[\quad] (b)
Prove that the product of two compact spaces is compact.






\end{description}    
%\end{Large}
\end{document}














