
\documentclass[12pt]{article}
\usepackage{amsxtra,amssymb,amsthm,amsmath,latexsym}

\setlength{\parindent}{0cm}
\def\ds{\displaystyle}

\def\R{{\mathbb R}}
\def\N{{\mathbb N}}
\def\Z{{\mathbb Z}}
\def\C{{\mathbb C}}
\def\ds{\displaystyle}
\def\nd{\noindent}
\def\iff{ {\hbox{\,iff\,}} }
\def\ra{\rightarrow}
\def\lra{\longrightarrow}

\pagestyle{empty}

%\textwidth=7.25truein\hoffset=-1truein
\textwidth=6.5truein\leftmargin=0in\hoffset=-.5truein
%\textheight=9.5truein\voffset=-1truein
\textheight=9truein\voffset=-.5truein

\begin{document}

\begin{large}

\begin{center}
{\bf
Topology Qualifying Exam\\
Miller \& Maginnis, Fall 2002
}\end{center}

\vspace{.15in}


\begin{description}

\item[1.]
Given subsets $A$, $V$ and $W$ of a topological space $X$ such that
$A$ is closed, $V$ is open, $W$ is open and $A\cap W\subset V$.
Prove that $(A\cap W) \cup(V\backslash A)$ is open.

\vspace{.15in}
\item[2.]
Let $X$ be a topological space such that every sequence of points of
$X$ has a convergent subsequence. Show that every countable open
cover of $X$ has a finite subcover.

\vspace{.15in}
\item[3.]
Prove that a countable product of separable  spaces is separable
(recall that a separable space is one which has a countable dense subset).

\item[4.]
Suppose that $X$ is a regular topological space and that $A\subset X$ is a
closed subset. Show that the quotient space $X/A$ is  Hausdorff.

\vspace{.15in}
\item[5.]
Show that every locally compact Hausdorff space is regular.

\vspace{.15in}
\item[6.]
Let $f:S^2\to S^2$ by $f(x,y,z)=(-x,-y,-z)$.
Define an equivalence relation on
$S^2$ by $vRw \iff v=w$ or $w=f(v)$ for $v,w\in S^2$.
Let $X=S^2/R$ with the quotient topology and $p:S^2\to X$
be the quotient mapping.

\item[\quad ]a)
Show that $p$ is a covering map.

\item[\quad] b)
Use covering space theory to determine the fundamental group of $X$.

\vspace{.15in}
\item[7.]
Let $X$ be a compact, Hausdorff, locally connected space, $Y$ be a
Hausdorff space and $f:X\to Y$ a continuous onto mapping.
Prove that $Y$ is locally connected.

\end{description}
\end{large}
\end{document}
