% latex file
%Muenzenberger & Strecker

\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{{\mathbb R}}
\def\Q{{\mathbb Q}}
\def\F{{\mathbb F}}
\def\K{{\mathbb K}}
\def\C{{\mathbb C}}
\def\Z{{\mathbb Z}}
\def\E{{\mathbb E}}
\def\N{{\mathbb N}}
\def\divs{\vert}
\def\notdivs{\not\vert}
\def\rt{\,\root\of}
\def\inv{^{-1}}
%out of date:
%\def\K{{\Bbb K}}
%\def\C{{\Bbb C}}
\def\vs{\vspace{0.2in}}
\def\vvs{\vspace{0.5in}}

\begin{document}
\begin{large}


\begin{center}\begin{LARGE}
  {\bf  Topology Qualifying Exam}\\ 
  {\bf August 2000}\\ \end{LARGE}
  {\bf (Muenzenberger and Strecker)}
\end{center}

\vspace{0.1in}
\noindent\hrulefill\\
{\bf Instructions:
Choose and work any 6 of the following 14 problems. Start each problem on
a new sheet of paper. \underline {Do not turn in more
that six problems.}
In the problems below, a space means a topological space.
}

\vspace{0.2in}



\begin{description}

\item[1.]
Prove or disprove that  the continuous image of a compact space is compact.

\item[2.(a)]
True - False.

\item[\quad \quad (i)]
An open and closed one-to-one function between topological spaces must be
an embedding.

\item[\quad \quad (ii)]
Each space that is locally-Hausdorff (in the sense that each point has
a neighborhood base of Hausdorff subspaces) must be Hausdorff.

\item[\quad \quad (iii)]
Each quotient of a locally connected space must be locally connected.

\item[\quad \quad (iv)]
Each locally compact Hausdorff space is completely regular.

\item[\quad \quad (v)]
The product of metrizable spaces is metrizable.

\item[\quad \quad (vi)]
The product of continua is a continuum. (A continuum is a compact, connected,
Hausdorff space.)

\item[\quad \quad (vii)]
Every metrizable space is normal.

\item[\quad \quad (viii)]
Every subspace of a separable space is separable.

\item[\quad (b)]
For each false entry, give a counterexample or other explanation (no proofs).

\item[3.]
Let $(X_i, \tau_i)_{i \in I}$ be a family of 
topological spaces and let $\tau$ be the
product topology on $\Pi X_i$. Let $Y_i \subseteq X_i$ for each $i \in I$.
There is a ``natural" subspace topology $\sigma$ on $\Pi Y_i$ that is
induced from $\tau$ and the fact that $\Pi Y_i$ is a subset of
$\Pi X_i$. There is also a ``natural" product topology $\rho$ on
$\Pi Y_i$ that comes from the subspace topologies $(Y_i, \rho_i), i \in I$
induced by the topologies $(\tau_i, i \in I)$. Prove or disprove that
$\sigma = \rho$.

\item[4.]
Let $f : X \to Y$ be a continuous surjective map from a compact space $X$
to a Hausdorff space $Y$. Prove that $f$ is a quotient map.

\item[5.]
Use Zorn's lemma to prove that for every set $X$ and relation $R$ on $X$
there is a maximal $A \subseteq X$ such that $A \times A \subseteq R$.

\item[6.]
Prove that the interval $[0,1]$ with the usual topology is connected.

\vfill\pagebreak

\item[7.]
A continuous map $f:(X, \tau) \to (Y, \sigma)$ is said to be final provided that
for each topological space $(Z, \mu)$ each set-function $g:Y \to Z$ is
continuous whenever $g \circ f:(X,\tau) \to (Z, \mu)$ is continuous.
Prove any two of the following:

\item[\quad (a)]
The composition of final (continuous) maps is final.

\item[\quad (b)]
The ``second factor" of a final map is final; i.e., if
$(X, \tau) \stackrel{f}{\longrightarrow} (Y, \sigma)
\stackrel{h}{\longrightarrow}
(W,\rho)$
are continuous maps and $h \circ f$ is final, then $h$ is final.

\item[\quad (c)]
Every quotient map is final.

\item[8.]
Prove or disprove each of the following:

\item[\quad a)]
Every subspace of a compact Hausdorff space is locally compact.

\item[\quad b)]
Every subspace of a connected Hausdorff space is locally connected.

\item[9.]
Describe the fundamental groups of the following spaces, where each has the
usual topology.

\item[\quad (a)]
The circle, $S^1$;

\item[\quad (b)]
The Mobius Strip, $M$;

\item[\quad (c)]
The figure eight, $\infty$;

\item[\quad (d)]
The torus, $S^1 \times S^1$;

\item[\quad (e)]
The projective plane, ${\mathbb P}$.

\item[10.]
Prove that if a filter $F$ is contained in a unique ultrafilter $G$, then
$F=G$.

\item[11.]
Prove that every continuous bijection $f: {\R} \to {\R}$ is a
homeomorphism, where $\R$ has the usual topology.

\item[12.]
Prove that if $A$ and $B$ are compact subsets of a Hausdorff space $X$, then
there are open disjoint sets $U$ and $V$ such that $A \subseteq U$ and
$B \subseteq V$.

\item[13.]
Prove that if $X$ and $Y$ are spaces where $Y$ is compact and
$\{x\} \times Y \subseteq W$, an open set in $X \times Y$, then there is
$U$ open in $X$ and $V$ open in $Y$ such that
$\{x\} \times Y \subseteq U \times V \subseteq W$.

\item[14.]
Prove that if a normal space $X$ contains a closed homemorphic image of
$\R$ (with the usual topology), then there is a continuous function
$f :X \to X$ such that $f(x) \neq x$ for all $x \in X$.





    
\end{description}
\end{large} \end{document}
















