%ÿWPCX
%ûÿ2BVPZ¦Courier 10cpiðÿÿ‰?xxx,Ûwôxþ6X@É“8Ç;X@þþþþþþþÿþÿÿÿþÿÿþÿÿÿÿÿÿÿÿÿÿÿÿÿÿÿÿÿÿÿÿHP LaserJet IIIDHPLASII0.PRSÛxŒ
%@É‡Ï,\,ð°6mX@ÐÐÓÓûÿ2FV#|x%QX©F95
%Crane & Wu
% LaTeX file 


%\documentstyle[12pt]{article}
\documentclass[12pt]{article}
\topmargin -.5in
\textwidth 6.1in
\textheight 9.0 in
\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)


\renewcommand{\labelenumi}
{\mbox{ \mbox{{\rm {\roman{enumi}}}})}}
\renewcommand{\theenumi}{\roman{enumi}} 
%This changes the numbering in itermizing to roman.
\newcounter{problem}\setcounter{problem}{0}
\renewcommand{\theproblem}{\arabic{problem}}
\newcommand{\problem}{%\setcounter{subsubsection}{0}
     \par \vspace{1em} 
      \stepcounter{problem} 
          \noindent {\bf {\theproblem}. }
     }


\begin{document}

 \begin{center}

 {\bf TOPOLOGY QUALIFYING EXAM \ \ Fall 1995}

     (Crane and Wu)

 \end{center}

Work out problem {\bf 1} and then choose {\bf 4} (and only 4) 
additional problems among the remaining ones. Start each 
problem on a new sheet of paper. 
{\bf Do not turn in more than 4 additional problems.}
A space always means a topological space below.

\vspace{20pt}
{\bf This one is required!}
\problem 
Answer the following questions and give a brief explanation or
counterexample:
\begin{enumerate}
\item What is the smallest possible cardinality of a noncompact
Hausdorff
     space?
\item \begin{enumerate}
     \item Is the disjoint union of finitely many compact spaces
compact?
     \item Is the Cartesian product of finitely many compact
spaces
          compact? 
     \item Is the Cartesian product of any family of compact
spaces
          compact?
     \item Is the intersection of any family of compact subsets
of a
          Hausdorff space compact?
     \end{enumerate}
\item Is the intesection of two connected spaces connected?
\item Is the Cartesian product of finitely many connected spaces
connected?
\item Is the Cartesian product of uncountably many connected
spaces connected?
\item Is every space metrizable?
\item Is the inverse of any continuous bijective function
continuous?
\item Is the composite of two continuous functions continuous?
\item What is a quotient map?
\item Is a quotient map always an open map?
\end{enumerate}

\vspace{12pt} {\bf Choose 4 problems from below:}

\problem
Let 
$S:=\left\{ (x, y)\in {\Bbb R}^2 \ | 
    \ y=\sin{\frac{1}{x}}, x > 0 \right\}\subset {\Bbb R}^2 $. 
   The closure $\overline{S}$ of $S$ (in ${\Bbb R}^2$) 
   is called the {\bf topologist's sine curve}. 
It is endowed with the subspace topology. Show that the
topologist's sine
   curve is connected, but not path connected. 
   Is $\overline{S}$ locally connected?

\problem
Let $f, g:X\rightarrow Y$ be continuous. Assume that $Y$ is
Hausdorff.
Show that the set $\{ x\in X\ | \ f(x)=g(x) \}$ is closed in $X$.

\problem
Let $(X, d)$ be a metric space. Show that the following are
equivalent:
\begin{enumerate}
 \item $X$ has a countable basis;
 \item $X$ is Lindel\"{o}f;
 \item $X$ has a countable dense subset.
\end{enumerate}

\problem
Prove or disprove the following statements:
\begin{enumerate}
\item The continuous image of a locally connected space is
locally connected;
\item The quotient of a locally connected space is locally
connected.
\end{enumerate}

\problem
Let $f: X\rightarrow Y$ be a map, where $Y$ is a compact
Hausdorff
   space and $X$ any topological space. Show that $f$ is
continuous
   iff the graph of $f$,
   $$ G_f :=\{ (x, f(x))\ |\ x\in X\}\subset X\times Y$$
   is closed in $X\times Y$. 

\problem
Show that if $Y$ is compact,then the projection map
$$ \pi_1: X\times Y\longrightarrow X $$
is a closed map. Is the compactness of $Y$ necessary?

\problem
Let $X$ be a regular space with a countable basis and $U$ a
   non-empty open subset. Show that
\begin{enumerate}
  \item $U$ is a countable union of closed subsets of $X$;
  \item There is a continuous function $f: X\rightarrow [0, 1]$
        such that $f(x)>0$ for $x\in U$ and $f(x)=0$ for
$x\not\in U$.
\end{enumerate}

\problem
A space $X$ is called {\it locally metrizable} if each point $x$
of $X$
has a neighborhood that is metrizable in the subspace topology.
Show that a regular Lindel\"{o}f space is metrizable if it is
locally
metrizable.

\problem
Let $f:X\rightarrow {\Bbb R}$ be a continuous real-valued
function on
the compact Hausdorff space $X$. Show that every nonempty open
subset of $X$ contains a nonempty open set on which $f$ is
bounded.

\problem
Prove or disprove the following statements:

(a) ${\Bbb R}^2$ (with the standard topology) is a countable
union
     of closed subsets each having empty interior.

(b)  ${\Bbb R}^2$ (with the standard topology) is a countable
union
     of subsets each having empty interior.

\problem
Let $Y$ be a metric space and $f_n:X\rightarrow Y$ a sequence of
continuous 
functions and $f: X\rightarrow Y$ a function (not necessariy
continuous).
Suppose $\{ f_n\}$ is equicontinuous and $f_n(x)\rightarrow f(x)$
for each
$x\in X$. Show that $f$ is continuous.


\end{document}

  





