% latex file
%Maginnis

\documentclass[12pt]{article}
\usepackage{amsxtra,amssymb,amsthm,amsmath,latexsym}

\topmargin  0in
\textwidth 7.0truein
\textheight 9.5truein

\voffset=-1truein
\hoffset=-1.25truein
\mathsurround=2pt

\def\R{{\mathbb R}}
\def\N{{\mathbb N}}
\def\Z{{\mathbb Z}}
\def\C{{\mathbb C}}
\def\diam{\hbox{\,diam\,}}
\def\lub{\hbox{\,lub\,}}
\def\ds{\displaystyle}

\begin{document}

\begin{large}

\begin{center}
   TOPOLOGY QUALIFYING EXAM\\
   Spring 2001 \\
   (Maginnis and Muenzenberger)\\
\end{center}

\end{large}

\vspace{.2in}

Choose and work any 6 of the following 10 problems.
Start each problem on a new sheet of paper.
 \underbar{\bf Do not turn in more than six problems.}
A space always means a topological space below.

\begin{description}

\item[] 1.
Assume that every open cover of the space $X$  has a countable subcover.
Let $A\subseteq X$ be an uncountable subset. Prove $A$ has a limit point.

\vspace{.1in}

\item[] 2.
Let $f:X\to Y$ be an open, continuous, surjective map. Let
$R\subseteq X\times X$ be the set $R=\{(x_1,x_2)\in X\times X|
f(x_1)=f(x_2)\}$.
Assume $R$ is closed in the product topology on $X\times X$.
Prove $Y$ is a Hausdorff space.

\vspace{.1in}

\item[] 3(a)
Prove an open subspace of a separable space is separable.

\item[\quad] (b)
Give an example of a subspace of a separable space which is not separable.

\vspace{.1in}

\item[] 4.
Let $X$ be a complete metric space, and let
$\{ C_n\subseteq X| n\in \N \}$
be a nested sequence of nonempty, closed, bounded subsets of $X$, with
$C_1\supseteq C_2\supseteq C_3\supseteq \dots$.
The diameter of a bounded subset of a  metric space is defined as the least
upper bound of the distances between points,
$\diam(C)=\lub\{ d(x,y) | x,y \in C \}$.
Assume $\ds\lim_{n\to\infty}\diam(C_n)=0$.
Prove the intersection $\ds\bigcap^\infty_{n=1}C_n$ is
\underbar{not} empty.

\vspace{.1in}

\item[] 5.
Let $X$ be a locally compact Hausdorff space.
Prove a subset $U\subseteq X$ is open if and only if for every compact
subspace $C\subseteq X$, $U\cap C$ is an open subset of $C$.

\vspace{.1in}

\item[] 6.
Let $X$ be a connected normal $T_2$ space having more than one point.
Prove that $X$ is uncountable.

\vspace{.1in}

\item[] 7.
Let $X$ be a connected space, and assume $X-\{x_0\}$ is not
connected for some point $x_0\in X$. Let $X-\{x_0\}=U\cup V$
be a separation, so that $U$ and $V$ are disjoint, nonempty, open
(and closed) subsets of $X-\{x_0\}$. Prove $A=U\cup \{x_0\}$
is a connected subspace of $X$.

\vspace{.1in}

\item[] 8.
Prove a compact metric space has a countable basis.

\vspace{.1in}

\item[] 9.
Prove a space $X$ is compact if and only if every net in $X$
has a cluster point.


\vspace{.1in}

\item[] 10.
Let $X$ be a paracompact space, and let $Y$ be a compact Hausdorff space.
Use the Tube Lemma to prove that $X\times Y$, in the product topology,
is paracompact.




\vfill

\end{description}
\end{document}



