% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Topology Qualifying Exam}\\ 
{\bf Spring 1988}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Do 9 of the following 17 problems.

\begin{description}
\item[1.]
Prove that the following statements are equivalent for any topological space
$X$:

\item[\quad] a)
$X$ is Hausdorff

\item[\quad] b)
for any space $Y$ and for each pair of continuous functions
$f, g: Y \to X$ that agree on a dense subset of $Y, f=g$.

\item[2.]
Prove that the product of normal $T_1$ spaces must be $T_1$ and completely
regular, but need not be normal.

\item[3.]
Prove or give a counterexample for each of the following  implications:

separable $\stackrel{(a)}{\Longrightarrow}$ second countable
$\stackrel{(b)}{\Longrightarrow}$ first countable $\stackrel{(c)}{\Longrightarrow}$
separable $\stackrel{(d)}{\Longrightarrow}$ first countable
$\stackrel{(e)}{\Longrightarrow}$ second countable $\stackrel{(f)}{\Longrightarrow}$
separable.

\item[4.]
Let $A$ be a subset of a topological space $X$. Show that the following
statements are equivalent:

\item[\quad] a)
$x \in \overline A$

\item[\quad] b)
there is a filter $\mathcal F$ on $X$ such that $A \in \mathcal F$ and
$\mathcal F$ converges to $x$.

\item[\quad] c)
There is a net $\alpha : \Lambda \to A$ that converges to $x$.

\item[5.]
Let $(X_\alpha) _{\alpha \in A}$ be a collection of topological spaces and
let $x$ be fixed point in $\prod_{\alpha \in A} X_\alpha$. Show that
$$D= \left\{ y \in \prod_{\alpha \in A} X) \alpha |\  y \hbox{\ and\ } x
\hbox{\ differ\ } \hbox{in\ } \hbox{at\ } \hbox{most\ } \hbox{finitely\ }
\hbox{many\ } \hbox{coordinates\ } \right\}$$
is dense in $\prod_{\alpha \in A} X_\alpha$.

\item[6.]
Let $\{A_\alpha\}$ be a point-finite covering of a Hausdorff space $X$. Show
that $\{A_\alpha\}$ has a subcovering $\{B_\gamma\}$ so that no subset of
$\{B_\gamma\}$ covers $X$.

\item[7.]
Prove that the following statements are equivalent for any topological space
$X$:

\item[\quad] a)
$X$ is Tychonoff (i.e., $T_1$ and completely regular)

\item[\quad] b)
$X$ can be embedded in a compact Hausdorff space.

\item[8.]
Show that for $n \geq 1$ the $n$-sphere, $S^n$, is homeomorphic to the
one-point compactification of Euclidean $n$-space, $E^n$.

\item[9.]
Find an error in the following purported ``proof" that $2^R$ is metrizable,
where $2=\{0,1\}$ is a two point discrete space, and $R$ is the set of
real numbers.

{\bf Proof}. Consider the inclusion $N \subseteq R$, where $N$ is the natural
numbers. This induces a ``natural" embedding $2^N \hookrightarrow 2^R$
where $f \mapsto \hat f$ and
$$\hat f (r) = \begin{cases}
                f(r) &\hbox{\ if\ } r \in N \\
                0 &\hbox{\ if\ } r \in R-N.
                \end{cases}.$$
But $2^N$ is the Cantor space $C$. Thus $C$ is embedded in $2^R$. Let $U$ be
an open subset of $2^R$. By the definition of product topology, the projection
of $U$ is $\{0,1\} = 2$ in all but finitely many coordinates. Thus
$U \cap 2^N \neq \emptyset$. So $2^N$ is dense in $2^R$. But $2^N$ is compact
and $2^R$ is Hausdorff, so $2^N$ is closed in $2^R$. Thus $2^N = 2^R$. But
$C$ is metrizable, and so $2^R$ is metrizable.

\item[10.]
Prove that every quotient of a locally connected space is locally connected.

\item[11.]
Let $E^n$ denote Euclidean $n$-dimensional space. For what values of $n$ is
it true that $E^n - \{p\}$ is simply connected for any point $p \in E^n$.
Prove your answer.

\item[12.]
Let $E^n$ denote Euclidean $n$-dimensional space. For what values of $n$ is
it true that $E^n-C$ is connected for any countable subset $C$. Prove your
answer.

\item[13.]
Show that in a locally compact Hausdorff space, countable intersections of
dense open sets are dense and open.

\item[14.]
A covering $\{U_\alpha\}_{\alpha \in A}$ of a topological space $X$ is called
shrinkable if there is a covering $\{V_\alpha\}_{\alpha \in A}$ of $X$
such that $\overline{V_\alpha} \subseteq U_\alpha$ for each
$\alpha \in A$ and $V_\alpha \neq \emptyset$ whenever
$U_\alpha \neq \emptyset$. Show that a Hausdorff space is normal if and only
if each point-finite covering is shrinkable.

\item[15.]
Prove that every separable metrizable space can be embedded in a countable
product of closed unit intervals.

\item[16.]

\item[\quad] a)
Give an example of an ultrafilter on the set $\R$ of real numbers.

\item[\quad] b)
Give an example of a filter on $\R$ that is not an ultrafilter.

\item[\quad] c)
Prove that every filter on $\R$ is contained in an ultrafilter on $\R$.

\item[17.]
Show that every compact connected Hausdorff space is the continuous image of
$[0,1]$.



\end{description}    
%\end{Large}
\end{document}














