%From fangbing@math.ksu.edu Mon Feb 23 11:30:22 1998
%Subject: Top Qual for 98 Spring

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%-*-latex-*-            This is using LaTeX mode for emacs%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NeedsTeXFormat{LaTeX2e} %LaTeX2.09 can't be used for this document 
\documentclass[12pt]{article}   %% class {article}
\usepackage{mathrsfs}                   %% use script fonts
\usepackage{amsthm,amsfonts             %% AMS fonts etc.
%%      ,amsmath                        %% use AMS-extra symboles
%%      ,amscd                          %% use commutative diagrams
         }

\topmargin -.5in
\textwidth 6.1in
\textheight 9.0 in

\renewcommand{\labelenumi}
{\mbox{ \mbox{{\rm {\roman{enumi}}}})}}
\renewcommand{\theenumi}{\roman{enumi}} 
%This changes the numbering in itermizing to roman.

\newcounter{problem}\setcounter{problem}{0}
\renewcommand{\theproblem}{\arabic{problem}}
\newcommand{\problem}{%\setcounter{subsubsection}{0}
        \par \vspace{1.2em} 
         \stepcounter{problem} 
                \noindent {\bf {\theproblem}.  \ }
        }

\begin{document}

 \begin{center}

 {\bf TOPOLOGY QUALIFYING EXAM \ \ Spring 1998}

	(Strecker and Wu)

 \end{center}


 Choose and work any {\bf 6} of the following problems. Start each 
  problem on a new sheet of paper. {\bf Do not turn in more than six
  problems.}
Below a ``space'' always means a ``topological space''. 

\problem
Let $(X, \leq)$ be a linearly ordered set and let $\tau$ be the order
topology on $X$ determined by $\leq$. Let $Y$ be a subset of $X$. Then there
is a ``natural'' topology $\sigma$ on $Y$ that is induced by the inherited
linear ordering on $Y$ and another ``natural'' topology $\rho$ on $Y$ that
is the subspace topology induced by $\tau$. Then prove or disprove that
$\sigma = \rho$.

\problem Prove or disprove each of the following:
\begin{enumerate}
\item
The intersection of any family of topologies on a set
$X$ is a topology on $X$.

\item
The union of two topologies on a set $X$ is a topology on $X$.
\end{enumerate}

\problem
Prove that every second countable space is Lindel\"{o}f. (If you used the
Axiom of Choice, show precisely where it is used. If you do not use the
Axiom of Choice, state so.)


\problem Prove or disprove each of the following:
\begin{enumerate}
\item $\mathbb{R}$ in the standard topology cannot be written as a
countable union of open sets;
\item $\mathbb{R}$ in the standard topology cannot be written as a
countable union of closed sets;
\item $\mathbb{R}$ in the standard topology cannot be written as a
countable union of sets with empty interiors;
\item $\mathbb{R}$ in the standard topology cannot be written as a
countable union of closed sets with empty interiors.
\end{enumerate}

\problem
Prove or disprove each of the following:
\begin{enumerate}
\item If $A$ is a connected subset of a space $X$ and $Q\subseteq X$ such
that $A\subseteq Q\subseteq \overline{A}$, then $Q$ must be connected.
\item Each component of a space must be closed.
\item Each component of a space must be open.
\end{enumerate}

\problem
Let $\rho_1, \rho_2$ be two metrics on the set $X$ such that both metric
spaces $(X, \rho_1)$ and $(X, \rho_2)$ are complete. If $X$ is totally
bounded with respect to the metric $\rho_1$, is it also totally bounded with
respect to the metric $\rho_2$? Prove you assertion.

\eject
\problem
\begin{enumerate}
\item State the Tychonoff Product Theorem (TPT)
\item If AXC is the Axiom of Choice, show either 
	$AXC\Rightarrow TPT$ or $TPT \Rightarrow AXC$.
\end{enumerate}

\problem
Prove or disprove that the product of two normal spaces is normal.

\problem
Let $X$ be a normal $T_1$ space and $W=\cup_{n\in \mathbb{Z} } C_n\subseteq X$,
where each $C_n$ is closed and $W$ is open in $X$. Show that there is a
continuous function $f: X \rightarrow [0, 1]$ such that $f(x)> 0$ for all
$x\in W$ and $f(x)=0$ for all $x\not\in W$.

% \problem
% Let $\mathbb{R}$ be the set of real numbers with the standard topology.
% Let $\mathbb{R}^J$ be endowed with the product topology where $J$ is an
% index set. When is $\mathbb{R}^J$ metrizable and when it is not metrizable?
% Prove your statement.

\problem 
Let $(X_1, \tau_1)$ and $(X_2, \tau_2)$ be topological spaces and let $\tau$
be the product topology on $X_1\times X_2$. Let $Y_1\subseteq X_1$ and
$Y_2\subseteq X_2$. There is a ``natural'' subspace topology $\sigma$ on
$Y_1\times Y_2$ that is induced from $\tau$ and the fact that 
$Y_1\times Y_2$ is a subset of $X_1\times X_2$. There is also a ``natural''
product topology $\rho$ on $Y_1\times Y_2$ that comes from the subspace
topologies $(Y_1, \rho_1)$ and $(Y_2, \rho_2)$ induced by $\tau_1$ and
$\tau_2$ respectively. Then prove or disprove that $\sigma = \rho$.

\problem
Let $X$ be a Hausdorff space. Suppose there exists a countable open covering
$\{ U_n\}_{n=1}^{\infty}$ of $X$ such that for each $n$, $\overline{U_n}$ is
compact and $\overline{U_n}\subseteq U_{n+1}$. Show that $X$ is paracompact.

\problem
Prove or disprove that the contimuous image of a compact space is compact.


\end{document}

  






