% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}



\begin{center}\begin{LARGE}
{\bf Algebra Qualifying Exam}\\ 
{\bf Fall 1992}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

All rings are assumed to have a multiplicative identity, denoted 1. The
fields $\Q$, $\R$ and $\C$ are the fields of {\it rational, real} and
{\it complex} numbers, respectively.

\begin{description}

\item[1.]
Let $G$ be a group of order $2n$, where $n$ is odd. Prove that $G$ has
a normal subgroup of index 2.

\item[2.]
Let $G$ be a group of order $5 \cdot 7 \cdot 11$. Prove that
$7 \parallel Z(G)|$, where $Z(G)$ denotes the {\it center} of $G$.

\item[3.]
Let $F$ be a field and let $R$ be the ring
$$R= \left\{ \begin{bmatrix} a&0 \\ 0&b \end{bmatrix} | a \in F \right\}.$$

Define the $R$-modules
$$M_1 = \left\{\begin{bmatrix} a \\ 0 \end{bmatrix} | a \in F \right\},
  M_2 = \left\{\begin{bmatrix} 0 \\ b \end{bmatrix} | b \in F \right\}.$$
Prove that $M_1 \ncong M_2$.

\item[4.]
Let $R$ be a ring and let $M$ be an irreducible left $R$-module. Prove
that there exists a maximal left ideal $I$ such that
$R/I \cong M$ as left $R$-modules.

\item[5.]
Let $R$ be a ring. An ideal $P \subseteq R$ is called {\it primary} if for
any $a,b \in R$ with $ab \in P$, $a \notin P$, then $b^n \in P$ for some
positive integer $n$. Show that $P$ is primary if and only if every
zero-divisor of $R/P$ is nilpotent.

\item[6.]
Let $V$ be an $n$-dimensional vector space over a field $F$, and let
$$V= V_0 \supseteq V_1 \supseteq \dots \supseteq V_n = 0$$
be a chain of subspaces of $V$, with $\dim(V_i/ V_{i+1}) = 1$ for
$i=0,1, \dots, n-1$. Suppose that $T:V \to V$ is linear transformation
satisfying $T(V_i) \subseteq V_{i+1}$ for all $i = 0,1, \dots, n-1$.
Compute the characteristic polynomial of $T$.

\item[7.]
Let $V$ be a complex vector space and let $T \in End_C$.
If $T^{1992} = 1_V$, prove that $T$ is diagonalizable.

\item[8.]
Let $f(x), g(x) \in F[x]$, and assume that $F \subseteq K$ is an extension of
fields. If $f(x)$ divides $g(x)$ in $K[x]$, prove that $f(x)$ divides
$g(x)$ in $F[x]$.

\item[9.]
Let $K \subseteq C$ be the splitting field over $Q$ for the polynomial
$x^6-1 \in Q[x]$. Compute the Galois group of $K$ over $Q$ and show
exactly how it operates on $K$.





\end{description}    
%\end{Large}
\end{document}














