% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}


\begin{center}\begin{LARGE}
{\bf Algebra Qualifying Exam}\\ 
{\bf Fall 1991}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

All rings are assumed to have a multiplicative identity, denoted 1. The
fields $\Q$, $\R$ and $\C$ are the fields of {\it rational, real} and
{\it complex} numbers, respectively.


\begin{description}

\item[1.]
Let $G$ be a group and let $N_1, N_2$ be normal subgroups of $G$ with
$N_1 \cong N_2$. Prove, or give a counterexample to the assertion
$$G/N_1 \cong G/N_2.$$

\item[2.]
Let $G$ be a finite group and let $\cal C$ be a conjugacy class of elements
in $G$. If $|{\cal C}| = \frac{1}{2} |G|$, show that every element of
$\cal C$ is an involution (i.e., an element of order 2).

\item[3.]
Let $R$ be a commutative ring, and let $x \in R$. Define what it means
for $x$ to be an {\it irreducible} element, and define what it means for
$x$ to be {\it prime}. If $R$ is a {\it unique factorization domain},
show that $x$ is irreducible if and only if $x$ is prime.

\item[4.]
Let $R$ be a principal ideal domain, and let $0 \neq I \subset R$ be a prime
ideal. Prove that $I$ is a maximal ideal.

\item[5.]
Let $R$ be a ring and let $M$ be an irreducible left $R$-module. Prove
that there exists a maximal left ideal $I$ such that $R/I \cong M$
as left $R$-modules.

\item[6.]
Consider the map $T: \C \to \C$ defined by setting $T(\alpha) = (2+i) \alpha$.
If we regard $T$ as an $\R$-linear transformation of the 2-dimensional
$\R$-vector space $\C$, compute $\det(T)$.

\item[7.]
Let $V$ be an $n$-dimensional vector space over a field $F$, and let
${\cal B} = \{x_1, x_2, \dots, x_n\}$ be a basis for $V$. Let $V^\ast$
denote the dual space of $V$, that is, $V^\ast$ is the vector space
$Hom_F(V,F)$ of all linear transformations $\lambda : V \to F$.
Define elements $\lambda_1, \dots, \lambda_n$ of $V^\ast$ by setting
$$\lambda_i \left( \sum^n_{j=1} a_jx_j \right) = a_i,$$
$1 \leq i \leq n, a_j \in F$, and put
${\cal B}^\ast = \{\lambda_1, \dots, \lambda_n\}$. Show that ${\cal B}^\ast$
is a basis of $V^\ast$.

\item[8.]
Consider the matrix of rational entries
$$A= \begin{bmatrix} 0&-1 \\ 1&-1 \end{bmatrix}.$$

If $R=\Q[A]$ is the ring of polynomials in $A$ with rational coefficients,
prove that $R$ is a field. (Hint: Consider the homomorphism $\Q[x] \to \Q[A]$,
where $x \mapsto A$.)

\item[9.]
Let $f(x) = x^5 - 2 \in \Q[x]$.

\item[\quad] (a)
Construct a splitting field $K \supseteq \Q$ for $f(x)$ over $\Q$.

\item[\quad] (b)
Find an element $\alpha \in K, \alpha \notin \Q$, such that
$\Q[\alpha]$ is a normal extension of $\Q$.





\end{description}    
%\end{Large}
\end{document}














