% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}



\begin{center}\begin{LARGE}
{\bf Algebra Qualifying Exam}\\ 
{\bf Spring 1990}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\
All rings are assumed to have a multiplicative identity, denoted 1. The
fields $\Q$, $\R$ and $\C$ are the fields of {\it rational, real} and
{\it complex} numbers, respectively.

\begin{description}
\item[1.]
Let $G$ be a group, and let $H$, $K$ be subgroups of $G$. If
$K \triangleleft G$, prove that $HK/K \cong H/(H \cap K)$.

\item[2.]
Let $G$ be a finite group and let $\cal C$ be a conjugacy class of elements
in $G$. If $|{\cal C}| = \frac{1}{2} |G|$, show that every element of
$\cal C$ is an involution (i.e., an element of order 2).

\item[3.]
Let $R$ be a commutative ring, and let $x \in R$. Define what it means for $x$
to be an {\it irreducible} element, and define what it means for $x$
to be {\it prime}. If $R$ is a {\it unique factorization domain}, show that
$x$ is irreducible if and only if $x$ is prime.

\item[4.]
State and prove the {\it Eisenstein criterion} for irreducibility of
polynomials.

\item[5.]
Let $K \subseteq \C$ be the splitting field over $\Q$ for the polynomial
$x^6-1 \in \Q[x]$. Compute the Galois group of $K$ over $\Q$ and show exactly
how it operates on $K$.

\item[6.]
Let $f(x), g(x) \in F[x]$, and assume that $F \subseteq K$ is an extension of
fields. If $f(x)$ divided $g(x)$ in $K[x]$, prove that $f(x)$ divides $g(x)$
in $F[x]$.

\item[7.]
Consider the map $T: \C \to \C$ defined by setting $T(\alpha) = (2 + i) \alpha$.
If we regard $T$ as an $\R$-linear transformation of the 2-dimensional
$\R$-vector space $\C$, compute $\det(T)$.

\item[8.]
Let
$$A= \begin{bmatrix} 2&3&4 \\ 0&2&3 \\ 0&0&1 \end{bmatrix}$$

be a real matrix. Find $3 \times 3$ real matrices $N$, $D$ with
$A=N+D, DN = ND$, where $N$ is nilpotent and $D$ is diagonalizable.

\item[9.]
A band of 17 pirates decided to divide their gold coins into equal portions.
When they found that they had 3 coins remaining, they agreed to give them to
their Chinese cook Wun Tu. But 6 of the pirates were killed in a fight. Now
when the treasure was divided equally among them, there were 4 coins left
that they considered giving to Wun Tu. Before they could divide the coins,
there was a shipwreck and only 6 pirates, the coins, and the cook were saved.
This time equal divison left a remainder of 5 coins for the cook. Now Wun
Tu took advantage of his culinary position to concoct a poison mushroom
stew so that the entire fortune in gold coins became his own .What is the
smallest number of coins that the cook would have finally received?






\end{description}    
%\end{Large}
\end{document}














