% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf  Algebra Qualifying Exam}\\ 
{\bf August 22, 2000}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\
{\bf Instructions:} You are given 10 problems from which you are to do 8.
 Please indicate those  8 problems which you would like  to be graded 
by circling the problem numbers on the  problem sheet. 

\noindent{\bf Note:} All rings on this exam are associative and
have multiplicative identity 1. All  integral domains are 
assumed to be commutative.
\vspace{0.2in}

\vs


\begin{enumerate}

\item Prove that a finite group of order $224=2^5\cdot 7$ cannot be a simple group.

\item Let $G$ be a finite group, and let $\phi:G\to G$ be a homomorphism
satisfying $\phi^2=1_G$. Assume also that $\phi(g)=g$ if and only if
$g=e$, the identity element of $G$. 
\begin{enumerate}
\item Show that for all $g\in G$, there exists $h\in G$ with $g=h^{-1}\cdot f(h)$.
\item Prove that $G$ must be abelian.
\end{enumerate}
 
\item Let $G$ be a group, and let $H$ be a cyclic normal subgroup of $G$. Prove
that the commutator subgroup $G'$ is a subgroup of the centralizer 
$C_G(H)=\{g\in G\,|\, gh=hg, \mbox{ for all }h\in H\}.$ (Hint:  note that
${\rm Aut}(H)$ is an abelian group.)

\item Let $R$ be a ring and assume that $R=
 \bigoplus _{i=1}^nI_i $, where $I_1,\ldots I_n$ are minimal left ideals
of $R$.  If $M$ is an irreducible left $R$-module, prove that $M\cong
I_j$ for some index $j,\ 1\leq j\leq n$.

\item Let $R$ be a unique factorization domain and assume that for every
pair of elements $a,b\in R$, their greatest common divisor $d$ can be
expressed in the form $d=ra+sb$, for suitable $r,s\in R$. Prove that
$R$ must be a principal ideal domain.  

\item  Let $R$ be a simple ring and let $C=Z(R)$,
the center of $R$.  Prove that $C$ is a field. (Hint: if $0\not=a\in C$,
consider the mapping $R\to R$, given by $r\mapsto ar$. Why must this
mapping be bijective?)


\item Let $A$ be a nilpotent $n\times n$ matrix, and assume that its
Jordan canonical form consists of a single Jordan block. Find the
Jordan canonical form of $A^2$. (Hint: there are two cases, depending
on the parity of $n$.)

\item Let $V$ be an $\F$-vector space, where $\F$ is an algebraically closed
field. Assume that $T:V\to V$ is a linear transformation such that for any
$T$-invariant subspace $W\subseteq V$, there is a $T$-invariant subspace
$W'\subseteq V$ such that $V=W\oplus W'$. Prove that $T$ must be diagonalizable.

 \item Let $\F$ be any field and let $\F[x]$ be the ring of polynomials in the
indeterminate $x$, with coefficients in the field $\F$. Prove that there are
infinitely many irreducible polynomials in $\F[x]$. (Hint: Euclid's proof of the
corresponding result for the ring $\Z$ of integers works here, too.)

\item Let $\alpha=\root\of{2+\root\of 2}$, and prove that the
Galois group ${\rm Gal}(\Q(\alpha)/\Q)$ is cyclic of order 4.






    
\end{enumerate}
%\end{Large}
\end{document}














