% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass[12pt]{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\pagestyle{empty}
\vcorrection{-1.0in}
\hcorrection{-0.8in}
%\textwidth 6.0in
%\textheight 9.0in
\textwidth 6.5in
\textheight 9.25in

\begin{document}
%\begin{Large}

\begin{center}
%\begin{LARGE}
{\bf  Algebra Qualifying Exam}\\ 
{\bf August 23, 2001}\\
%\end{LARGE}
\end{center}

\vspace{0.1in}
\noindent\hrulefill\\
{\bf Instructions:} You are given 10 problems from which you are to do 8.
 Please indicate those  8 problems which you would like  to be graded 
by circling the problem numbers on the  problem sheet. 

\noindent{\bf Note:} All rings on this exam are associative and
have multiplicative identity 1. All  integral domains are 
assumed to be commutative.
\vspace{0.2in}

%\vs


\begin{enumerate}

\item Suppose that $G$ is a group and that $N_1,\, N_2\trianglelefteq G$ are
two normal subgroups with $G/N_1\cong G/N_2$. Must it follow that 
$N_1\cong N_2$? Prove or supply a counterexample.

\item Let $G$ be a group, and let $H$ be a cyclic normal subgroup of $G$. Prove
that the commutator subgroup $G'$ is a subgroup of the centralizer
$C_G(H)=\{g\in G\,|\, gh=hg, \mbox{ for all }h\in H\}.$


\item Let $G$ be a finite group and let $P$ be a $p$-Sylow subgroup
of $G$. If for all $x\in G$ we have $P\cap P^x=1$ or $P$, show that
$|\mbox{Syl}_p(G)|\equiv 1\, (\mbox{mod }|P|)$.

%\item Let $P$ be a prime ideal of the commutative ring $R$. Show that
%$P[x]$ is a prime ideal of the commutative ring $R[x]$.

\item $\K$ be a field and let ${\cal O}\subseteq\K$ be
a subring. Assume
that for all $0\not=\alpha\in\K$, we have either $\alpha\in {\cal O}$ or
$\alpha^{-1}\in{\cal O}$. Prove that the set of nonunits of ${\cal O}$
forms an ideal of ${\cal O}$.

\item Let $\F$ be a field and let $R$ be the ring of all $2\, \times\, 2$
matrices over $\F$. Prove that, up to isomorphism, $R$ has only one
irreducible left $R$-module $M$, viz., 
$$M\ =\ \left\lbrace\left[ \begin{array}{c}
                 a \\
                 b
                 \end{array}\right]\vert\ a,b\in {\F}\right\rbrace,$$
acted on by ordinary matrix multiplication.

\item Let $\F$ be a field, let $V$ be a finite-dimensional $\F$-vector space,
and let $T:V\to V$ be a linear transformation. If the minimal polynomial
$m_T(x)$ splits into {\em linear} factors, prove that there exists an
ordered basis of $V$ relative to which $T$ is represented by an upper
triangular matrix.

%\item Let $\F=\F_2$ be the finite field of two elements,  and let $k$ be
%a positive integer. Let $W$ be a $k$-dimensional $\F$-vector space, and
%let $\{w_1,w_2,\ldots, w_n\}$ be the set of nonzero vectors of $W$.
%If $V$ is the $\F$-vector space  
%$$V=\{(a_1,a_2,\ldots, a_n)\,|\ a_1,a_2,\ldots ,a_n\in \F\},$$ 
%and if $T:V\to W$ is the linear transformation defined by
%$$T(a_1,a_2,\ldots ,a_n)=\sum\limits_{i=1}^na_iw_i\in W,$$
%compute the dimension of $\ker T$ in wholly terms of the parameter $k$.

\item Let $V$ be an $n$-dimensional vector space and let $V^*$ be the
dual space.  Let $\{f_1,f_2,\ldots f_r\}$ be linearly independent
functionals
in $V^*$, and let $W:=\cap  _{i=1}^r\mbox{ann}({f_i})$, where
$\mbox{ann}({f_i})$
is the subspace of $V$ annihilated by $f_i$.  Prove that $\mbox{dim
}W=n-r.$

\item Let ${\F}\subseteq {\E}\subseteq {\K}$ be fields,
let $\alpha\in {\K}$, and let $f(x)$ be the minimal polynomial of
$\alpha $ over ${\F}$.  Assume that $[{\E}:{\F}]$ and
$\mbox{deg }f(x)$ are relatively prime. Prove that $f(x)$ is also the
minimal polynomial of $\alpha $ over ${\E}$.

\item Let $f(x)\in\Q[x]$ be an irreducible polynomial 
of degree at least 2, and let 
$\alpha\in\C$ be a root of $f(x)$. Give an example to show that
the Galois group of $\Q(\alpha)$ over $\Q$ can be trivial.

\item Let $f(x)\in\Q[x]$ be an irreducible polynomial and let
$\K\subseteq \C$ be the splitting field of $f(x)$. Show that
${\rm Gal}(\K/\Q)$ must act transitively on the roots of $f(x)$.






    
\end{enumerate}
%\end{Large}
\end{document}














