% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
\begin{Large}






\begin{center}\begin{LARGE}
{\bf  Algebra Qualifying Exam}\\ 
{\bf August 31, 1999}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\
{\bf Instructions:} You are given 10 problems from which you are to do 8.
 Please indicate those  8 problems which you would like  to be graded 
by circling the problem numbers on the  problem sheet. 

\noindent{\bf Note:} All rings on this exam are associative and
have multiplicative identity 1. All all integral domains are 
assumed to be commutative.
\vspace{0.2in}

\vs


\begin{enumerate}

\item Let $P$ be a $p$-Sylow subgroup of the finite group $G$.
Prove that $N_G(N_G(P))=N_G(P).$

\item Let $G$ be a finite group and let ${\cal C}$ be a conjugacy class
of elements in $G$.  If $|{\cal C}|=\frac{1}{2}|G|$, show that every
element of ${\cal C}$ is an involution (i.e., an element of order 2).

\item Let $x$ be an element of $p$-power order in the 
finite group $G$, where $p$
is prime. Assume that $|\{g^{-1}xg|\ g\in G\}|=p$. Show that $x$ lies in
a normal $p$-subgroup of $G$.

\item Prove, or give a counterexample to the assertions below:
\begin{description}
\item{(a)} $\Z[x]$ is a principal ideal domain.
\item{(b)} If $I$ is a maximal ideal of $\Z$, then $I[x]$
is a maximal ideal of $\Z[x]$.
\end{description}

\item Consider the commutative ring $R=\{a+b\,\root\of{-5}|\ a,b\in
\Z\}
\subseteq \C.$  Show that the element $1+2\,\root\of{-5}$ is
irreducible but not prime in $R$.


%\item Let $R$ be a ring and assume that $R=
%\oplus _{i=1}^nI_i$, where $I_1,\ldots I_n$ are minimal left ideals
%of $R$.  If $M$ is an irreducible left $R$-module, prove that $M\cong
%I_j$ for some index $j,\ 1\leq j\leq n$.

\item Let $R$ be a ring and let $M$ be an irreducible left $R$-module.
If $K$ is the kernel of the action of $R$ on $M$ ({\em i.e.}, 
$K=\ker (R\to\mbox{End}_\Z(M)))$, prove that $R/K$ is semisimple,
{\em i.e.}, the Jacobson radical is trivial. (Hint: the problem
itself is trivial.)
\clearpage

\item Let $\F$ be a field and let
$$0\to V_1\to V_2\to V_3\to V_4\to 0$$
be an exact sequence of finite dimensional vector spaces over $\F$. Prove
that 
$$\dim V_1-\dim V_2 +\dim V_3 -\dim V_4 =0.$$

\item Let $\F$ be a field and let $T:V\to V$ be a linear transformation
on $V$. Assume that $T$ has elementary divisors $x-a, (x-a)^2,(x-a)^2,
(x-b)^2, x-c, x-c$, where $a,b,c\in\F$ are distinct elements of $\F$.
\begin{description}
\item{(i)} What is the dimension of $V$?
\item{(ii)} What is the minimal polynomial of $T$?
\item{(iii)} What are the invariant factors of $T$?
\item{(iv)} Compute the Jordan canonical form of $T$.
\end{description}

\item Let $\F\subseteq \K$ be fields such that the extension degree
$[\K:\F]<\infty $. Prove that every element of $\K$ is algebraic 
over $\F$.

\item Let $G$ be a finite {\em Hamiltonian} group, {\em i.e.}, one such that
every subgroup of $G$ is normal. Now assume that $f(x)\in\Q[x]$ is an
irreducible polynomial whose Galois group is isomorphic to $G$. 
Prove that $\mbox{deg }f(x)=|G|$.






    
\end{enumerate}
\end{Large} \end{document}
















