


% latex file
%\def\hcorrection#1{\advance\hoffset by #1 }
%\def\vcorrection#1{\advance\voffset by #1 }

\documentstyle[12pt,amscd,amssymb] {amsart}
\pagestyle{empty}
%\input amssym.def
%\input amssym

\oddsidemargin -.5in
\evensidemargin -.5in
\topmargin -0.5in 
\textwidth 7.0in
\textheight 9.0in
\begin{document}
%\begin{large}
\newcommand{\C}{{\Bbb C}}
\newcommand{\Q}{{\Bbb Q}}
\newcommand{\Z}{{\Bbb Z}}
\newcommand{\F}{{\Bbb F}}
\newcommand{\Gal}{\mbox{\rm Gal}}
\newcommand{\Aut}{\mbox{\rm Aut}}


\begin{center}\begin{LARGE}
{\bf  Algebra Qualifying Exam}\\ 
{\bf January  29, 2002}\\
\end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\
{\bf Instructions:} You are given 10 problems from which you are to do 8.
 Please indicate those  8 problems which you would like  to be graded 
by circling the problem numbers on the  problem sheet.  

\vspace{0.2in}

{\bf Note:} All rings in this exam are associative and with 1 and 
all integral domains are commutative.


% $\Bbb Q $ and $ \Bbb C $ are
%the fields of  rational and complex  numbers respectively.

%$\Bbb Z $ and $ \Bbb Q $ are
%the sets of the integers and rational numbers respectively.

\vspace{0.2in}

\begin{enumerate}
 \item Let $G$ be an arbitary group and $H_1, H_2, \dots, H_n$ be subgroups
 of $G$ of with finite index in $G$. Show that there is a normal subgroup
 $K$ of finite index such that $ K \subseteq H_i $ for all $ i=1, 2, \dots, n$.

\vspace{0.2in}

%\item Let $r\geq 3$ be a positive integer and  $ \hat{W}$ be the subgroup of all  
%permutations $w$ of the set of
% integers $\Z $ such that 
%\[w(i+r)=w(i)+r \text{ and } w(i)\equiv i \pmod{i} \text{ for all }  \in \Z. \]
%Show that $\hat{W} $ is a free abelian group of rank $r$ with basis $ \{w_1,
%\dots, w_r\} $ defined by 
%\[ w_i(j)=\left \{ \begin{array}{lcl}
%j+r & \text{ if } & j\equiv i \pmod{r}\\
%j+r & \text{ if } & \text{otherwise.} \end{array} \right . \]

\item  Let $\F$ be a field and $G$ be the group 
\[ G=\left\lbrace \begin{pmatrix} 1&a&b&c\\0&1&d&e\\0&0&1&f\\0&0&0&1\end{pmatrix}\,|\  a,
 b, c, d, e, f \in \F \right\rbrace \]
with usual matrix multiplication. Prove that the group $G$ is nilpotent.
% by computing the lower contral series.

\vspace{0.2in}
 

\item Let $ R$ be a commutative ring. Show that if $I_1, \cdots, I_n $ are
ideals with $ \cap _{i=1}^n I_i \subseteq P$ for some prime ideal $P$, then
there is an $i $ such that $ I_i \subseteq P$.

\vspace{0.2in}
 
\item Let $p$ be a fixed prime number. Define 
$\Z(p)=\{ ap^{-n} \in \Q\, |\ a \in \Z, n \in \Z\}$. 

\vspace{0.2in}

\begin{enumerate}
\item Show that $ \Z(p) $ is a
subgring of $ \Q $ the field of rational numbers $\Q$ containing $\Z$ as
a subring.\\
\item Assume that $A$ is a ring and that $ \phi: \Z \rightarrow A $
is a ring homomorphism such that $\phi(p) $ has an muliplicative inverse in $A$. 
Show that there
exists a unique ring homomorphism $ \psi: \Z(p)\rightarrow A$ such that 
$\psi(x)=\phi(x) $ for all $ x \in \Z$.
\end{enumerate}

\vspace{0.2in}

\item Let $A$ be a commutative ring, and let $ A[x]$ be the ring of all
polynomials $f(x)=\sum_{i=0}^na_i x^n$ in the indeterminate $x$ with 
coefficients $a_i \in A$.
Show that for any two polynomials $f(x) $ and $ g(x)$
such that the leading coefficient of $g(x)$ is invertible in $A$, then
there is a unique pair of polynomials in $q(x)$ and $ r(x) $ such that 
\[f(x)\ =\ g(x)q(x)\ +\ r(x),\]
where $\deg(r(x))<\deg(g(x))$. Can you conclude that the ring $A[x]$ is an
Euclidean domain? Justify your conclusion.

\vspace{0.2in}

\item Let $D$ be a unique factorization domain and $ Q(D) $ be the field of
factions (quotient field of $D$). Suppose  $ a, b, f, g $ are in $ D$ 
with $ b\neq 0 $ and $ g\neq 0 $ such that $ \frac{a}{b}=\frac{f}{g}\in Q(D)$. If $
a $ and $ b$ are relatively prime, show that $ a|f$ and $b|g$. 

\vspace{0.2in}

%\item Let $F$ be a field and $F[x]$ be the ring of polynomials with
%coefficients in $F$. Let $ F(x)$ be the field of fractions of $ F[x]$. If $
%p(x), q(x), f(x), g(x) $ are in $F[x]$ such that $
%\frac{a(x)}{b(x)}=\frac{f(x)}{g(x)}$ in $F(x)$, show that $ \deg(a(x))\leq
%\deg(f(x))$ and  $\deg(b(x))\leq
%\deg(g(x))$.
 
\item Let $\F$ be field and $R=M_n(\F) $ be the ring of all $n\times
n$ matrices with entries in $\F$. Show that the vector space $ \F^n $ of
$n\times 1$ column vectors with coefficients in $\F$
is an irreducible module of $R$ (relative to ordinary matrix multiplication). 
Then show that any
finite-dimensional $R$-module  is a finite-dimensional vector space of
dimension divisible by $n$. 

\vspace{0.2in}

\item Let $\F_q$ be a finite field with $q$ elements. Compute the order of
the group ${\rm GL}_n(q)$ of all invertible linear transformations of the vector space 
$\F_q^n$. 
%(Hint, If $\{v_1,\dots, v_r\} $ a linearly independent set of
%vectors, then for any vector $v$, $\{v_1,\dots, v_r, v\}$ is linearly
%dependent if and only if $ v \in \text{Span}(v_1, \dots, v_r)$.
    
\vspace{0.2in}
%\item Let $ \F_q$ be a finite field with $q$ elements. Compute the number of
%$k$-dimensional subspaces of $\F_q^n$ in terms of $q$.

\item Let $ \F $ be a field, let $x$ be an indeterminate over $\F$,
and set $ R=\F[x]/(x^5(x-2)^6) $. Describe all
 $ R $-modules of $\F$-dimension 25 up to $ R $-module
isomorphism.  
%(Any $R$-module can be regarded as a $\F[T]$-module, what are
%the possible matrices of $T$ on $\F^{25}$.)

\vspace{0.2in}

\item Let $q$ be a prime power. Show that the finite field $ \F_{q^r}$
is isomorphic to a subfield of $ \F_{q^n}$ if and only if $ r$ divides $n$.
If $ r $ divides $n$, then show that $ \F_{q^n}$ is a Galois extension of
$\F_{q^r}$ and compute the Galois group of $\F_{q^n} $ over $\F_{q^r}$.
  \end{enumerate}
%\end{large}
\end{document}





