% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}

\begin{center}\begin{LARGE}
{\bf Algebra Qualifying Exam}\\ 
{\bf Spring 1996}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

{\bf Note:} All rings in this exam are associative and with 1 and all
integral domains are commutative. $\Bbb Q$ and $\Bbb C$ are the fields
of rational and complex numbers respectively.


\begin{description}

\item[1.]
Show that for any group $G$, the quotient group $G/Z(G)$ is never a
nontrivial cyclic group. Here $Z(G)$ is the center of the group $G$.

\item[2.]
Let $G$ be a finite group and $p$ a prime number dividing $|G|$. If $P$
is a $p$-Sylow subgroup of $G$, show that $N_G (N_G(P)) = N_G(P)$.
Here $N_G (H)$ is the normalizer of a subgroup $H$ in the group $G$.

\item[3.]
Let $R$ be a commutative ring with 1 and $x \in R$. Suppose that $x$ lies
in every maximal ideal of $R$. Show that $1-x$ is a unit of $R$.

\item[4.]
This problem conisits of two parts:

\item[\quad] (i)
Give a definition to what a unique factorization domain (UFD) means.

\item[\quad] (ii)
Give an example of an integral domain that is not a UFD.

\item[5.]
Let $R$ be a ring with 1. Suppose that an $R$-module $M = M_1 \oplus M_2$
is a direct sum of two non-isomorphic irreducible submodules $M_1$ and
$M_2$. Show that $M_1$ and $M_2$ are the only two proper submodules of $M$.

\item[6.]
This problem consists of two parts:

\item[\quad] (i)
State the definition of what an indecomposable module over a ring means.

\item[\quad] (ii)
Give an example of a ring $R$ and an indecomposable module $M$ over $R$.

\item[7.]
Let $T: V \to V$ be a linear transformation on a finite dimensional vector
spaces over a field $F$. Suppose $T$ has the following invariant factors:
$$1+x, \quad x(1+x), \quad x(1+x)^2.$$
Answer the following questions:

\item[\quad] (i)
What is $\dim_F V$?

\item[\quad] (ii)
Is $T$ onto?

\item[\quad] (iii)
Does $T$ have a Jordan form over the field $F$ with respect to an
appropriate basis of $V$? If yes, write down such a matrix.

\item[\quad] (iv)
Is $V$ indecomposable as an $F[T]$- module?

\item[\quad] (v)
What is the minimal polynomial of $T$?

\item[8.]
Let $V$ be the 4-dimensional vector space of all $2 \times 2$- matrices
over a field $F$. We define the function $f(\cdot, \cdot) : V \times V \to F$
by $f(A,B) =$ trace $(AB)$ for all $A, B \in V$.

\item[\quad] (i)
Show that $F$ is a symmetric bilinear form;

\item[\quad] (ii)
Let $\{E_{11}, E_{12}, E_{21}, E_{22} \}$ be the standard basis of $V$.
Compute the matrix of the bilinear form $f$ with respect to the standard
basis.

\item[\quad] (iii)
Describe the radical of the bilinear form $f$. Here the radical of a
bilinear form $f$ is defined to be the set
$\{v \in V | f(v,V) = 0\}$.

\item[9.]
This problem consists of two parts:

\item[\quad] (i)
Define what it means for a field extension $E$ over $F$ to be separable.

\item[\quad] (ii)
Is the splitting field of the polynomial $p(x) = x^9 - x^3 + 1$ over
$\Bbb F_3$ separable?

\item[10.]
Show that, if $F \lneqq K \leq F(x)$ is tower of fields, where $x$ is an
indeterminate and $F(x)$ is the field of rational functions (i.e., $x$ is
transcendental over $F$), then $K$ cannot be an algebraic extension over
$F$.




\end{description}    
%\end{Large}
\end{document}














