
%-*-latex-*-            This is using LaTeX mode for emacs%

%\documentstyle[12pt]{article}
\documentclass[12pt]{article}
\topmargin -.5in
\textwidth 6.1in
\textheight 9.0 in
\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)


\renewcommand{\labelenumi}
 {\mbox{ \mbox{{\bf {\roman{enumi}}}})}}
 \renewcommand{\theenumi}{\roman{enumi}} 
% This changes the numbering in itermizing to roman.
\newcounter{problem}\setcounter{problem}{0}
\renewcommand{\theproblem}{\arabic{problem}}
\newcommand{\problem}{%\setcounter{subsubsection}{0}
 	\par \vspace{1em} 
 	 \stepcounter{problem} 
 		\noindent {\bf {\theproblem}. }
 	}


\begin{document}

 \begin{center}

 {\bf Geometry/Manifolds QUALIFYING EXAM \ \ Spring 1997}

	(Yetter and Wu)
 \end{center}

% Answer {\it all} questions in {\bf Part A} and choose {\bf 3} problems in
% {\bf Part B}.

Work out problem {\bf 1} and then choose {\bf 5} (and only 5) 
additional problems among the remaining ones. Start each 
problem on a new sheet of paper. 
{\bf Do not turn in more than 6 problems including the required number 1.}
A space always means a topological space below.

\vspace{20pt}
\bigskip \problem {\bf This one is required!}

Answer the following questions and give a brief explanation or
counterexample:
\begin{enumerate}
\item 	\begin{enumerate}
	\item Give an examples of orientable, connected, simply-connected
		smooth 2-manifold with trivial tangent bundle.
	\item Give an examples of orientable, connected, simply-connected
		smooth 2-manifold with non-trivial tangent bundle.
	\end{enumerate}
\item What is the dimension of the total space of the exterior bundle
	$\Lambda^3(M)$ if $M$ is a smooth $6$-manifold?
\item Give an example of a compact non-orientable manifold.
\item Consider the vectorfield on ${\Bbb R}^2\setminus \{ (0,0) \}$ given
	in polar coordinates by $(r-1)d\theta$.  Describe in detail its flows.
	Is this vectorfield complete?
\item \begin{enumerate}
	\item What is the fundamental group of $S^1 \times S^1$?
	\item What is the fundamental group of 
	$S^1 \times S^2$?
	\item What is the fundamental group of 
	${\Bbb R}^2 \setminus \{ (0, 1), (0, -1)\} $?
       \end{enumerate}
\item Define a connection on the tangent bundle of a manifold.  Define the
Levi-Civita connection.
\item Give an example of a space which is locally euclidean, but is not
a manifold.

\end{enumerate}


\vfill {\hfill (continued on next page)}
\eject

{\bf Choose 5 problems from below:}
\bigskip

\bigskip \problem Let $\Gamma$ be the ellipsoid $x^2 + \frac{y^2}{4} + \frac{z^2}{9} = 1$
in ${\Bbb R}^3$ calculate

\[ \int_\Gamma z dx\wedge dy - y dz\wedge dx \]

\bigskip \problem \begin{enumerate}
	\item Describe the natural Lie algebra structure on the set of
vectorfields on a smooth manifold $M$.
	\item In the case where $M$ is a Lie group, use the group law
and the Lie algebra structure in {\bf i)} to construct a Lie algebra structure
on the tangent fibre at
the identity $T_e(M)$.
\end{enumerate} 

\bigskip \problem
Suppose G is a connected compact Lie group.  Show that the fundamental
group of G, $\pi_1(G)$ is abelian.

\bigskip \problem
Find the scalar curvature of the surface $z = x^2 + y^2$ at $(0,0,0)$.

\bigskip \problem
Show that every  $1$-form on ${\Bbb R}^1$ is {\it exact}. 
	Show that every  {\it closed} $1$-form on ${\Bbb R}^3$ is {\it
	exact}.

\bigskip \problem
Let $\varphi: M\rightarrow N$ be a (smooth) map. Then the vector
field $X$ on $M$ and $Y$ on $N$ are said to be $\varphi$-related if
$ d\varphi_m(X_m) = Y_{\varphi(m)}$ for all $m\in M$.

Let $X_1, X_2$ be vector fields on $M$ and $Y_1, Y_2$ vector fields on $N$.
Assume that $X_1$ is $\varphi$-related to $Y_1$ 
and $X_2$ is $\varphi$-related to $Y_2$. Show that
$[X_1, X_2]$ is $\varphi$-related to $[Y_1, Y_2]$.

\bigskip \problem
Suppose $f: X^d\rightarrow {\Bbb R}^{d+1}$ is a (smooth) embedding of the
$d$-dimensional manifold $X$ into ${\Bbb R}^{d+1}$. A normal vector field
along $(X, f)$ is a smooth map $N: X\rightarrow T({\Bbb R}^{d+1})$ such that
for each $p\in X$, $N(p)\in T_{f(p)}{\Bbb R}^{d+1}$ and it ($N(p)$) is
orthogonal to the subspace $df(T_pX)\subset T_{f(p)}{\Bbb R}^{d+1}$.
Prove that the manifold $X$ is orientable if and only if there is a smooth
{\it nowhere-vanishing} normal vector field along $(X, f)$.

\bigskip \problem
Prove that $S^n$ ($n\geq 1$) is orientable.

\bigskip \problem
Let $f: M\rightarrow N$ be a smooth map such that for all $m\in M$,
$df_m: T_mM\rightarrow T_{f(m)}N$ is {\it surjective}. Show that for any 
$n\in N$, $f^{-1}(n)\subset M$ is a smooth submanifold of $M$. What is the
dimension of $f^{-1}(n)$?

\end{document}

  






