% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Geometry Manifolds Qualifying Exam}\\ 
{\bf Spring 1995}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Part A. Do all parts.

\begin{description}

\item[1.]
A Riemannian metric is a cross section of what bundle on a manifold?

\item[2.]
Is the two handled torus (connected sum of two $S^1 \times S^1$'s)
parallelizable?

\item[3.]
Give an example of a compact manifold which is not orientable.

\item[4.]
If we are given a 3-form on the unit ball in $R^3$, when will Stoke's
theorem allow us to rewrite its integral as an integral on $S^2$?

\item[5.]
Describe the universal covering space of (a) $S^2$, (b) $S^2 \times S^1$.

\item[6.]
Consider the vector field on $R^3 (y^2+z^2+1) \partial / \partial x$. Describe
the family of its flows.

\item[7.]
What is the dimension of the fiber of the bundle of 7-forms on a 9-manifold?

\item[8.] (a)
Give an example of a lie group which is contractible as topological space
and has dimension seven.

\item[\quad] (b)
Give an example of a lie group which is not contractable as a
topological space.  Is its lie algebra contractible?

Part B. Calculate -- Do 3 of the following 5.
\item[1.]
Let an atlas for the 2-sphere be given by choosing stereographic
projection from two antipodal points. Pick a geodesic joining the two as
0-ray and write polar coordinates on each patch.

\item[\quad] (a)
Find the transition function $(r, \theta) \to (r^\prime, \theta^\prime)$.

\item[\quad] (b)
Write the round metric of radius 1 in each patch.

\item[\quad] (c)
Find the $\left\{ \theta^\theta_r \right\}$ component of the Levi-Civita
connection in one patch.

\item[2.]
Write generators and relations for $\pi_1$ of the once punctured torus.

\item[3.]
Give a set of generators for the lie algebra $su(2)$, and compute the
bracket of each pair.

\item[4.]
Use Stokes theorem to compute $\int_{S^2} zdx \wedge dy - ydx \wedge dz$
on the unit sphere in $R^3$.

\item[5.]
Find the scalar curvature of the surface $z=x^2-y^2$ at the point
$(0,0,0)$.



\end{description}    
%\end{Large}
\end{document}














