\documentclass[12pt]{article}
\usepackage{amsxtra,amssymb,amsthm,amsmath,latexsym}


\setlength{\parindent}{0cm}

\def\ds{\displaystyle}
\def\R{\mathbb R}
\def\Z{\mathbb Z}


\pagestyle{empty}

\textwidth=7truein
\textheight=9.25truein
\hoffset=-1truein
\voffset=-.75truein

\begin{document}

\begin{large}

\begin{center}
 {\bf
\begin{Large}
  Geometry of Manifolds\\
  Fall 2001  Qualifying Exam \end{Large} \\
  Auckly \& Miller}
\end{center}


\vspace{.2in}



Notes:
Work as many of the following problems as you can. Good luck.

%\vspace{.75in}

\begin{description}

\item[1.]
Define $f:\R^3\to \R^2$ by
$ f(x,y,z)=(x+z^2, y-z^2)=(u,v) $
where $(x,y,z)$ are coordinates on $\R^3$ and $(u,v)$ are
coordinates on $\R^2$. Let $\alpha=du+u\,dv$. Find

\item[\quad] a) $f^\ast(\alpha)$
\item[\quad] b) $f^\ast(d\alpha)$
\item[\quad] c) $\int_Sf^\ast(d\alpha)$ where $S$ is the surface
$x^2+y^2+z^2=1$, $z\geq 0$
oriented with the upward pointing normal.

\item[2.] a)
Construct an example of a covering projection, $P:E\to X$
with $\hbox{Deck}(E)\cong \Z_2\oplus \Z_3$.

\item[\quad] b)
Compute $H^\infty(X:\R)$ for your example.


\item[3.]
On $\R^3$ let
$\ds X=xz^2\frac{\partial}{\partial x}+y\frac{\partial}{\partial y}
+z\frac{\partial}{\partial z}.$

\item[\quad] a) Calculate $\ds\int_BL_X(dx \wedge dy \wedge dz)$
where $B=\{(x,y,z)|x^2+y^2+z^2\leq 1\}$
with orientation given by $dx \wedge dy \wedge dz$.

\item[\quad] b)
Compute the flow of $X$, $F:\R\times\R^3\to \R^3$.


\item[4.]
Consider a $2\hbox{-dimensional}$ Riemannian manifold $(M,g)$ where
$M$ is an open subset of
$\R^2=\{(x,y)|x\in \R, y\in \R\}$. Suppose $g$ is such that
$\ds e_1=\frac{\partial}{\partial x}$
and $\ds e_2=f(x,y)\frac{\partial}{\partial x}+\frac{\partial}{\partial y}$
is an orthonormal frame.

\item[\quad] a)
Find the coframe $(\theta^1,\theta^2)$ dual to $(e_1,e_2)$.

\item[\quad] b)
Find a $1\hbox{-form } w$ so that $d\theta^1=w\wedge\theta^2$ and
$d\theta^2=-w\wedge\theta^1$.

\item[\quad] c)
Find a function $K(x,y)$ so that $dw=-K\theta^1\wedge \theta^2$.

\item[\quad] d)
Show how to choose $f(x,y)$ so that $K(x,y)=-1$ for all $(x,y)$.


\vfill\eject

\item[5.]
Let $K$ be a simplicial complex.

\item[\quad] a)
Define the Euler characteristic of $K$.

\item[\quad] b)
Let $p:L\to K$ be an $n\hbox{-fold}$ simplicial cover.
Show that $\chi(L)=n\chi(K)$.

\item[\quad] c)
Determine all {\bf closed} $2\hbox{-manifolds}$ that cover $T^2$.

\item[\quad] d)
Give an example of a non-trivial cover $q:T^2\to T^2$.

\item[\quad] e)
Compute $\chi(S^n)$ as a function of $n$.

\item[\quad] f)
Let $G$ be a finite group that acts freely on $S^{2k}$.
Assume that the map $\pi:S^{2k}\to S^{2k}/G$ is a covering projection.
Prove that $G\cong \Z_2$ or $G$ is trivial.


\item[6.] a)
Given that $X$ is a non-vanishing vector field on $M$, prove that there
is a diffeomorphism, $f:M\to M$ without fixed points (i.e., $x$ such that
$f(x)=x$) that is homotopic to $\hbox{id}:M\to M$.

\item[\quad] b)
Given that $f:S^n\to S^n$ has no fixed points, show that $f$ is homotopic
to $p:S^n\to S^n$; $p(x)=-x$.

\item[\quad] c)
Compute $H^n(p):H^n(S^n,\R)\to H^n(S^n,\R)$.

\item[\quad] d)
What can be said about the dimension of a sphere that admits
a non-vanishing vector field.

\item[\quad] e)
Let $G$ be a Lie group that acts freely (i.e., $x\cdot g=x$ implies $g=1$)
on $S^{2k}$. Prove that $\dim G=0$.


\item[7]
Let $G$ be a Lie group.



\item[\quad] a)
Define the Lie algebra of $G$, say $\mathfrak g$.

\item[\quad] b)
Let $\mathfrak h$ be a Lie subalgebra of $\mathfrak g$.
Explain how to construct a distribution on $G$ using $\mathfrak h$ and prove
that this distribution is integrable.

\item[\quad] c)
Let $H$ be the maximal connected leaf of the distribution from part b)
passing through $1$.
Construct an atlas on $H$.

\item[\quad] d)
Prove that $H$ is closed under multiplication.
\newline
Hint: Consider $L_gH:=\{gh|h\in H\}$.






\end{description}
\end{large}
\end{document}


