
\documentclass[12pt]{article}
\usepackage{amsxtra,amssymb,amsthm,amsmath,latexsym}

\setlength{\parindent}{0cm}
\def\ds{\displaystyle}

\def\R{{\mathbb R}}
\def\N{{\mathbb N}}
\def\Z{{\mathbb Z}}
\def\C{{\mathbb C}}
\def\ds{\displaystyle}
\def\nd{\noindent}

\def\ra{\rightarrow}
\def\lra{\longrightarrow}

%\pagestyle{empty}

\textwidth=7.25truein\hoffset=-1truein
\textheight=9.5truein\voffset=-1.25truein

\begin{document}

\begin{large}

\begin{center}
{\bf
Geometry of Manifolds Qualifying Exam\\
Auckly and Crane, Fall 2002
}\end{center}

\vspace{.15in}


\begin{description}

\item[1.] (A)
What is $H^2(\R^3)$?

\vspace{.15in}
\item[\quad] (B)
Write Stokes' Formula.

\vspace{.15in}
\item[\quad] (C)
What kind of differential form can one integrate on a $3\hbox{-surface}$
embedded in a $5\hbox{-manifold}$?

\vspace{.15in}
\item[\quad] (D)
Which of the following topological spaces are simply connected?

\begin{tabular}{llll}
\rule{1.15in}{0in}&\rule{1.15in}{0in}&\rule{1.15in}{0in}&\rule{1.15in}{0in}\\
a) $S^3$  & b) $RP^3$ & c)$S^1\times S^2$ & d) $\C-\{0\}$
\end{tabular}

\vspace{.15in}
\item[\quad] (E)
Which lie algebras are contractible as topological spaces?

\vspace{.15in}
\item[2.]
Compute $H^\ast(RP^2\times S^2, \Z)$ and $\pi_1(RP^2\times S^2)$.

\vspace{.15in}
\item[3.]
Let $F=\R\times (0,2\pi)$ and let the metric $g$ be given by
$\ds\frac{2z^2+1}{z^2+1}dz^2+(z^2+1)d\theta^2$.

\vspace{.15in}
\item[\quad] (A)
Compute $\nabla_{\partial_\theta}\partial_\theta$ for the Levi-Civita
connection corresponding to $g$.

\vspace{.15in}
\item[\quad] (B)
Compute the sectional curvature of $(F,g)$.

\vspace{.15in}
\item[4.]
Give an example of a topological space, every point of which has a
neighborhood homeomorphic to $(0,1)$, which is not a manifold.

\vspace{.15in}
\item[5.]
Let $(X,X_0)$ be a pointed topological space, $G$ a lie group
and $C^0[(X,X^0),\,(G,1)]$ the set of continuous pointed maps.
This forms a group under pointwise multiplication.
Since $\pi_1$ is a functor, we obtain a map
$\widetilde\pi:C^0[(X,X^0),\,(G,1)]\to \hbox{Hom}(\pi_1(X,X_0),\,
\pi_1(G,1))$ given by $\widetilde\pi,(u)([\ell(t)])\to[u(\ell(t))]$
where $\ell(t)$ is a loop in $(X,X_0)$.
Prove that $\widetilde\pi$ is a group homomorphism when
$\hbox{Hom}[\pi_1(X),\, \pi_1(G)]$ is viewed as a group under pointwise
multiplication. (Recall that $\pi_1$ of a lie group is abelian so this is
in fact a group.)

\vspace{.15in}
\item[6.]
Prove that if $p:G_1\to G_2$ is a smooth homomorphism of connected
lie groups which induces an isomorphism of lie algebras,
then $p$ is a covering projection. [Recall that a covering projection is
a surjective continuous map such that for any $x\in G_2$ there is
a neighborhood $U$ such that $p^{-1}(U)$ is a disjoint union of
open sets $V_\alpha$ and $p\vert_{V_\alpha}$
is a homeomorphism for each $\alpha$ onto $U$.]

\vspace{.15in}
\item[7.]
Let $X$ be a complete vector field on a manifold $M$.
This means that $X$ has a global flow,
$\varphi :\R\times M\to M$.
Let $f:M\to \R$ be a smooth function with $X(f)(p)=1$ for all
$p\in M$. Prove that $\phi_t(f^{-1}(a))=f^{-1}(a+t)$ for all $a$, $t\in\R$.
Give an example to show that the vector field must be complete for this
to hold.

%\vfill\clearpage



\end{description}
\end{large}
\end{document}
