% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Geometry of Manifolds Qualifying Exam}\\ 
{\bf Fall 1997}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

{\bf Part A}. Do all nine (9) questions in part A.

\begin{description}

\item[1.]
What is the fundamental group of

\item[\quad] (a)
$\Bbb R \Bbb P^2$ (the real projective plane)

\item[\quad] (b)
$S^1 \times S^1$

\item[\quad] (c)
$T(M)$, the total space of the tangent bundle to a simply connected smooth
manifold, $M$.

\item[2.]
Describe in detain the flows of the vectorfield on $\Bbb R^2$ given by
$$-y \frac{\partial}{\partial x} + x \frac{\partial}{\partial y}$$

\item[3.]
Let $\omega$ be the 1-form on $\Bbb R^2$ given by $x(x-1)(y-1)dx$,
and let $R$ be the region
$$\{(x,y) | 0 \leq x \leq 1, 0 \leq y \leq 1\}.$$
Find $\int_R d\omega$.

\item[4.] (a)
Give an example of a compact orientable manifold with non-trivial
tangent bundle.

\item[\quad] (b)
Give an example of a compact orientable manifold with trivial tangent
bundle.

\item[\quad] (c)
Give an example of a compact non-orientable manifold.

\item[5.]
If $M= S^1 \times S^4$, what is the dimension of the fibres of the third
exterior bundle $\Lambda^3(M)$?

\item[6.]
How many non-zero vectorspaces of differential forms are there in the
deRham complex of $S^2 \times S^2$?

\item[7.]
What is the scalar curvature of the surface $3x+2y - z = 0$ in
$\Bbb R^3$ at the point $(0,0,0)$?

\item[8.]
Give an example of a locally Euclidean topological space which is not a
topological manifold.

\item[9.]
State the deRham Theorem.

\end{description}

{\bf Part B}. Choose four (4) and only four of the following problems.

\begin{description}

\item[1.]
On $\Bbb R^3$ with standard Euclidean coordinates $(x,y,z)$, consider the
2-form $\alpha = f(x,y,z) dx \wedge dy + yzdx \wedge dz+ x^2 dy \wedge dz$.
Choose a function $f(x,y,z)$ so that $d \alpha =0$ and
$\alpha |_{z=1} = dx \wedge dy$.

\item[2.] (a)
Define the deRham cohomology groups of a differentiable manifold.

\item[\quad] (b)
Calculate the deRham cohomology groups of the circle $S^1$ directly
from the definition in part (a).

\item[3.]
Give a detailed computation of the fundamental group of the closed compact
surface of genus 2 (a.k.a the ``two-holed torus").

\item[4.] (a)
Write down the deRham cohomology groups for the 4-sphere $S^4$.

\item[\quad] (b)
Suppose that $\omega$ is a differential 2-form on $S^4$ and that
$d \omega = 0$. Show that

\item[\quad] \quad i.
$\omega \wedge \omega = d \phi$ for some 3-form $\phi$.

\item[\quad] \quad ii.
$\int_{S^4} \omega \wedge \omega = 0$.

\item[\quad] \quad iii.
There is at least one point $x \in S^4$ such that
$\omega \wedge \omega (x) = 0$.

\item[5.] (a)
Define what we mean by a Lie group.

\item[\quad] (b)
If $G$ is a Lie group, define its Lie algebra $g$.

\item[\quad] (c)
Apply the construction of b) to determine the Lie algebra of $SO(3)$,
including a derivation of the bracket.

\item[\quad] (d)
Show that the tangent bundle to a Lie group is equivalent to a trivial
(product) bundle.

\item[6.]
Let $(M,g)$ be a Riemannian manifold and $V(M)$ be the smooth vectorfields
over $M$.

\item[\quad] (a)
For $X,Y \in V(M)$ define the Riemannian curvature operator
$R(X,Y): V(M) \to V(M)$.

\item[\quad] (b)
Show that if $M= \Bbb R^n$ and $g$ is the Euclidean metric, then
$R(X,Y)Z = 0$ for all vectorfields $X,Y,Z$.

\item[\quad] (c)
Suppose that $R(X,Y)Z = 0$ for all vectorfields $X,Y,Z$ on an arbitrary
Reimannian manifold $(M,g)$. {\it Sketch} a proof that shows that for
$x \in M$ there is a coordinate system $(x_1, \dots, x_n)$ around
$x$ such that
$$g = \sum^n_{i=1} dx^i \otimes dx^i$$





\end{description}    
%\end{Large}
\end{document}














