% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Goemetry of Manifolds Qualifying Exam}\\ 
{\bf Fall 1994}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\
Part A. Short answers. Answer all of the following.

\begin{description}
\item[1.]
What is the fundamental group of

\item[\quad] (a)
$S^3$

\item[\quad] (b)
$S^1 \times S^1$

\item[\quad] (c)
The Euclidean plane with two pts. deleted.

\item[2.]
A differential 1-form is a section of what bundle?

\item[3.]
What is the topology of the underlying manifold of the lie group
$SU(2)$? What is the topology of the lie algebra
$\begin{small}{\cal {SU}} \end{small} (2)$?

\item[4.]
State Stokes' theorem.

\item[5.] (a)
What kind of differential form can we integrate on a 4-manifold?

\item[\quad] (b)
What kind of differential form can we integrate on a surface in a 4-manifold?

\item[6.] (a)
Give an example of a compact surface whose tangent bundle is trivial.

\item[\quad] (b)
Give an example of a compact surface whose tangent bundle is not trivial.

\item[7.]
Suppose we have a non-zero vector field on $R^n$ all of whose covariant
derivatives with respect to the standard metric and Levi-Civita
connection vanish. Describe the family of its integral curves.

\item[8.]
If $M$ is a 4-dimensional manifold, what is the dimension of the fibers of
its bundle of differential 2-forms.

\item[9.]
Consider the complex of differential forms used to define the de Rham
cohomology of a 5-manifold. How many of these spaces are non-vanishing.

\end{description}

Part B. Do any two of the following problems.
\begin{description}

\item[1.] (a)
Write the metric tensor for the Euclidean plane in polar coordinates.

\item[\quad] (b)
Compute the $\theta^\theta r$ component of the Levi-Civita connection for
the Euclidean plane in polar coordinates.

\item[2.]
Give an example of a topological space every point of which has a
neighborhood homeomorphic to $R^2$ which is not a manifold.

\item[3.]
Prove the Jacobi identity holds for ${\cal {SU}} (3)$.

\item[4.]
Compute the DeRham cohomology of the Torus $T^2$.

\item[5.]
Use Stokes' theorem to compute the area of the unit ball in $R^2$.





\end{description}    
%\end{Large}
\end{document}














