% latex file
% Geometry of Manifolds - Auckly & Miller

\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)

\input prepictex.tex
\input pictex.tex
\input postpictex.tex

%\documentstyle[bbb]{report}
\documentclass[bbb]{report}

\vcorrection{-.7in}
\hcorrection{-1in}

\topmargin  0in
\textwidth 6.4in
\textheight 9in

\mathsurround=2pt

\def\ds{\displaystyle}

\def\R{{\Bbb R}}
\def\Z{{\Bbb Z}}
\def\C{{\Bbb C}}
\def\D{{\Bbb D}}
\def\nd{\noindent}

\begin{document}

%\pagestyle{empty}

\begin{large}

\hfill Name \rule{2.5in}{.01in}
\par
\vspace{.1in}

\begin{center}
   GEOMETRY OF MANIFOLDS QUALIFYING EXAM \\
   Fall 1999 \\
   (Auckly \& Miller ) \\
\end{center}


\vspace{.1in}

\begin{large}\begin{center}
\noindent {\bf Work as many as you can in the 2 hours.\\    %instructions
Best of luck.}
\end{center}\end{large}

\vspace{.05in}

\begin{description}

\item[1.] (A)
On $\R^2$ let
$$\begin{array}{rlll}
  X &=x^2y\frac{\partial}{\partial x}+(x+y)\frac{\partial}{\partial y}&
  \qquad\qquad\qquad\qquad\qquad&\\
  Y &=y\frac{\partial}{\partial x}+x\frac{\partial}{\partial y}&&\\
  \alpha &=-y\,dx+x\,dy&&
  \end{array} $$
\item[] Calculate:
$$\begin{array}{lll}
\hbox{ a)  $[X,Y]$ \qquad\qquad\qquad\qquad   }
& \hbox{b) $L_Y \alpha$\qquad\qquad\qquad\qquad}
& \hbox{c) $L_Y d\alpha$\qquad\qquad\qquad\qquad}
 \end{array}$$

\vspace{.25in}

\item[2.]
On $\R^2$ consider the metric
$g=(1+x^2)dx^2+\frac{1}{2} (dx\otimes dy+dy\otimes dx)+dy^2$.

\item[\quad] a) Compute $\nabla_{\frac{\partial}{\partial x}}(y\,dx)$.
\item[\quad] b) Calculate the sectional curvature.

\vspace{.25in}

\item[3.]
On $\{(x,y)\}\bigg|x^2+y^2<1\}$, let
$g=\ds\frac{dx^2+dy^2}{(1-x^2-y^2)^2}$  be a metric.

\item[\quad] a) Show that
  $\phi(z)=\ds\frac{az+b}{\overline{b}z+\overline{a}}$
  is an isometry if $|a|^2-|b|^2=1$ and $z=x+iy$.

  Hint: Show that if $ w =\phi(z)$ is analytic then
  $d w \otimes d\overline{ w }
    =|\phi'(z)|^2dz\otimes d\overline{z}$.

\item[\quad] b) Using this metric, compute the radius and area
of a circle with Euclidean radius $R<1$ centered at the origin.

\vspace{.25in}

\item[4.]
Let $\alpha$ be a differential 2-form on the 2-sphere $S^2$ with
$\int_{S^2}\alpha=1$. Suppose $f:S^3\to S^2$ is smooth, $S^3$ the 3-sphere.

\item[\quad] a) Show that there exists a 1-form $\theta$ on $S^3$ such that
$f^\ast\alpha=d\theta$.

\item[\quad] b) Define $Q(f)=\int_{S^3}\theta\wedge d\theta$.
Show that this is independent of the choice of $\alpha$ and $\theta$.

Hint: First show that for $\alpha$ fixed it is independent of choice
of $\theta$.


\vspace{.25in}
\item[5.]
Suppose that $F:N\to M$ is a smooth covering mapping and that $M$
is a Riemannian manifold with metric $g$.

\item[\quad] a) Show that there exists a unique metric on $N$ so that
$F$ is a local isometry.

\item[\quad] b) Suppose that $N$ is connected and compact. Determine
the relation between volume $(M)$ and volume $(N)$ in terms of the
fundamental groups of $M$ and $N$.

\vspace{.25in}
\item[6.]
Suppose $M$ is a smooth oriented $n\hbox{-dimensional}$ manifold and
$X$ is a complete vector field which generates the 1-parameter
group of diffeomorphisms $(F_t)_t$.
Suppose that $\mu$ is a differential $n\hbox{-form}$ on $M$
and that $U$ is a relatively compact ($\overline{U}$ is compact) open
subset of $M$.

\item[\quad] a) Show that
$\ds\frac{d}{dt}\bigg|_{t=0}\ds\int_{F_t(U)} \mu=\ds\int_UL_X\mu$.

\item[\quad] b) For $M=R^3$ and $\mu=dx\wedge dy\wedge dz$,
the usual volume element, calculate an expression for $L_X\mu$ for
any vector field
$X=f\frac{\partial}{\partial x}
  +g\frac{\partial}{\partial y}
  +h\frac{\partial}{\partial z}$.
Thus obtain a formula for
$\frac{d}{dt}\bigg|_{t=0}\ \hbox{volume}\,(F_t(U))$.

\vspace{.25in}
\item[7.]
Prove that smooth connected manifolds are topologically homogeneous.
That is, given $p, q\in M$ there is a diffeomorphism
$f:M\to M$ so that $f(p)=q$.

\vspace{.25in}
\item[8.]
On $\R^2$ with coordinates $(x^1,x^2)$ let a connection
$(\Gamma^i_{jk})$ be given by $\Gamma^1_{11}=\frac{\partial f}{\partial x^1}$
and $\Gamma^2_{22}=\frac{\partial f}{\partial x^2}$ and all other
$\Gamma^i_{j k}=0$
Here $f:\R^2\to R$ is some given function.

Let $P_0(x^1_0, x^2_0)$ be a given point. If
$v:[a,b]\to \R^2$ is a smooth curve such that
$v(a)=v(b)=P_0$, let $T_v$ be the $2\times 2$ matrix
which represents (with respect to
$\left(\frac{\partial}{\partial x^1},\frac{\partial}{\partial x^2}\right)$ )
the holonomy transformation $T_{p_0}\R^2\to T_{p_0}\R^2$
of parallel transport around $v$.

\item[\quad] a) Show that $T_v$ is a diagonal matrix
  with determinant equal to 1.

\item[\quad] b) For $f(x^1,x^2)=x^1x^2$
let

$v_c:[0,4]\to \R^2$ by

$v_c(t)=((\ln c)t,0)$ if $0\leq t<1$,

$v_c(t)=(\ln c, t-1)$ if $1\leq t<2$,

$v_c(t)=((\ln c)(3-t),1)$ if $2<t<3$.

$v_c(t)=(0,4-t)$ if $3\leq t\leq 4$.

Find $T_{v_c}$.


\vfill
\clearpage


\end{description}
\end{large}
\end{document}


