%From dyetter@math.ksu.edu Mon Feb 23 11:20:03 1998
%Date: Mon, 23 Feb 1998 11:19:25 -0600 (CST)
%Geometry of Manifolds Qual, S98

%-*-latex-*-            This is using LaTeX mode for emacs%

\documentstyle[12pt]{article}
\topmargin -.5in
\textwidth 6.1in
\textheight 9.0 in
\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)

\begin{document}

\centerline{\large \bf Geometry of Manifolds Qualifying Examination}

\centerline{\bf Spring 1998}

\centerline{(Yetter and Miller)}
\bigskip

	This examination is in two parts:  in part A, answers with a brief
explanation will suffice; in part B, detailed calculations or proofs are
expected unless a sketch of a proof is explicitly requested.
\smallskip

\noindent{\bf Part A} Do all six (6) questions in part A. 
\smallskip

\begin{enumerate}
\item What is the fundamental group of
	\begin{enumerate}
	\item ${\bf S^3}$ (the 3-sphere)
	\item ${\bf S^2}\times {\bf S^1}$
	\item $T^*({\Bbb R}{\Bbb P}^2)$, the total 
space of the cotangent bundle 
to the real projective plane
	\end{enumerate}

\item Describe in detail the flows of the vectorfield on ${\Bbb R}^{\bf 3}$
given by

\[ (x^2 + y^2 - 4) \frac{\partial}{\partial z} \]


\item \begin{enumerate}
	\item Give an example of a simply connected compact manifold 
		with non-trivial tangent bundle.
	\item Give an example of a simply connected compact manifold 
		with trivial tangent bundle.
	\item Give an example of a non-simply connected compact manifold.
	\end{enumerate}
Note: Manifold means manifold without boundary.

\item 
\begin{enumerate}
\item If $M = {\bf S^3}$, what is the dimension of the total space
of the second exterior bundle $\Lambda ^2(M)$?
\item What are sections of the second exterior bundle called?
\end{enumerate}

\item Let $Z$ be the vectorspace of closed 1-forms on 
${\bf S^1}\times {\bf S^1}$ and $B$ be the vectorspace of exact 1-forms on
${\bf S^1}\times {\bf S^1}$. What is the dimension of $Z/B$?

\item Let ${\bf S^2} = \{(x,y,z) | x^2+y^2+z^2 = 1 \} \subset {\Bbb R}^3$.
with the induced orientation.  Calculate

\[ \int_{\bf S^2} z dx\wedge dy - y dx\wedge dz + x dy\wedge dz \]  
\end{enumerate}
\newpage

\noindent {\bf Part B.} Choose four (4) and only four of the following
problems.

\begin{enumerate}
\item Let $G$ be a Lie group
\begin{enumerate}
	\item Define its Lie algebra.
	\item Describe the correspondence between the Lie algebra and the
		1-parameter subgroups of $G$.
	\item Define the exponential mapping for $G$.
	\item Derive the from of the exponential mapping for the group of
		non-singular $n \times n$ real matrices.
\end{enumerate}

\item	Prove that ${\bf  S^n}$ for $n > 1$ is an orientable manifold.

\item 	Let $p \in {\Bbb R}{\Bbb P}^2$, and let $X = 
{\Bbb R}{\Bbb P}^2\times \{0,1\}/\equiv$
where $\equiv$ is the equivalence relation generated by $(p,0)\equiv (p,1)$.
Compute in detail $\pi_1(X)$.

\item 	Let $(M,g)$ be a Riemannian manifold, and let $X$ be a smooth 
vectorfield on $M$.  Define a function on $M$ by $f(m) = \| X(m) \|^2$
and let $Y$ be the gradient vectorfield of $f$.  Show that $X(m) = 0$
implies $Y(m) = 0$.

\item 	Let $B$ be the 2-dimensional subbundle of the tangent bundle to
${\Bbb R}^3$ whose fibre at each point is spanned by the vectorfields
$X_1 = z\frac{\partial}{\partial x} + \frac{\partial}{\partial z}$ and
$X_2 = \frac{\partial}{\partial y} + \frac{\partial}{\partial z}$.
Use the Frobenius Theorem to determine whether $B$ is integrable.  Here
$(x,y,z)$ is the standard Euclidean coordinate system.

\item Consider a 2-dimensional Riemannian manifold $(M,g)$, where $M$ is
an open subset of ${\Bbb R}^2$ and $g = f^2(dx^2 + dy^2)$ for some
positive function $f(y)$ depending only on $y$.
\begin{enumerate}
\item Write down the simplest moving orthonormal frame you can think of.

\[ e_1 = \;\;\;\;\;\;\;\;\;\; e_2 = \]

\item The moving coframe dual to $(e_1,e_2)$ is

\[ \theta^1 = \;\;\;\;\;\\;\;\;\;\; \theta^2 = \]

\item Now find a 1-form $\omega$ such that 
$d\theta^1 = \omega \wedge \theta^2$ and 
$d\theta^2 = -\omega \wedge \theta^1$.

\item Find $K$ so that $d\omega = -K \theta^1 \wedge \theta^2$.

\item What is the Gaussian curvature when 
$ds^2 = \frac{1}{y^2}(dx^2 + dy^2)$ ?
\end{enumerate} 

\end{enumerate}





\end{document}
