%From sasha@ksu.edu Wed Sep  6 10:51:01 2000
%Yetter & Miller
% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)

%\documentstyle[12pt,bbb]{report}
\documentclass[12pt,bbb]{report}

\vcorrection{-0.5in}
\hcorrection{-.9in}

\topmargin  0in
\textwidth 6.4in
\textheight 8.6in

\mathsurround=2pt

\def\ds{\displaystyle}

\begin{document}

%\pagestyle{empty}

%\begin{Large}

\begin{center} Geometry Qualifying Exam \\
               Fall 2000  \end{center}

%\vspace{.2in}

\noindent {\bf Part A.  Short answers.  Work all of the following.}

%\vspace{.2in}

\begin{description}

\item[1.] What is the fundamental group of

\begin{description}

\item[a)] $S^2 \times S^2$

\item[b)] $T^\ast (S^3 \times S^1)$  the total space of the cotangent
bundle of  $S^3 \times S^1$.

\item[c)] $R^3$  with  $2$  parallel lines deleted.

\end{description}

%\vspace{.2in}

\item[2.] A Riemannian metric on a manifold is a cross section of what
bundle?

%\vspace{.2in}

\item[3.] In  $R^3$  with the standard euclidean flat metric, describe
the flow and integral curves of a covariant-constant vector field.
(Covariant constant means all covariant derivatives vanish).

%\vspace{.2in}

\item[4.] Let  $\Gamma$  be the ellipsoid  $x^2 + \frac {y^2}4 + \frac
{z^2}9 = 1$  in  $R^3$  calculate  $\int_\Gamma zdx\wedge dy - ydz\wedge
dx$.

%\vspace{.2in}

\item[5.] Find the scalar curvature of the surface  $z = x^2 + y^2$  at
$(0, 0, 0)$.

%\vspace{.2in}

\item[6.] Can we integrate a  $3$-form on a surface in a  $4$-manifold?
Why or why not?

%\vspace{.2in}

\item[7.] \ {\bf a)} \ What is the fiber dimension of the bundle of
$5$-forms on  $S^7$?

\begin{description}

\item[b)] What is the fiber dimension of the bundle of  $7$-forms on
$S^7$?

\end{description}

\item[8.] Are all vector spaces (a) parallelizable?  Why or why not?
(b) Simply connected?  Why or why not?

%\vspace{.2in}

\item[9.] What is the scalar curvature of the euclidean plane in polar
coordinates.

%\vspace{.2in}

\end{description}

\noindent {\bf Part B.  Choose 3 (and only 3) of the following.}

%\vspace{.2in}

\begin{description}

\item[1.] Let  $G$  be a Lie group.  Prove  $G$  is orientable.

%\vspace{.2in}

\item[2.] Compute the Levi-Civita connection at a point on the standard
unit  $2$-sphere in  $R^3$  in latitude-longitude coordinates.

%\vspace{.2in}

\item[3.] Compute the De-Rham cohomology of  $S^1 \times S^2$.

%\vspace{.2in}

\item[4.] Show an explicit isomorphism between the Lie algebras  $so(3)$
and  $su(2)$.

%\vspace{.2in}

\item[5.] Let  $C$  be the  $2$-dimensional subbundle of the tangent
bundle to  $R^4$  determined by  $V_1 = x_2 \ \frac \partial{\partial
x_1} + \frac \partial{\partial x_3}$  and  $V_2 = \frac
\partial{\partial x_2} + \frac \partial{\partial x_3}$.  Use the
Frobenius theorem to determine if  $C$  is integrable.

\end{description}

%\end{Large}
\end{document}

********************************************************************************
Reta McDermott, Office Specialist            e-mail: sasha@ksu.edu 
KSU, Dept. of Mathematics                            reta@math.ksu.edu
138 Cardwell Hall                            phone:  785-532-0559
Manhattan, KS  66506-2606                    fax:    785-532-0546
********************************************************************************


