% latex file
% Geometry of Manifolds - Auckly & Miller

\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)

\input prepictex.tex
\input pictex.tex
\input postpictex.tex

%\documentstyle[bbb]{report}
\documentclass[bbb]{report}

\vcorrection{-.7in}
\hcorrection{-1in}

\topmargin  0in
\textwidth 6.4in
\textheight 9.5in

\mathsurround=2pt

\def\ds{\displaystyle}

\def\R{{\Bbb R}}
\def\Z{{\Bbb Z}}
\def\C{{\Bbb C}}
\def\D{{\Bbb D}}
\def\nd{\noindent}

\begin{document}

%\pagestyle{empty}

\begin{Large}

\hfill Name \rule{2.5in}{.01in}
\par
\vspace{.1in}

\begin{center}
   GEOMETRY OF MANIFOLDS QUALIFYING EXAM \\
   Fall 1998 \\
   (Auckly \& Miller ) \\
\end{center}


\vspace{.1in}

\begin{large}\begin{center}
\noindent {\bf Work as many as you can in the 2 hours.\\    %instructions
Best of luck.}
\end{center}\end{large}

\vspace{.05in}

\begin{description}

\item[1.] (A)
Define the deRham cohomology groups of a differential manifold.

\vspace{.1in}
\item[\quad] (B)
Determine all of the deRham cohomology groups of $S^2\times S^2$.
For those that are nonzero specify representatives for generators.



\vspace{.5in}

\item[2.]
Describe in detail the flows of the vector field
$$ -y\frac{\partial}{\partial x}
   +x\frac{\partial}{\partial y}
   -z\frac{\partial}{\partial z} \ \hbox{on}\ \R^3=
   \{(x,y,z)\vert x,y,z\in\R\}.
   $$
Describe the behavior of the orbits as $t\to +\infty$.

\vspace{.5in}

\item[3.]
On $\{(x,y)\vert x,y\in\R,0<y<\pi\}\subset\R^2$
let $g$ be the metric
$$ g= dx^2+\cos y(dx\otimes dy+dy\otimes dx)+dy^2 $$

\vspace{.1in}
\item[\quad] (A)
Compute
$\left[ x^2\frac{\partial}{\partial x}+y^2\frac{\partial}{\partial y},
 \quad xy\frac{\partial}{\partial x} \right] $

\vspace{.1in}
\item[\quad] (B)
Compute $g \left( \nabla_{\sin y\frac{\partial}{\partial y}}
  \left( \cos y\frac{\partial}{\partial x} \right),
  \frac{\partial}{\partial y} \right)$


\vspace{.5in}

\item[4.]
Let $D=\{(x,y)\in\R^2\vert x^2+y^2\leq 4\}/\sim$
where $\sim$ is the equivalence relation generated by
$(x,y)\sim(-x,-y)$ if $x^2+y^2=1$ or $x^2+y^2=4$.
Determine the fundamental group of $D$.

\vspace{.5in}
\item[5.]
Suppose that $f$ and $g:M\to N$ are smooth mappings
between two $n-\hbox{dimensional}$ manifolds and that $w$ is a
closed $n-\hbox{form}$ on $N$.
If $f$ and $g$ are homotopic show that
$\ds\int_Mf^\ast w=\ds\int_Mg^\ast w$.

\vspace{.5in}
\item[6.]
Let $B$ be a smooth vector subbundle of $TM$, the tangent bundle
of the manifold $M$.

\vspace{.1in}
\item{\quad} (A)
Define what we mean when we say that $B$ is integrable.

\vspace{.1in}
\item[\quad] (B)
State the Frobenius theorem which gives necessary and sufficient
conditions for $B$ to be integrable in terms of the bracket of
vector fields.

\vspace{.1in}
\item[\quad] (C)
Suppose that $\alpha$ is a smooth 1-form and $\alpha(m)\not= 0$
for all $m\in M$. If $B=\ds\bigcup_{m\in M} \hbox{kernel}(\alpha(m))$
show that $B$ is integrable if and only if $d\alpha\vert B=0$.


\vspace{.5in}
\item[7.]
Prove that $O_n=\{A\vert A \ \hbox{is an $n\times n$ real matrix and}
A^TA=I\}$ is a manifold. What is the dimension of $O_n$?

\item[\quad] \underbar{Hint}: Consider the mapping
$f:GL(n,\R)\to\hbox{\it symmetric matrices}$ by $f(A)=A^TA$.

\vspace{.5in}
\item[8.]
Prove or disprove the statements

\vspace{.1in}
\item[\quad] (A)
There is a Lie group $G$ which is diffeomorphic to $S^2$, the 2-sphere.

\vspace{.1in}
\item[\quad] (B)
There is a Lie group $G$ which is diffeomorphic to $S^3$, the 3-sphere.

\vspace{.5in}
\item[9.]
Prove that a simply connected 2-manifold with nonpositive
curvature can have at most one geodesic (parametrized by arc length)
from a point $A$ to a second point $B$.

\vspace{.1in}
\item[\quad] \underbar{Hint}: Consider the Gauss-Bonnet Theorem.

\vfill
\clearpage


\end{description}
\end{Large}
\end{document}


\vspace{.1in}
\item[\quad] ()
