%Real Analysis - Auckly and Miller

\documentclass[bbb]{report}
\usepackage{amsxtra,amssymb,amsthm,amsmath,latexsym}

\voffset=-1.25truein
\hoffset=-1.25truein

\topmargin  0in
\textheight 10truein
\textwidth 7in
\mathsurround=2pt

\pagestyle{empty}

\def\ds{\displaystyle}
\def\nd{\noindent}
\def\R{{\mathbb R}}
\def\Z{{\mathbb Z}}
\def\C{{\mathbb C}}
\def\D{{\mathbb D}}
\def\N{{\mathbb N}}
\def\F{{\mathcal F}}
\def\M{{\mathcal M}}

\begin{document}


%\begin{Large}
%\begin{Large}

%\hfill Name \rule{2.5in}{.01in}
%\par
%\vspace{.15in}

\begin{center}
   GEOMETRY QUALIFYING EXAM \\
  SPRING 2001
   (Auckly  \& Miller) \\
\end{center}


\vspace{.1in}


%\begin{large}
\nd {\bf
Work as many of the following problems as you can.
You do not have to solve every problem to pass.
}
%\end{large}

\vspace{.1in}

\renewcommand\baselinestretch{1.15}

\begin{description}

\item[]
\hspace{-.2truein}
\begin{tabular}{ll}
{\bf 1.} On $\R^2$ let &\\& \vspace{-.5in}\\&
$\begin{array}{ll}
X=&y\ds\frac{\partial}{\partial x}+y^2x\frac{\partial}{\partial y},\\
Y=&(x+y)\ds\frac{\partial}{\partial x} + x\frac{\partial}{\partial y}.\\
\alpha=&\,-dx+x\,dy
\end{array}$
\end{tabular}

\item[]
Calculate:\ \ a)$[X,Y]$, \ \ b)$L_Y\alpha$,\ \ c)$L_Y(d\alpha)$.

\item[2.] a)
Construct an atlas for $S^2\times S^1$.

\item[\quad] b)
Parametrize $S^2$ by $(\theta,\varphi)$ where
$x=\cos\theta\cos\varphi$,
$y=\sin\theta\cos\varphi$,
$z=\sin \varphi$.
Parametrize $S^2\times S^1$ by $(\theta_1, \varphi_1, \alpha)$
in a similar way
$(\alpha\in(\theta,2\pi))$
and define $f:S^2\times S^1\to S^2$;
$f(\theta_1,\varphi_1,\alpha)=(2\theta_1,\varphi_1)$.
Notice that $\mu=\ds\frac{1}{4\pi}\sin\varphi\,d\theta \wedge d\varphi$
is the normalized area form on $S^2$.

\item[]
Compute $\ds\int_{S^2\times S^1} \frac{1}{2\pi}d\alpha\wedge f^\ast\mu$.
(Use the orientation $d\alpha\wedge d\theta\wedge d\varphi$.)

\item[\quad] c)
Show that the answer remains unchanged if $\mu$ is replaced by
$\mu +d\omega$.

\item[\quad] d)
What does this integral represent geometrically?

\item[3.]
Let $g=dt^2+\sin^2t\,d\alpha^2+\cos^2t\,d\beta^2$.

\item[\quad] a)
Pick an orthonormal coframe, $\theta^k$, suitable for computation with the
metric, $g$.

\item[\quad] b)
Compute the connection form, $\omega$, relative to the coframe that you
chose in part a).

\item[\quad] c)
Compute the curvature form, $\Omega$, relative to the same coframe.


%\vfill\eject

\item[4.]
Let $E$ and $F$ be distributions on $(M,g)$ so that
$X\in E$ and $Y\in F$ implies $g(X,Y)=0$
and so that
$X,Y\in\Gamma(E)$ implies $\nabla_XY\in\Gamma(E)$
and
$X,Y\in \Gamma(F)$ implies that $\nabla_XY\in\Gamma(F)$.

\item[\quad] a)
State Frobenius' Theorem.

\item[\quad] b)
Show that $E$ and $F$ are integrable.

\item[\quad] c)
Show that any point of $M$ has a product neighborhood
$V\times U$ so that $g|_{V\times U}=g|_v\oplus g|_u$.


\item[5.]
Let $X$ and $Y$ be the vector fields on $\R^4-\{0\}\times\R^2$
given by
$$  \begin{array}{ll}
X=&x^2\ds\frac{\partial}{\partial x^2}+x^3\frac{\partial}{\partial x^3}\\
Y=&x^3 \ds\frac{\partial}{\partial x^1} - x^2\frac{\partial}{\partial x^4}
\end{array} $$

Let $E$ be the subbundle of the tangent bundle generated by $X$
and $Y$.

\item[\quad] a)
Use the Frobenius Theorem to show that $E$ is integrable.

\item[\quad] b)
Find parametric equations for the integral manifold containing the point
$(1,2,3,4)$.

\item[6.]a)
State Sard's Theorem.
Let $f:S^2\to\R$ be a smooth map.

\item[\quad] b)
Show that for any $y_0\in\R$ and any $\varepsilon>0$ there is a
$y\in B_\varepsilon(y_0)$ so that
$f^{-1}(y)$ is a finite disjoint union of circles.

\item[7.]a)
Give an example of a space, $X$, with $\pi_1(Y)=\langle a,b,c|a^3=1$,
$b^4=1$,
 $c^5=1$, $ (abc)^2=1\rangle$.

\item[\quad] b)
Give an example of a space, $Y$,
with $\pi_1(Y)=\langle a,b|aba^{-1}b=1\rangle$.

\item[\quad] c)
Closed, Hausdorff, separable $2\hbox{-manifolds}$ have been classified.
State the classification theorem.

\item[\quad] d)
There is a closed $2\hbox{-manifold}$ with the fundamental group
from part b). Which $2\hbox{-manifold}$ has this fundamental group?




\vfill

\end{description}

%\end{Large}

\end{document}
