% latex file
%Geometry of Manifolds S99 - Auckly & Miller

\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)

\input prepictex.tex
\input pictex.tex
\input postpictex.tex

%\documentstyle[bbb]{report}
\documentclass[bbb]{report}

\vcorrection{-.5in}
\hcorrection{-1.15in}

\topmargin  0in
\textwidth 7in
\textheight 9in

\mathsurround=2pt

\def\ds{\displaystyle}
\def\R{{\Bbb R}}
\def\Z{{\Bbb Z}}

\def\nd{\noindent}
\def\qed{\hfill\ \vbox{\hrule\hbox{\vrule\kern4pt\vbox{\kern4pt{}
\kern4pt}\kern4pt\vrule}\hrule}}


\begin{document}

%\pagestyle{empty}

\begin{Large}
\hfill Name \rule{2.15in}{.01in}
\par
\vspace{.25in}

\begin{center}
   GEOMETRY OF MANIFOLDS QUALIFYING EXAM \\
   Spring 1999 \\
   (Miller and Auckly) \\
\end{center}

\end{Large}

\vspace{.1in}

\begin{large}
\noindent {\bf
WORK THE FIRST FOUR AND AS MANY OF THE OTHERS AS YOU CAN.}

\vspace{.2in}

\begin{description}

\item[1.]
On $\R^3$ with coordinates $(x,y,z)$,  let

\item[\quad]  $X=y\ds\frac{\partial}{\partial x}
             -x\ds\frac{\partial}{\partial y}$
\item[\quad] $Y=x^2y\ds\frac{\partial}{\partial x}
             +z\ds\frac{\partial}{\partial y}
             + y\ds\frac{\partial}{\partial z}$
\item[\quad] $\alpha = (x^3+y^3+z^3)(dx\land dy+dz\land dx+dy\land dz)$

\vspace{.1in}
\item[\quad] a) calculate the Lie bracket $[X,Y]$.

\item[\quad] b) describe the flow of the vector field $Y$ through
the point $(x_0, y_0,z_0)$.

\item[\quad] c) calculate $d\alpha$ as a function times the usual
volume element.

\item[\quad] d) if $f:\R^2\to\R^3$ by $f(s,t)=(s,t,s^2+t^2)$,
compute $f^\ast\alpha$.

\vspace{.2in}
\item[2.]
On $\{(x,y)|x,y\in\R,y>0\}$ let $g$ be the metric
$g=y\,dx^2+dy^2$.
Determine the differential equations for parallel transport of a vector
field
$\xi=\xi^1\ds\frac{\partial}{\partial x}
+\xi^2\ds\frac{\partial}{\partial y}$ along a curve $x=x(t)$, $y=y(t)$.


\vspace{.2in}
\item[3.]
Let $C=\{(x,y,z)|x^2+y^2=1\}$ with the orientation
$\Omega_C=x\,dy\land dz-y\,dx\land dz$.
Compute
$$ \int_C\frac{(x+1)}{(1+z^2)(x^2+y^2)}
   (x\,dy\land dz-y\,dx\land dz) $$


\vspace{.2in}
\item[4.]
Construct a cell complex $X$, with

\item[\quad] $\pi_1(X)=\langle a,b|a^2=b^3\rangle $
\item[\quad] $H_0(X)=Z$ and $H_2(X) =Z\oplus Z_2$.


\vfill
\pagebreak


\vspace{.75in}
\item[5.]
On a Riemannian manifold define the scalar curvature to be
\ $ S=-\ds\sum_{n,k}g(R(e_n,e_k)e_n,e_k) $ \ where $(e_n)$ is
an orthonormal basis.

\item[\quad] a)
Prove that $S$ is independent of the choice of basis.

\item[\quad] b)
Let $S^n_r=\{ x\in\R^{n+1}\mid \| x\|=r\}$,
with the induced metric. Compute the scalar curvature of
$S^2_a\times S^3_b$.



\vspace{.2in}
\item[6.]
Let $M$ be a simply connected manifold, $\omega$ a closed differential
2-form, $X$ a vector field and $H$ a smooth real valued function on
$M$. Suppose they are related by $dH=i_X\omega=\omega(X,\quad )$.
Further suppose that $Y$ is a second vector field such that $L_Y\omega=0$
and $Y(H)=0$.

\item[\quad] a)
Show that there is a smooth function $f$ on $M$ such that $i_Y\omega=df$.

\item[\quad] b) Show that the function $f$ of part a) is
constant along the flows of $X$.

\item[\quad] Note: If you do not remember the formula giving the Lie
derivative $L_Y$ acting on differential forms in terms of $i_Y$ and $d$,
you may ask at the cost of a penalty.



\vspace{.2in}
\item[7.]
Let $G$ be a Lie group and $\pi_G:P\to M$ be a principal $G$-bundle,
and $H$ be a closed subgroup of $G$. We say that the structural
group of $P$ may be reduced to $H$ if and only if there is a principal
$H$-bundle $\pi_H:Q\to M$ and a bundle map $i:Q\to P$ so that
\ $i(q\cdot h)=i(q)\cdot h$.
Let $E=P\times_G(G/H)=P\times(G/H)/\sim$
 where\ $(p,[g])\sim (pk,[k^{-1}g])$, $p\in P,g,k\in G$.
Prove that $E\to M$ admits a global section if and only if the
structural group of $P$ may be reduced to $H$.

\vspace{.2in}
\item[8.]
On $\R^3$ with coordinates $(x,y,z)$ let
$\alpha=x\,dy+dz$ and $E=\{v\in T\R^3|\alpha(v)=0\}$.

\item[\quad] Prove or disprove:

\item[\quad] a) There is a codimension 2 foiliation $\cal F$
of $\R^3$ so that any leaf $N$ of $\cal F$
satisfies $TN\subseteq E\mid N$.

\item[\quad] b)
There is a codimension $1$ foiliation $\cal F$ of $\R^3$
so that any leaf of $\cal F$ satisfies $TN\subseteq E\mid N$.


\vspace{.2in}
\item[9.]
Let $S0_3=\{A:\R^3\to\R^3|A\ \hbox{is linear,}
\ A^\ast A=I\ \hbox{and}\ \det A=1\}$.
Prove that
$$\{A\in S0_3\mid A^\ast=A, A\not= I\}$$
 is a compact manifold.
What is its dimension?

\vfill
\clearpage


\end{description}
\end{large}
\end{document}
