% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Geometry Manifolds Qualifying Exam}\\ 
{\bf Spring 1996}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

Work out problem 1 and then choose 4 (and only 4) additional problems
among the remaining ones. State each problem on a new sheet of paper.
{\bf Do not turn in more than 4 additional problems}. A space always means
a topological space below.

{\bf This one is required!}

\begin{description}

\item[1.]
Answer the following questions and give a brief explanation or
counterexample:

\item[\quad] i) \quad (a)
Give 2 non-diffeomorphic examples of orientable connected 2-dimensional
compact smooth manifolds.

\item[\quad] \quad (b)
Give an example of a non-orientable connected 2-dimensional compact
smooth manifold.

\item[\quad] ii)
What is the dimension of the total space of the exterior bundle
$\Lambda^2 (M)$ if $M$ is a smooth 5-manifold?

\item[\quad] iii)
Consider the 1-form $\omega = (x^2 + y^2 - 1) dx$ on $\Bbb R^2$. Let $D$ be
the standard unit disk, $D= \{(x,y) |x^2+y^2 \leq 1 \}$. Find
$\int_D d \omega$.

\item[\quad] iv)
Let $S^3 = \{(x_1, x_2, x_3, x_4) \in \Bbb R^4 |
    x^2_1 + x^2_2+ x^2_3 + x^2_4 = 1\}$
be the unit sphere in $\Bbb R^4$.

\item[\quad] \quad (a)
What is the fundamental group of $S^3$?

\item[\quad] \quad (b)
What is the fundamental group of $S^3 \backslash \{(0,0,0,1)\}$?

\item[\quad] \quad (c)
What is the fundamental group of
$S^3 \backslash \{(0,0,0,1), (0,0,0,-1) \}$?

\item[\quad] v)
Let $M$ be a compact {\bf connected} $n$-dimensional smooth manifold.
Is there a nowhere vanishing $n$-form on $M$? What if $M$ is also
{\bf simply connected}?

\item[\quad] vi)
Let $\theta$ be a closed 5-form on $\Bbb R^6$. Is $\theta$ exact?

\item[\quad] vii)
Describe in detail the flows of the vector field
$X=(x^2+y^2-1) \frac{\partial}{\partial z}$ on $\Bbb R^3$.

\item[\quad] viii)
Let $\gamma (t)$ be an integral curve of a complete vector field $X$ on a
smooth manifold $M$. Suppose that $\gamma(t_0)=0$ for some $t_0$. What
can you say about $\gamma (t)$?

\end{description}
{\bf Choose 4 problems from below:}

\begin{description}
\item[2.]
Let $\omega$ be the 2-form on $\Bbb R^3 \backslash \{(0,0,0)\}$ given by
the formula
$$\omega = \frac{x_1 dx_2 \wedge dx_3 - x_2dx_1 \wedge dx_3 + x_3dx_1
   \wedge dx_2}{(x^2_1 + x^2_2 + x^2_3)^{\frac{3}{2}}}.$$
Show that $\omega$ is closed but not exact. (Hint: Consider the restriction
to the unit sphere $S^2 \subset \Bbb R^3$.)

\item[3.]
Compute the de Rham cohomology of $S^2$.

\item[4.]
What is the fundamental group of

\item[\quad] i)
$\Bbb R P^2$ (real projective space)?

\item[\quad] ii)
the 2-sphere $S^2$ with 3 distinct points removed?

Use generators and relations if you like.

\item[5.]
Denote by $S^1 = \{e^{it} \in \Bbb C | t \in \Bbb R\} \subset \Bbb C$
the unit circle and $\Bbb T = S^1 \times S^1$ the 2-torus. Fix any real
number $\alpha \in \Bbb R$ and let $\varphi_\alpha : \Bbb R \to \Bbb T$ be
the map defined by
$$\varphi_\alpha (t) = (e^{2\pi it}, e^{2 \pi i \alpha t}).$$

Show that the image $\varphi_\alpha (\Bbb R) \subset \Bbb T$ is either compact
or dense in $\Bbb T$.

\item[6.]
Let $M$ be a compact smooth $n$-dimensional oriented manifold without
boundary. Show that for any $(n-1)$- form $\omega$ on $M$, there is
$p \in M$ such that $d\omega(p) = 0$.

\item[7.]
Let $f:S^n \to S^n$ be a smooth map such that $f(x) = f(-x)$. Show that the
degree of $f$ is even.




\end{description}    
%\end{Large}
\end{document}














