% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Real Analysis Qualifying Exam}\\ 
{\bf Fall 1991}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

In what follows $(X, {\cal A}, \mu)$ is an arbitrary measure space and
$\lambda$ is Lebesgue's outer measure on $\R$.

\begin{description}
\item[1.] (a)
What does it mean to say that a function $f:X \to [-\infty, \infty]$
is ${\cal A}$-measurable?

\item[\quad] (b)
Use your definition given in (a) to prove that if $f$ and $g$ are two
such functions, then
$$\{x \in X : f(x) < g(x) \} \in {\cal A}$$
and
$$\{x \in X : f(x) = g(x)\} \in {\cal A}.$$

\item[2.]
Suppose $f : X \to [0, \infty]$ is ${\cal A}$-measurable. Discuss how
$\int_X f d\mu$ is defined. You should suppose your reader knows about
measure spaces and measurable functions, but has never heard of simple
functions or integrals.

\item[3.]
What is meant by (a) $\sigma$-algebra, (b) Borel set, (c) Lebesgue measurable
set? How are these three concepts related?

\item[4.]
If $\mu (X) = 1$ and $0 < p < q< \infty$, prove that
$\parallel f\parallel_p \leqq \parallel f \parallel_q$ for every
${\cal A}$-measurable $f:X \to \C$. [Hint: Apply H\"older's Inequality with
$r = q/p > 1$.]

\item[5.]
Prove (a) Fatou's Lemma by applying the Monotone Convergence Theorem, and
also (b) the Monotone Convergence Theorem by applying Fatou's Lemma. Begin
by clearly stating both results.

\item[6.]
Prove the completeness of $Lp(\mu)$ for $1 \leq p < \infty$.

\item[7.]
Suppose $(f_n)^\infty_1$ is a sequence in $L_1(\mu)$, $f \in L_1(\mu)$,
and $\parallel f_n - f\parallel_1 \to 0$. Prove that
$$\lim_{n \to \infty}\sup \int \log |f_n| d \mu \leq \int \log |f| d \mu,$$
where $\log 0 = -\infty$. [Hints: First show that $(f_n)^\infty_1$ has a
subsequence with converges pointwise a.e. and then apply Fatou's Lemma
to an appropriate subsequence of
$(|f_n| - \log |f_n|)^\infty_1$.]

\item[8.]
Suppose $\mu$ is a $\sigma$-finite measure, and ${\cal C}$ is a subcollection
of ${\cal A}$. Prove that there exists $G_0 \in {\cal A}$ such that

\item[\quad] (i)
If $B \in {\cal C}$, then $B \subset G_0$ a.e., that is,
$\mu (B \backslash G_0) = 0$;

\item[\quad] (ii)
If $G \in {\cal A}$ is another set satisfying (i), then $G_0 \subset G$
a.e.

[Hints: Let ${\cal C}^\sim$ be the collection of all countable unions of sets
chosen from ${\cal C}$. For each $E \in {\cal A}$ with
$\mu (E) < \infty$, show that $\sup\{\mu (B \cap E) : B \in {\cal C}^\sim\}$
is attained at some $G=G_E \in {\cal C}^\sim$. Finally consider an
appropriate countable union of such $G$'s.]

\item[9.]
Let $S=\{(x,y) \in \R^2 : 0 < x< 1, 0<y<1\}$ and define $f$ on $S$ by
$$f(x,y) = \frac{x^2 - y^2}{(x^2 +y^2)^2}.$$
Use Fubini's Theorem to prove that $f$ is not Lebesgue integrable over
$S$.

[Hint: $\frac{\partial}{\partial y} \, \frac{y}{x^2+y^2} = f(x,y)$.]






\end{description}    
%\end{Large}
\end{document}














