%From nagy@math.ksu.edu Thu Feb 07 13:07:57 2002

%REAL ANALYSIS QUAL SPRING 2002
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\newcommand{\Hh}{\mathbb{H} }
\newcommand{\D}{\mathbb{D} }
\newcommand{\T}{\mathbb{T} }
\newcommand{\C}{\mathbb{C} }
\newcommand{\N}{\mathbb{N} }
\newcommand{\Z}{\mathbb{Z} }
\newcommand{\St}{\mathbb{S} }
\newcommand{\R}{\mathbb{R} }






\begin{document}
\vspace*{-1in}
\hfill\hbox to 3in{Name\hrulefill}
\vspace{1in}
\begin{center}{\large Real Analysis Qualifying Exam}
\\\bigskip
{\large January, 2002}
\end{center}

\

\noindent{\bf  Instructions:}  Below you will find 8 problems.
Each problem is worth 10 points. Only the
best 6 scores will be added.

\noindent{\bf Time:} 2 hours.

\

\noindent{\sc Notations:} $\R =$ set of all real numbers;
%$\C =$ set of all complex numbers;
$\lambda =$ Lebesgue measure on $\R$;
$L^p(I,d\lambda)=$ the space of real-valued Lebesgue
measurable functions on $I$ with $\int_I|f|^p\,d\lambda<\infty$ ($I=$ any interval,
$p>0$).
%\D=\{z\in\C\,:\,|z|<1\}$ (the unit disk); $\
%T=\{z\in\C\,:\,|z|=1\}$ (the unit circle).

\

\begin{itemize}
\item[{\bf 1.}] Give an example of a set $A\subset\R$, which is not Lebesgue measurable.

\item[{\bf 2.}] Let $K$ be a compact Hausdorff space, and let
$(f_n)_{n=1}^\infty$ be a sequence of continuous real-valued functions
on $K$, such that for every $x\in K$ one has:
\begin{itemize}
\item[(i)] $f_1(x)\geq f_2(x)\geq f_3(x)\geq\dots$
\item[(ii)] $\lim_{n\to\infty}f_n(x)=0$.
\end{itemize}
Prove that $\lim_{n\to\infty }f_n=0$ uniformly.

\item[{\bf 3.}] Suppose $(f_n)_{n=1}^\infty$ is a sequence in
$L^1(\mathbb{R},d\lambda)$, such that
\begin{equation}
\lim_{n\to\infty}\int_\R |f_n|\,d\lambda=0.
\tag{$*$}
\label{f}
\end{equation}
\begin{itemize}
\item[(a)]
Prove that there exists a subsequence $(f_{n_k})_{k=1}^\infty$ of
$(f_n)_{n=1}^\infty$, with $\lim_{k\to\infty}f_{n_k}=0$, a.e.
\item[(b)] Given an example of a sequence $(f_n)_{n=1}^\infty$, satisfying
\eqref{f}, but which is not convergent a.e. to $0$.
\end{itemize}
\newpage

\item[{\bf 4.}] Let $C[0,1]$ denote the Banach space of all
real-valued continuous functions on $[0,1]$. (The norm on $C[0,1]$ is
defined by
$\displaystyle{\|f\|=\max_{t\in[0,1]}|f(t)|}$.)
\begin{itemize}
\item[(a)] Prove that any finite dimensional linear subspace $V$ of
$C[0,1]$ is closed in the norm topology.
\item[(b)] Prove that if $V$ is a linear subspace of $C[0,1]$, which
has a countable infinite linear basis, then $V$ is not closed in the
norm topology. ({\sc Hint}: Use Baire's Theorem.)
\end{itemize}

\item[{\bf 5.}]
Let $f\in L^2([-\pi,\pi],d\lambda)$ be a function with the property:
$$\int_{-\pi}^\pi f(x)\sin nx\,dx=0,\,\,\,\forall\,n\in\mathbb{N}.$$
Prove that $f(-x)=f(x)$, a.e.

\item[{\bf 6.}] Let $f\in L^1(\R,d\lambda)$, and
let $(f_n)_{n=1}^\infty$ be a sequence in $L^1(\R,d\lambda)$, with
\begin{itemize}
\item[(i)] $\lim_{n\to\infty}f_n=f$, a.e.
\item[(ii)] $\lim_{n\to\infty}\|f_n\|_1=\|f\|_1$.
\end{itemize}
Prove that $\lim_{n\to\infty}\|f_n-f\|_1=0$.

\item[{\bf 7.}] Let $f\in L^1(\R,d\lambda)$. Prove that
$$
\lim_{n\to\infty}\int_\R f(x)\sin nx\,dx=0.$$

\item[{\bf 8.}] Let $(X,\mathcal{M},\mu)$ be a $\sigma$-finite
measure space, and let $\{E_\lambda\}_{\lambda\in\Lambda}\subset
\mathcal{M}$ be a disjoint collection.
Prove that the set
$$
S=\{\lambda\in\Lambda\,:\,\mu(E_\lambda)>0\}$$
is at most countable.
\end{itemize}
\end{document}

