% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Differential Equations Qualifying Exam}\\ 
{\bf Spring 1990}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

\begin{description}
\item[1.] (i)
Formulate maximum principle for Cauchy's problem for parabolic equations.

\item[\quad] (ii)
Give an example of its applications without proof.

\item[2.]
Solve the exterior Dirichlet problem
$$\begin{cases}
        \Delta u = 0 &\hbox{\ for\ } |x| > 1 \\
        u(x) = c &\hbox{\ for\ } |x| = 1
        \end{cases}
        x \in \Bbb R^3.$$

\item[3.] (i)
State the Huygens Principle. How does it depend on the space dimension?

\item[\quad] (ii)
Define the energy integral $E(t)$ for the wave equation
$$\begin{cases}
        u_{tt} &= c^2 \Delta u \\
        u(0,x) &=\phi(x), \quad u_t (0,x) = \psi (x) \\
        u(t,x)|_{x \in \partial \Omega} &= 0
        \end{cases}$$
Show that the $E^\prime (t) =0$.

\item[4.] (i)
For $u \in L^1 (\Omega)$ define the weak derivative $D_iu$.
$(1 \leq i \leq n)$ where $\Omega \subset \Bbb R^n$ is a bounded domain.

\item[\quad] (ii)
Define the Sobolev space $W^{2,p} (\Omega)$ where $p \geq 1$ and give the
definition of the norm $\parallel u \parallel_{W^{2,p}}$.

\item[5.]
A semidisc of radius $a$ is well isolated from the surrounding medium. The
temperature at $ADB$ and segment $\overline{AB}$ are kept at
$u = c_2$ and $u=c_1$ respectively. Find the stationary temperature
distribution $u$ in the semidisc. (See Fig. 1.)

\item[6.]
Solve the following equation in the upper half-plane
$$\begin{cases}
        \Delta u (x,y) &=0, \quad -\infty < x < \infty, \quad 0< y< \infty \\
        u(x,0) &=f(x)
        \end{cases}$$

\item[7.] (i)
Give the definition of the well posed problem in the sense of Hadamard.

\item[\quad] (ii)
Prove that the solution of the Dirichlet problem
$$\begin{cases}
        \Delta u &= 0 \\
        u|_{\partial \Omega} &= f(x)
        \end{cases}$$
depends on $f$ continuously.

\item[8.]
Solve the following equation
$$\begin{cases}
        xu_y - yu_x &= u \\
        u(x,0) &=h(x)
        \end{cases}
        \quad (x,y) \in \Bbb R^2$$
where $u = u(x,y), h$ is a given function.        









\end{description}    
%\end{Large}
\end{document}














