% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Partial Differential Equations Qualifying Exam}\\ 
{\bf Fall 1989}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

\begin{description}
\item[1.]
Show that the Cauchy problem of Laplace equation
$$\begin{cases}
        u_{tt} + u_{xx} &= 0 \\
        u(0,x) &= \phi(x) \quad t > 0, x \in R^1 \\
        u_t (o,x) &= \psi (x)
        \end{cases}$$
is not well posed.

\item[2.]
A semidisc of radius $a$ is well isolated from the surrounding. The temperature
at $ADB$ and segment $\overline{AB}$ are kept at constant $c_2$ and $c_1$
respectively. Find the stationary temperature distribution $u$ in this
semidisc. (See Fig. 1.)

\item[3.] (i)
Under the boundary condition
$\left(au + b \frac{\partial u}{\partial b}\right)_{|\partial \Omega}=0$
where $a \neq 0 \neq b$ are constants, show that the Laplacian $\Delta$
is selfadjoint operator with all eigenvalue being negative and that
$-\Delta$ is positive definite operator.

\item[\quad] (ii)
Can you use the result from part (i) to show that the equation
$$\begin{cases}
        \Delta u  =f \hbox{\ in\ } \Omega \\
        \left(au + b\frac{\partial u}{\partial n} \right)_{|\partial \Omega}
        = 0
        \end{cases}$$
exists a unique solution $u$ where $f \in L^2(\Omega)$ and $a \cdot b$
are nonzero constants.

\item[4.]
Find the solution $u(x,y,z)$ for the equation $xu_x+yu_y + u_z =u$ with the
Cauchy data $u(x,y,0) = \phi(x,y)$ where $\phi$ is given $C^1$-function.

\item[5.]
Let $\phi(x)$ be a bounded function in $\R^n$, define its average over
$\R^n$ as
$$\phi_{av} = \lim_{r \to \infty} \frac{1}{|B_r|} \int_{B_r}
  \phi (x) dx$$
where $B_r$ is the ball $\parallel x \parallel \leq r$. For heat conduction
in $\R^n$ without sources, show that the average of temperature remains
constant in time.

\item[6.] (a)
Define a weak solution of $\Delta u=0$.

\item[\quad] (b)
Define a strong solution of $\Delta u=0$.

\item[\quad] (c)
Show any weak solution is also a strong solution.

\item[7.]
Let $\frac{\partial^2 u}{\partial t^2} = \frac{\partial^2u}{\partial x^2}$
and let $B_t = \{x: |x| \leq 1-t\}$ for $0 \leq t \leq 1$ let
$E(t) = \frac{1}{2} \int_{B_t}
\left[\left(\frac{\partial u}{\partial t}\right)^2 +
\left(\frac{\partial u}{\partial x}\right)^2\right] dx$.
Show $\frac{dE}{dt} \leq 0$.

\item[8.]
Find the $C^2$ function $u$ on $D=\{x^2+y^2 \leq 1\}$
satisfying $u(e^{i\theta}) = \theta^2 - \pi^2 (-\pi < \theta \leq \pi)$
that minimizes
$$\int_D |\nabla u|^2 dxdy.$$






\end{description}    
%\end{Large}
\end{document}














