%From pietro@math.ksu.edu Mon Feb  7 15:27:57 2000

\documentclass[12pt]{amsart}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-0.5in}


\newcommand{\Hh}{\mathbb{H} }
\newcommand{\D}{\mathbb{D} }
\newcommand{\T}{\mathbb{T} }
\newcommand{\C}{\mathbb{C} }
\newcommand{\N}{\mathbb{N} }
\newcommand{\Z}{\mathbb{Z} }
\newcommand{\St}{\mathbb{S} }
\newcommand{\R}{\mathbb{R} }



\newcommand{\rea}{\operatorname{Re}}
\newcommand{\ima}{\operatorname{Im}}
\newcommand{\Arg}{\operatorname{Arg}}
\newcommand{\area}{\operatorname{Area}}
\newcommand{\spec}{\operatorname{Spectrum}}
\newcommand{\inte}{\operatorname{int}}
\newcommand{\dist}{\operatorname{dist}}
\newcommand{\Span}{\operatorname{Span}}
\newcommand{\Cap}{\operatorname{Cap}}
\newcommand{\bd}{\partial }



\begin{document}
%\vspace*{-1in}
\hfill\hbox to 3in{Name\hrulefill}
\vspace{.5in}

\begin{center}{\large Complex Analysis Qualifying Exam}
\\\bigskip
{\large January, 2000}
\end{center}

\

\noindent{\bf  Instructions:}  Below you will find 8 problems.
Each problem is worth 10 points. Do as many problems as you can.

\noindent{\bf Time:} 2 hours.

\

\noindent{\sc Notations:} $\R =$ set of all real numbers;
$\C =$ set of all complex numbers; $\D=\{z\in\C\,:\,|z|<1\}$ (the unit disk); $\
T=\{z\in\C\,:\,|z|=1\}$ (the unit circle).

\

\begin{itemize}
\item[{\bf 1.}] Let $f$ be a holomorphic function on the open unit
disk $\D $ with $|f (z)|<1$ for all $z\in\D$. If $f (z_{1})=z_{1}$ and
$f (z_{2})=z_{2}$ with $z_{1}\neq z_{2}$, show that $f (z)=z$ for all $z\in\D$.


\item[{\bf 2.}] Use the Residue Formula to compute
\[
\int_{-\infty}^\infty\frac{x^2}{1+x^4}\,dx.
\]








\item[{\bf 3.}] Let $C$ be the unit circle, oriented counter
clockwise. For any $z$ in the complex plane, $|z|\neq 1$, evaluate
\[
\int_{C}\frac{\overline {\zeta }}{\zeta -z}d\zeta .
\]

\item[{\bf 4.}] How many roots (counting multiplicities) does
\[
z^{4}+3z-1
\]
have in the annulus $\Omega =\{z: 1<|z|<2 \}$.


\item[{\bf 5.}] Let $G$ be the circle in the complex plane which
passes through $0$ and $1$, such that the angle from the real axis to
$G$ is $\pi/4$ (measured counterclockwise). Let $\Omega $ be the
region inside $G$ and above the real axis. Find a conformal map of
$\Omega $ onto the upper half-plane. You may express the conformal map
as a composition of simpler maps.


\item[{\bf 6.}] Given an entire function $g$, describe all entire functions $f$
with the property:
$$
|f(z)|\leq |g(z)|,\,\,\,\forall\,z\in\C.$$



\item[{\bf 7.}] Let $\Omega\subset\C$ be an open set which contains the open
unit disk $\D$, and let $f:\Omega\to\C$ be a holomorphic function such that
$f(0)=1$ and
$$
|f(z)|>1,\,\,\,\forall\,z\in\T.$$
Prove that $f$ has at least one zero in $\D$.




\item[{\bf 8.}] Suppose $f,g:\D\to\C$ are holomorphic, without zeros. Assume
$$
\frac{f'(1/n)}{f(1/n)}=\frac{g'(1/n)}{g(1/n)},\,\,\,n=2,3,4,\dots$$
Find another simple relation between $f$ and $g$.


\end{itemize}
\end{document}
 


