% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Complex Analysis Qualifying Exam}\\ 
{\bf Fall 1985}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

\begin{description}
\item[1.]
Let $U$ be open $\subset \C, f : U \times [0,1] \to \C$ a continuous
function such that $z \mapsto f(z,t)$ is holomorphic in $U$ for each
$t \  \varepsilon \ [0,1]$.

\item[\quad] (a)
Show that then $D_1 f(z,t)$ is a continuous function on $U \times [0,1]$.

\item[\quad] (b)
Show that the function
$$F(z) = \int^1_0 f (z,t) dt, \quad z \ \varepsilon \  U$$
is holomorphic in $U$ and that
$$F^\prime (z) = \int^1_0 D_1 f(z,t) dt, \quad z \ \varepsilon \ U.$$

\item[2.]
Suppose that $f$ is holomorphic in $V \backslash \{a\}$ for some neighborhood
$V$ of $a \ \varepsilon \ \C$.

\item[\quad] (a)
Define: $f$ has a pole of order $k >0$ at $a$.

\item[\quad] (b)
Show that if $f$ has a pole of order $k$ at $a$, then Residue of
$f$ at
$a =\frac{1}{(k-1)!} \lim_{z \to a} \frac{d^k}{dz^k} \left[(z-a)^k f(z)
\right]$.

\item[3.]
If $f$ is entire and periodic, then $f$ has a fixed pint.

{\bf HINT:} Consider the range of the entire function $f(z) - z$.

\item[4.]
Say that the continuous function $f : U$ open $\subset \C \to IR$ has the
{\it restricted (circumferential) mean value} property if for each
$z \ \varepsilon \ U$ there is at least one $r=r(z) >0$ such that
$\{w \ \varepsilon \ \C : |w-z| \leq r\} \subset U$ and
$$f(z) = \frac{1}{2 \pi} \int^{2\pi}_0 f(z+ re^{i \theta}) d \theta.$$

\item[\quad] (i)
Show that if $U$ is bounded and connected and $f$ has this property, then
$f$ has no maximum in $U$ unless it is constant.

{\bf HINT:} Suppose that the subset $M$ of $U$ where the maximum is realized
is not empty. It is relatively closed in $U$ and if
$M \neq U$, there will be a point $z \ \varepsilon \ M$ closest to
$\C \backslash U$.

\item[\quad] (ii)
Suppose that $U$ is bounded region for which the Dirichlet problem is
solvable and that $g : \overline U  \to IR$ is a continuous function which
enjoys the restricted mean value property in $U$. Show that $g$ is harmonic
in $U$.

{\bf HINT:} If $h$ is the solution of the Dirichlet problem with boundary
data $g|_{\partial U}$, consider $f = h-g$ and $f= g-h$.

\item[5.]
Write an essay on ``simple connectivity for planar regions". Cover the
following aspects: topological characterizations involving both internal and
external properties; homotopy; analytic characterizations. Define the index
of a closed curve and explain its relation to simple connectivity. What
special role do Jordan curves play? Elucidate the relation to Cauchy's
integral theorem and the contributions of Bernhard Riemann and Carl Runge.
Sketch the proofs of 4 of the equivalent definitions of simple connectivity.

\item[6.]
Let $D_0 = \{0 < |z| < 1\}$ and $f,g \ \varepsilon \ H (D_0)$. Suppose that
both $fg$ and $f+g$ have poles at $z=0$. Prove that at least one on $f$
and $g$ has a pole at $z=0$.

\item[7.]
Let $f$ be holomorphic on $\{0 < |z| < 1\}$ and have an essential
singularity at $z=0$. Prove that if $g$ is an entire function such that
$g \circ f$ has a removable singularity at $z=0$, then $g$ is a constant.

\item[8.]
Prove the following form of the maximum modulus principle:

\item[\quad] (a)
If $\Omega$ is a bounded region in $\C, M \ \varepsilon \ IR$, $f$ is holomorphic
in $\Omega$ and $\overline{\lim}_{z \to w} |f(z)| \leq M$ for every
$w \ \varepsilon \ \partial \Omega$, then $|f(z)| \leq M$ for
every $z \ \varepsilon \ \Omega$.

\item[\quad] (b)
Give an example to show that the conclusion fails if the boundedness of
$\Omega$ is not hypothesized.

\item[\quad] (c)
Suppose the hypothese in (a) hold for all but one point
$w_0 \ \varepsilon \ \partial \Omega$ but there we at least have
$\overline{\lim}_{z \to w_0} |f(z)| < \infty$. Show that the conclusion
still follows.

{\bf HINT:} For $d :=$ diameter of $\Omega$ and for positive integers $n$,
apply (a), with appropriate new constants $M$, to the functions
$F_n (z) := [f(z)]^n \left(\frac{z-w_0}{d}\right)$.

\item[9.]
Calculate $\int^\infty_0 \frac{\cos bx}{(x^2 + a^2)^2} dx$,
where $a \neq 0$ and $b$ are real numbers. (Hint: WLOG: $a>0, b \geq 0$.)

\item[10.]
Consider a finite (Blaschke) product
$B(z) = \prod^n_{k=1} \frac{z-a_k}{1-\overline a_{k^z}}$,
$a_k, z \ \varepsilon \ ID := \{z \ \varepsilon \ \C : |z| < 1\}$. Show
that every point of $ID$ is taken as a value $n$ times (counted according to
multiplicity) by the function $B$.




\end{description}    
%\end{Large}
\end{document}














