%From nagy@math.ksu.edu Fri Sep 14 15:07:45 2001

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
%\pagestyle{empty}
%\textwidth=7truein\hoffset=-1in
%\textheight=10truein\voffset=-1in

\begin{document}
\vspace*{-1in}
\hfill\hbox to 3in{Name \hrulefill}
\vspace{1in}
\begin{center}{\large Complex Analysis Qualifying Exam}
\\\bigskip
{\large August, 2001}
\end{center}

\

\noindent{\bf  Instructions:}  Below you will find 8 problems.
Each problem is worth 10 points. Only the
best 6 scores will be added.

\noindent{\bf Time:} 2 hours.

\

\noindent{\sc Notations:} $\mathbb{R}=$ set of all real numbers;
$\mathbb{C}=$ set of all complex numbers; $\mathbb{D}=\{z\in\mathbb{C}\,:\,|z|<1\}$ (the unit disk); $\mathbb{T}=\{z\in\mathbb{C}\,:\,|z|=1\}$ (the unit circle).

\

\begin{itemize}
\item[{\bf 1.}] Let $f:\overline{\mathbb{D}}\to\mathbb{C}$ be a continuous
function, which is holomorphic on the open unit disk $\mathbb{D}$.
Prove that there exists a sequence $(P_n)_{n\geq 1}$ of polynomials, such that
$$P_n\xrightarrow[\,\,n\to\infty\,\,]{}f\text{, uniformly on the {\em whole\/}
closed unit disk }\overline{\mathbb{D}}.$$
{\footnotesize
{\sc Hint:} For every $r<1$ define the function $f_r(z)=f(rz)$. Prove first
that
$$f_r\xrightarrow[\,\,r\to 1\,\,]{}f\text{, uniformly on }\overline{\mathbb{D}}.$$
Then use the fact that each $f_r$ is holomorphic on an open disk $D_r$ that
contains $\overline{\mathbb{D}}$.}

\item[{\bf 2.}]
Let $\Omega\subset\mathbb{C}$ be an open set,
which contains the closed unit disk $\overline{\mathbb{D}}$, and
let $f:\Omega\to\mathbb{C}$ be a holomorphic function which satisfies
$$
|f(z)|>|f(0)|\text{, for all }z\in\mathbb{T}.$$
Prove that $f$ has at least one zero in $\mathbb{D}$.

\item[{\bf 3.}]
Suppose $f,g:\mathbb{D}\to\mathbb{C}$ are holomorphic and
$$|f(z)|=|g(z)|\text{, for all }z\in\mathbb{D}$$
Prove that every zero of $g$ is also a zero of $f$, having the {\em same\/}
multiplicity. Infer that there exists some $\lambda\in\mathbb{T}$, such that
$$f(z)=\lambda g(z)\text{, for all }z\in\mathbb{D}.$$

\item[{\bf 4.}]
Compute
$$
\int_0^\infty \frac{dx}{1+x^7}.$$
{\footnotesize
{\sc Hint:} For (large) $R>1$, use the boundary of the circular sector
$$C_R=\{re^{i\theta}\,:\,0<r<R,\,\,0<\theta<2\pi/7\}$$
as contour.}

\item[{\bf 5.}] Let $\Omega\subset\mathbb{C}$ be an open set,
which contains the closed unit disk $\overline{\mathbb{D}}$, and
let $f:\Omega\to\mathbb{C}$ be a holomorphic function which satisfies
$$f(z)\neq 0\text{, for all }z\in\mathbb{T}.$$
Prove that the number
$$
Z_f:=\frac 1{2\pi i}\oint_{\mathbb{T}}\frac{f'(z)}{f(z)}\,dz$$
is equal to the number of zeros (counted with multiplicities)
of $f$ in $\mathbb{D}$. (In the above line
integral, the unit circle $\mathbb{T}$ is
oriented counterclockwise and traversed.)


\item[{\bf 6.}] Give an example of a harmonic function $h:\mathbb{C}\smallsetminus\{0\}
\to\mathbb{R}$, which is not the real part of any holomorphic function
$f:\mathbb{C}\smallsetminus\{0\}\to\mathbb{C}$.

\item[{\bf 7.}] How many zeros (counted with multiplicities) does
$$f(z):=z^4+3z-1$$
have in the annulus
$\Omega=\{z\in\mathbb{C}\,:\,1<|z|<2\}$?


\item[{\bf 8.}]
Let $\Omega\subset\mathbb{C}$ be an open set, let $a\in\Omega$, and let
$f:\Omega\smallsetminus\{a\}\to\mathbb{C}$ be a holomorphic  function. Prove
that the following are equivalent:
\begin{itemize}
\item[(i)] $a$ is an essential singularity of $f$.
\item[(ii)] There is no integer $n\geq 1$ such that
$$\lim_{z\to a}(z-a)^nf(z)$$
exists in $\mathbb{C}$.
\end{itemize}

\end{itemize}
\end{document}
