% latex file
%Complex Variables qual S99 - Burckel & Nagy

\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\input amssym.def   %defines Bbb and gothic (frak)
\input amssym       %defines Bbb and gothic (frak)

\input prepictex.tex
\input pictex.tex
\input postpictex.tex

%\documentstyle[bbb]{report}
\documentclass[bbb]{report}

\vcorrection{-.5in}
\hcorrection{-1.15in}

\topmargin  0in
\textwidth 7in
\textheight 9in

\mathsurround=2pt

\def\ds{\displaystyle}
\def\R{{\Bbb R}}
\def\Z{{\Bbb Z}}
\def\T{{\Bbb T}}
\def\O{{\Bbb 0}}
\def\D{{\Bbb D}}
\def\U{{\Bbb U}}
\def\C{{\Bbb C}}
\def\N{{\Bbb N}}


\def\nd{\noindent}
\def\qed{\hfill\ \vbox{\hrule\hbox{\vrule\kern4pt\vbox{\kern4pt{}
\kern4pt}\kern4pt\vrule}\hrule}}


\begin{document}

%\pagestyle{empty}

\begin{Large}
\hfill Name \rule{2.15in}{.01in}
\par
\vspace{.25in}

\begin{center}
   COMPLEX VARIABLES QUALIFYING EXAM \\
   Spring 1999 \\
   (Burckel and Nagy) \\
\end{center}

%\end{Large}

\vspace{.1in}

%\begin{large}

\begin{center}
{\bf
$\N:=$ natural numbers (positive integers),
$\R:=$ reals,
$\C:=$ complexes,

$\D:=\{z\in\C:|z|<1\}$,
$\U:= \{z\in\C: Im\,z>0\}$.
}

\vspace{.2in}
{\bf Do any 8 of the 10 problems.}

\end{center}


\vspace{.2in}

\begin{description}

\item[1.]
Compute $\ds\sum^\infty_{n=1}\ds\frac{1}{n^4}$.

\item[\quad] {\it Hint}:
Integrate $f(z):=\ds\frac{\pi\cot\pi z}{z^4}$ over bigger and bigger
circles.


\vspace{.2in}
\item[2.]
Suppose $f:\D\to\C$ is analytic, $f(0)=1$, and
$|f(z)|<\ds\frac{1}{|z|}$ for all $z\in\D$, $z\not= 0$.
Prove that $f$ is identically equal to $1$.

\item[\quad]  {\it Hint}: Schwarz.


\vspace{.2in}
\item[3.]
Show that the function $u:\R^2\setminus\{(0,0\}\to\R$,
defined as $u(x,y):=\ln\left(\ds\sqrt{x^2+y^2}\right)$
is harmonic, but has no harmonic conjugate.

\vspace{.2in}
\item[4.]
Let $f$ be an entire function. Define $u:\R^2\to\R$ by
$u(x,y)= Re\,f(x+iy)$.
Prove that if $u$ is a polynomial function (in two variables),
then $f$ is a polynomial.


\vspace{.2in}
\item[5.]
Compute: \quad $\ds\int^{2\pi}_0 e^{-i\theta} e^{e^{i\theta}} d\theta$.

\vfill
\pagebreak

\vspace{.2in}
\item[6.]
Prove the following strong form of Morera's theorem:
If $f$ is continuous in $\D$ and $\int_{\partial R}f=0$ for every
rectangle $R$ lying in $\D$ and having sides parallel to the
coordinate axes, then $f$ is holomorphic.

\item[\quad]  {\it Hint}: Via an integral, construct an $F$ satisfying
$D_1F=-iD_2F=f$.


\vspace{.2in}
\item[7.]
$f$ is continuous on $\overline{\D}$, holomorphic in $\D\setminus\R$.
Show that $f$ is holomorphic in $\D$.

\item[\quad]  {\it Hint}: Use preceding exercise
(whether or not you did it).


\vspace{.2in}
\item[8.]
$f$ is holomorphic in $\D\setminus\{0\}$ with a pole at $0$.
Show that the range of $f$ contains the complement of a compact disk.

\item[\quad] {\it Hint}: Show that $1/f$ is
holomorphic near $0$ and apply the Open Mapping Theorem.

\vspace{.2in}
\item[9.]
Show that if $f$ is holomorphic in a deleted neighborhood of $a$,
then $f'$ cannot have a {\it simple} pole at $a$.


\vspace{.2in}
\item[10.]
Use Liouville's Theorem to show that the range of every
non-constant entire function is dense in $\C$. What is the
strongest conclusion about the size of the image set that is valid?



\vfill
\clearpage


\end{description}
\end{Large}
\end{document}
