% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in
\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Complex Analysis Qualifying Exam}\\ 
{\bf Spring 1992}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\
$\Bbb R$ is the set of all real numbers, $\Bbb C$ the set of all complex
numbers, $D(a,r) := \{z \in \Bbb C : |z-a| < r\}$ for any
$a \in \Bbb C$, $0 < r \in \Bbb R$, $\Bbb D := D(0,1)$.

\begin{description}
\item[1.]
$f:\Bbb D \to \Bbb D$ is holomorphic. Show that
$$\left| \frac{f(x) - f(w)}{1- \overline{f(z)} f(w)} \right| \leq
  \left| \frac{z-w}{1-\overline z w} \right| \quad
  \forall z, w \in \Bbb D.$$

What can you say about $f$ if there exists a pair $(z,w)$ for which
$z \neq w$ and equality holds above?

\item[2.]
Let $f$ be a one-to-one holomorphic function on a region $\Omega$. Show
that $f^\prime$ is zero-free.

\item[3.] (a)
$f$ is continuous on $\overline{\Bbb D}$, holomorphic in
$\overline{\overline{\Bbb D}}$.
Show that $f$ is uniformly approximable on $\overline{\Bbb D}$ by
polynomials.

\item[\quad] (b)
State and prove the converse of (a).

\item[4.]
Let $A$ denote the algebra of continuous functions on $\overline{\Bbb D}$
which are holomorphic in $\Bbb D$. Find all $\Bbb C$-algebra homomorphisms
$\Phi : A \to \Bbb C$ which are not identically 0. {\bf HINT:} Show first
that $\Phi (f) \in f(\overline{\Bbb D})$ for every $f$, then see what
$\Phi$ does to polynomials. Is $\Phi$ continuous?

\item[5.] (a)
State the Maximum Modulus Principle for holomorphic functions, and

\item[\quad] (b)
Give two different proofs of this principle.

\item[6.]
Suppose that $f$ is holomorphic on $\overline{\Bbb D}$, $|f(z)| < 1$
whenever $|z| = 1$, and that $\alpha \in \Bbb D$. Find the number of
solutions in $\overline{\Bbb D}$ of the equation
$$f(z) = \left( \frac{z-\overline \alpha}{\alpha z -1} \right)^2.$$

\item[7.]
Suppose $f$ is holomorphic in the disc $D (a, R)$. Prove that
$$|f(a)| \leq \frac{1}{\pi R^2} \int \int_{D(a,R)} |f(x + iy) | dxdy.$$
Is this inequality valid for harmonic functions? {\bf HINT:} Polar
coordinates.

\item[8.](a)
State and prove Harnack's Inequalities.

\item[\quad] (b)
Using (a) [whether or not you proved it] demonstrate the following: If
$h : \Bbb C \to \Bbb R$ is harmonic but {\bf not} constant, and
$$m(r) := \min_{\theta} h(re^{i \theta}), \quad
  M(r) := \max_\theta h (re^{i \theta}),$$
then
$$\overline{\lim_{r \to \infty}} \frac{m(r)}{r} < 0 < \lim_{r \to \infty}
  \frac{M(r)}{r}.$$

\item[\quad] (c)
For the $h$ in (b), infer from (b) [whether you proved (b) or not] that
$h$ must have at least one zero, then go on to prove that $h(\Bbb C) = \Bbb R$.





\end{description}    
%\end{Large}
\end{document}














