% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Complex Analysis Qualifying Exam}\\ 
{\bf Fall 1983}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

\begin{description}
\item[1.]
Prove that if $f : \C \to \C$ is a continuous function such that $f$ is
analytic on $\C \backslash [-1,1]$, then $f$ is an entire function.

\item[2.]
Let $f(z) \equiv \sum^\infty_{n=1} \frac{1}{n} z^n, z \varepsilon \C$.

\item[\quad] a)
For which values of $z$ does $\sum^\infty_{n=1} \frac{1}{n} z^n$ converge?

\item[\quad] b)
Find the largest region on which $f$ is analytic.

\item[\quad] c)
Can $f$ be analytically continued to domain including $z=1$? {\it You must
prove all your answers}.

\item[3.]
Let $f$ be analytic on $D_2 = \{z \varepsilon \C : |z| < 2\}$. Suppose
$|f(z)| < 1$ if $|z| = 1$. Show that $f$ has exactly one fixed point
inside the unit circle.

\item[4.]
Use Cauchy's Residue Theorem to evaluate
$\sum^\infty_{n=0} \frac{1}{(2n + 1)^2}$.

\item[5.] (a)
Prove that for $n = 1,2, \dots$
$$\frac{1}{\sqrt{n}} = \frac{2}{\sqrt{\pi}} \int^\infty_0 e^{-nt^2} dt$$
by using the identity $\Gamma (z) \Gamma (1-z) = \frac{\pi}{\sin \pi z}$,
$z$ not an integer.

\item[\quad] (b)
Find an analytic extension of
$f(z) = \sum^\infty_{n=1} \frac{z^n}{\sqrt{n}}$,
$|z| < 1$  to $\C \backslash [1, \infty)$. {\it Justify all steps.}

\item[6.]
Let $P$ and $Q$ be polynomials with (degree $Q$) $\geq$ (degree $P$) $=2$.
Calculate the sum of the residues of $P/Q$.

\item[7.]
Consider the class of all entire functions $F$ such that if $f \varepsilon F$
then for all $z \varepsilon \C$.
$$f(z) = f(z+1)$$
and
$$f(z) = f(z+i).$$

Characterize $F$.

\item[8.]
Find a harmonic function $u(x,y)$ in $D = \{z \varepsilon \C : |z| < 1\}$
with boundary values $u(x,y) = x^2$.

\item[9.] (i)
Give an example (as simple as possible) of an entire function with zeros
precisely at:

\item[\qquad] (a)
$n!, n = 1,2, \dots$

\item[\qquad] (b)
$\sqrt{n} \log n, n = 2,3, \dots$

\item[\qquad] (c)
$\log n, n = 2,3, \dots$

\item[\quad] (ii)
Give an example of an entire function of order one with zeros precisely
at the positive even integers.

{\bf Justify all answers!}

\item[10.]
Prove that an entire function of order $3/2$ must have infinitely many zeros.





\end{description}    
%\end{Large}
\end{document}














