% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Numerical Analysis Qualifying Exam}\\ 
{\bf Spring 1988}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

\begin{description}
\item[1.]
Let $A$ be an $n \times n$ matrix, and let $A^\circ$ be the conjugate
transpose of $A$, and $\rho(A)$ be the spectral radius of $A$. Recall that
$$\parallel A \parallel_n = \sup_{\parallel x \parallel_n = 1}
  \parallel Ax \parallel_n \hbox{\ and\ } \parallel x \parallel_n =
  \left\{\sum^n_{i=1} |x_i|^n \right\}^{\frac{1}{n}}$$

Prove:

\item[\quad] (a)
$$\parallel A \parallel_1 = \max_j \sum^n_{i=1} |a_{ij}|.$$

\item[\quad] (b)
$$\parallel A \parallel_2^\circ = \sqrt{\rho(A^\circ A)}.$$

\item[2.]
Let $A$ be a strictly diagonally dominant matrix. Prove the Jacobi iteration
method converges for $A$ independently of the choice of initial vector.

\item[3.]
Consider the application of the iteration
$$z_{k+1}= \frac{z^2_k + 2}{3}$$
to the equation $x^2 - 3x + 2 = 0$.

\item[\quad] (a)
Does the method converge? Why?

\item[\quad] (b)
Show that $x_k \to 1$ as $k \to \infty$ if $-2 < z_0 < 2$.
(First prove that $z_{k+1}$ is between $z_k$ and 1 when $k \geq 1$.)

\item[\quad] (c)
Show that $z_{k+1} = 2$ if $z_0 = \pm2$, $k=0,1, \dots$, but that the
convergence to the root $\alpha = 2$ for any other value of $z_0$ is
impossible.

\item[4.] (a)
Define a Householder transformatin.

\item[\quad] (b)
Show that for any square matrix $A$, there is a unitary matrix $U$ such
that $U^0 AU$ is upper triangular. ($U^0$ is the conjugate transpose
of $U$).

\item[5.]
Suppose $A=(A_{ij})$ is an $n \times n$ matirx and that
$$|a_{ij}| \geq C \sum_{j \neq i} |a_{ij}|$$
for $i = 1, \dots, n$ where $C>1$. Let
$K(A) = \rho (A) \rho(A^{-1})$ where $\rho(A)$ is the spectral radius of
$A$. Find a bound for $K(A)$ in terms of $C$ and the $a_{ij}$'s.

\item[6.]
State the Weierstrase approximation theorem, define Bernstein polynomials, and
explain how the latter are used in the proof of the former.

\item[7.]
Assume
$$f(0) = 1, f(1) = 3, f(2), \hbox{\ and\ } f(3) = 1.$$

\item[\quad] (a)
Write a formula for the polynomial, $P(x)$ of degree at most 3 that
interpolates $f$ at these points.

\item[\quad] (b)
Derive an expression for the error $f(x) - P(x)$ at an arbitrary point $x$.
Be sure to indicate the conditions assumed on $f$.

\item[8.]
Prove: Of all nth degree monic polynomials $P_n(x)$, The Chebyshev
polynomial
$$T_n (x) = \frac{1}{2^n} \cos (n \cos^{-1} x)$$
has the smallest maximum norm on the interval $[-1,1]$, that is,
$$\max_{-1 \leq x \leq 1} |T_n(x) | \leq \max_{-1 \leq x \leq 1}
 |P_n(x)|.$$

\item[9.]
Assume that $f(x)$ has a contiuous fourth derivative on an open interval
containing the interval $[c-h, c+h]$. Derive an expression for the error
in approximating $f^{\prime \prime} (c)$ by
$$\frac{f(c+h) + f(c-h) - 2f (c)}{h^2}.$$

\item[10.]
Use undetermined coefficients to find coefficients $H_1, H_2$, and $H_3$
so that the approximation
$$\int^b_c f(x) dx \approx H_1 f \left(\frac{2a + b}{3} \right) + H_2f
  \left(\frac{a+b}{2} \right) + H_3f \left( \frac{a+3b}{4} \right)$$
is exact for polynomials of degree less than or equal to 2.

\item[11.]
Show that the numberical integration method
$$y_{n+1} = y_{n-3} + \frac{4h}{3} \left(2y^\prime_n - y^\prime_{n-1} +
  2y^\prime_{n-2} \right)$$
to approximating the solution to the ordinary differential equation
$y^\prime = f(x,y), y_o = y(x_0)$ is exact for polynomials of degree 0 or 1.
Is the method stable? Explain.




\end{description}    
%\end{Large}
\end{document}














