% latex file
\def\hcorrection#1{\advance\hoffset by #1 }
\def\vcorrection#1{\advance\voffset by #1 }

\documentclass{article}
\usepackage{my,amsxtra,amssymb,amsthm}

\vcorrection{-1.0in}
\hcorrection{-0.8in}
\textwidth 6.0in
\textheight 9.0in

\def\R{\mathbb R}
\def\C{\mathbb C}
\def\N{\mathbb N}
\def\Z{\mathbb Z}
\def\Q{\mathbb Q}

\begin{document}
%\begin{Large}






\begin{center}\begin{LARGE}
{\bf Numerical Analysis Qualifying Exam}\\ 
{\bf Fall 1988}\\ \end{LARGE}
\end{center}
\vspace{0.1in}
\noindent\hrulefill\\

\begin{description}
\item[1.]
State and prove a theorem giving necessary and sufficient conditions for the
series
$$\sum^\infty_{k=0} A^k$$
to converge, where $A$ is a real matrix.

\item[2.]
Suppose we say $x^\ast$ is a good approximation to $x$ if the relative
error $(x^\ast - x)/x$ is much less than 1.

\item[\quad] (a)
if $x^\ast$ is a good approximation to $x$ and $y^\ast$ is a good approximation
to $y$ and all of $x^\ast, x, y^\ast, y$ are nonzeros, show that
$x^\ast/ y^\ast$ is a good approximation to $x/y$.

\item[\quad] (b)
if $x^\ast$ is a good approximation to $x$ and $y^\ast$ is a good
approximaiton to $y$, explain why $x^\ast - y^\ast$ may not be a good
approximation to $x-y$. Hint: A good way of stating that $x^\ast$ is a good
approximation to $x$ is to say
$x^\ast = x(1+ \delta)$ where $|\delta| << 1$.

\item[3.]
Of all polynomials of degree $\leq 1$, find the one which best approximates
$f(x) = x^2 + 1$ on the interval $[0,1]$, in the $\infty$-norm. Be sure
to justify your answer (by stating and using an appropriate theorem).

\item[4.]
Tell whether the following matrix is positive definite.
$$\left[\begin{array} {cccc}
        8&-2&3&3 \\
        -2&6&1&1 \\
        3&1&9&0 \\
        3&1&0&8
        \end{array} \right]$$

\item[5.]
Suppose we want to use the Gauss-Seidel iteration to solve $Ax=b$ where
$$A=\left[ \begin{array} {cccc}
        5&3&1&1 \\
        2&6&3&1 \\
        1&1&5&2 \\
        1&1&4&6
        \end{array} \right]$$
Will the Gauss-Seidel iteration converge? (justify your answer).

\item[6.]
Suppose that $x = g(x)$ has a root at $x = \alpha$ and in the interval
$$|x - \alpha| < \rho \eqno{(1)}$$
$g(x)$ satisfies
$$|g(x) - g(\alpha)| \leq \lambda |x - \alpha|,$$
with $\lambda < 1$.

\item[\quad] (a)
Prove that for the iteration scheme
$$x_{k+1} = g(x_k), k = 0,1,2, \dots,$$
with any $x_0$ in (1),

\item[\qquad] (i)
all the iterates $x_k$ lie in the interval (1),

\item[\qquad] (ii)
the iterates $x_k$ converge to $\alpha$.

\item[\quad] (b)
Consider the perturbed iteration scheme
$$X_{k+1} = g(X_k) + \delta_k, k=0,1,2, \dots,$$
where $\delta_k$ satifies
$$|\delta_k| \leq \delta, k =0,1,2, \dots$$
Prove that if $X_0$ is any point in the interval
$$|x-\alpha | \leq \rho_0,$$
where $\rho_0$ satisfies
$$0 < \rho_0 \leq \rho - \frac{\delta}{1-\lambda},$$
then the iterates $X_k$ satisfy
$$|\alpha - X_k| \leq \rho,$$
and
$$|\alpha - X_k| \leq \frac{\delta}{1-\lambda} + \lambda^k \left( \rho_0 -
  \frac{\delta}{1-\lambda} \right),$$
and $\lambda^k \to 0$ and $k \to \infty$.

\item[7.]
Let $A$ be a symmetric, positive definite real matrix, $b \in R^n$, and
define $f: R^n \to R$ by
$$f(x) = \frac{1}{2} x^T Ax - b^Tx.$$
Show that $f$ has a unique minimum and that it occurs where $x$ satisfies
$$Ax = b.$$

\item[8.]
Let $V$ be a real vector space with inner product $<>$. Let $E \subset V$
be a finite dimensional linear subspace. Prove: if $v \in V$, there is a
unique $e^\ast \in E$ such that for all $e \in E$ we have
$$\parallel v-e^\ast \parallel \leq \parallel v - e \parallel,$$
where $\parallel z \parallel = \sqrt{<z,z>}$.

\item[9.]
Suppose $I(f) = Af(a) + Bf(b)$ is a 2-point quadrature rule on $C[-1,1]$.
Suppose further that $I$ has degree of precision 3. Find $A, a, B$, and $b$.

\item[10.]
Let $f \in C[a,b]$ and for $n \geq 0$ let $Q_n(x)$ be the least-squares
polynomial for $f$ of degree $\leq n$. Recall this means of all polynomials
$P(x)$ of degree $\leq n$
$$\int^b_a |f(x) - Q_n(x)|^2 dx \leq \int^b_a |f(x) - P(x)|^2 dx.$$

Prove:$\lim_{n \to \infty} \int^b_a |f(x) - Q_n (x)|^2 dx = 0$.

Hint: if $a_n = \int^b_a |f(x) - Q_n(x)|^2 dx$, show the sequence
$\{a_n\}_n$ is non-increasing. The Weierstrass Approximation Theorem may then
be helpful.

\item[11.]
Let $I(f) = \sum^n_{i=1} w_i f(x_i)$ be an n-point quadrature rule on
$[a,b]$. Show that $I$ cannot give the exact answer for every polynomial
of degree $\leq 2n$; this says an n-point rule cannot have degree of
precision $\geq 2n$.




\end{description}    
%\end{Large}
\end{document}














